### 2024-05-18 actuarial.qmd
20240518 - placeholders

# Human Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}