### 2024-05-18 fantasy-football.qmd
20240518 - add video introductions

# Intro to Football and Fantasy {#sec-introFantasy}

This chapter provides a brief primer on football and fantasy football.
If you are already familiar with fantasy football, feel free to skip this chapter.

## Football {#sec-football}

::: {.content-visible when-format="html:js"}

Here is a video that provides a brief introduction to American football:

![Video that provides introduction to American football.](images/football.mp4){width=100%}

:::

## Fantasy Football {#sec-fantasyFootball}

### Overview of Fantasy Football {#sec-fantasyOverview}

As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]
The participants are in charge of managing and making strategic decisions for their imaginary team to have the best possible team that will score the most points.
Thus, the participants are called "managers".
Managers make decisions such as selecting which players to draft, selecting which players to play (i.e., "start") on a weekly basis, identifying players to pick up from the remaining pool of available players (i.e., waiver wire), and making trades with other teams.
Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.

::: {.content-visible when-format="html:js"}

Here is a video that provides a brief introduction to fantasy football:

![Video that provides introduction to fantasy football.](images/fantasyFootball.mp4){width=100%}

:::

### The Fantasy League {#sec-league}

A fantasy football "league" is composed of various imaginary (i.e., "fantasy") teams—and their associated manager.
In the fantasy league, the managers' fantasy teams play against each other.
A fantasy league is commonly composed of 8, 10, or 12 fantasy teams, but leagues can have more or fewer teams.

### The Roster of a Fantasy Team {#sec-roster}

On a given roster, a manager has a "starting lineup" and a "bench".
Each week, the manager decides which players on their roster to put in the starting lineup, and which to keep on the bench.
In many leagues, a starting lineup is composed of offensive players, a kicker, and defense/special teams:

<u>Offensive players:</u>

| Position           | Typical Number of Players in Starting Lineup |
|--------------------|----------------------------------------------|
| Quarterback (QB)   | 1                                            |
| Running Back (RB)  | 2                                            |
| Wide Receiver (WR) | 2                                            |
| Tight End (TE)     | 1                                            |
| Flex Position      | 1                                            |

A "flex position" is a flexible position that can involve a player from various positions: e.g., a RB or WR or TE.

<u>Kickers:</u>

- one Kicker (K)

<u>Defense/Special Teams:</u>

- one Team Defense (DST/D/DEF) or multiple Individual Defensive Players (IDP)

### Scoring {#sec-scoring}

#### Scoring Overview {#sec-scoringOverview}

In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
Managers receive points for only those players who are on their starting lineup (not players on their bench).
A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
Scoring settings can differ from league to league.

Below are common scoring settings for fantasy leagues.

#### Offensive Players {#sec-scoringOffense}

| Statistical category                              | Points               |
|---------------------------------------------------|----------------------|
| Rushing or receiving TD                           | 6                    |
| Returning a kick or punt for a TD                 | 6                    |
| Returning or recovering a fumble for a TD         | 6                    |
| Passing TD                                        | 4                    |
| Passing INT                                       | −2                   |
| Fumble lost                                       | −2                   |
| Rushing, passing, or receiving 2 point conversion | 2                    |
| Rushing or receiving yards                        | 1 point per 10 yards |
| Passing yards                                     | 1 point per 25 yards |

Note: "TD" = touchdown; "INT" = interception

Other common (but not necessarily standard) statistical categories include:

- receptions (called "point per reception" [PPR] leagues)
- return yards
- passing attempts
- rushing attempts

#### Kickers {#sec-scoringKickers}

| Statistical category                              | Points |
|---------------------------------------------------|--------|
| FG made: 50+ yards                                | 5      |
| FG made: 40–49 yards                              | 4      |
| FG made: 39 yards or less                         | 3      |
| Rushing, passing, or receiving 2 point conversion | 2      |
| Point after attempt made                          | 1      |
| Point after attempt missed                        | −1     |
| Missed FG: 0–39 yards                             | −2     |
| Missed FG: 40–49 yards                            | −1     |

Note: "FG" = field goal

#### Team Defense/Special Teams {#sec-scoringDST}

| Statistical category          | Points |
|-------------------------------|--------|
| Defensive or special teams TD | 3      |
| Interception                  | 2      |
| Fumble recovery               | 2      |
| Blocked punt, PAT, or FG      | 2      |
| Safety                        | 2      |
| Sack                          | 1      |

Note: "PAT" = point after touchdown; "FG" = field goal

#### Individual Defensive Players {#sec-scoringIDP}

| Statistical category  | Points |
|-----------------------|--------|
| Tackle solo           | 1      |
| Tackle assist         | 0.5    |
| Tackle for loss       | 1      |
| Sack                  | 2      |
| Interception          | 4      |
| Fumble force          | 2      |
| Fumble recovery       | 2      |
| TD                    | 6      |
| Safety                | 2      |
| Pass defended         | 1      |
| Block kick            | 2      |
| Extra point returned  | 2      |

Note: "TD" = touchdown

Other common (but not necessarily standard) statistical categories include:

- turnover return yards

#### Common Scoring Abbreviations {#sec-scoringAbbreviations}

- "TD" = touchdown
- "INT" = interception
- "yds" = yards
- "ATT" = attempts
- "2 pt conversion" = two-point conversion
- "FG"
- "PAT" = point after touchdown (i.e., extra point/point after attempt)