### 2024-05-25 bugs.qmd
20240525 - bugs to report

# Bugs {#sec-bugs}

Bugs to report (`nflreadr`):

- `players` database has a season column but it is only populated for some players, and it is only populated with 2020
- `player_stats` database of weekly stats does not have a `game_id` column