### 2024-05-28 fantasy-football.qmd
20240527 - glossary

# Intro to Football and Fantasy {#sec-introFantasy}

This chapter provides a brief primer on (American) football and fantasy football.
If you are already familiar with fantasy football, feel free to skip this chapter.

## Football {#sec-football}

Football is the most widely watched sport in the United States.^[<https://news.gallup.com/poll/610046/football-retains-dominant-position-favorite-sport.aspx> (archived at <https://perma.cc/X2UG-RAAK>); <https://www.statista.com/statistics/1430289/most-watched-sports-leagues-usa/> (archived at <https://perma.cc/JNU6-S96A>)]

### The Objective {#sec-footballObjective}

The goal in football is for a team to score more points than their opponent.
A game lasts 60 minutes, and it is separated into four 15-minute quarters.
The team with the most points when the time runs out wins.

::: {.content-visible when-format="html:js"}

Here is a video that provides a brief introduction to American football:

![Video that Provides An Introduction to American Football.](images/football.mp4){width=100%}

:::

### The Roster {#sec-footballRoster}

#### Overview {#sec-footballRosterOverview}

Each team has 11 players on the field at a time.
The particular players who are on the field will depend on the situation, but usually includes one of the three subsets of players:

1. Offense
1. Defense
1. Special Teams

An example formation is depicted in Figure @fig-footballFormation.

::: {#fig-footballFormation}
![](images/footballFormation.png)

An Example Football Formation for the Offense and Defense. The solid line indicates the line of scrimmage. The arrow indicates the direction the offense tries to advance the ball.
:::

#### Offense {#sec-footballRosterOffense}

The offense is on the field when the team has the ball.

Players on offense include:

- Quarterback (QB)
- Running Back (RB)
    - Halfback (HB) or Tailback (TB)
    - Fullback (FB)
- Wide Receiver (WR)
- Tight End (TE)
- Offensive Linemen (OL), part of the "Offensive Line"
    - Center (C)
    - Offensive Guard (OG)
    - Offensive Tackle (OT)

The quarterback is the most important player on the offense.
They help lead the team down the field.
The quarterback receives the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
Some quarterbacks are fast and are considered "dual threats" to pass or run.

Running Backs take a hand-off from the Quarterback to execute a running play (i.e., a rush).
They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
Halfbacks and Tailbacks tend to be quick and agile.
Fullbacks tend to be strong and powerful.

Wide Receivers catch passes from the Quarterback to execute a passing play.
On running plays, they provide protection for the player running the ball (e.g., the Running Back) so the ball carrier can get as far as possible without being tackled.
Wide receivers tend to be tall, fast, have good hands (can catch the ball well), and can jump high.

Tight Ends block for running and passing plays, and they catch passes from the Quarterback.
Tight ends tend to be strong and have good hands.

Offensive Linemen block for running and passing plays.
On passing plays, they provide protection for the Quarterback so the Quarterback has time to pass the ball without being tackled.
On running plays, they provide protection for the player running the ball (e.g., the Running Back) so the ball carrier can get as far as possible without being tackled.
Offensive Linemen tend to be large so they can provide adequate protection for the Quarterback and Running Back.

#### Defense {#sec-footballRosterDefense}

The defense is on the field when the team does not have the ball (i.e., when the opposing team has the ball).

Players on defense include:

- Defensive Linemen (DL), part of the "Defensive Line"
    - Defensive End (DE)
    - Defensive Tackle (DT)
- Linebacker (LB)
    - Middle (or Inside) Linebacker (MLB)
    - Outside Linebacker (OLB)
- Defensive Back (DB), part of the "Secondary"
    - Cornerback (CB)
    - Safety (S)
        - Free Safety (FS)
        - Strong Safety (SS)

The players on the defense attempt to tackle the offensive players for as short of gains as possible and attempt to prevent completed passes.

On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to tackle the Quarterback behind the line of scrimmage before the Quarterback can throw the ball (i.e., a sack).
On rushing plays, Defensive Linemen try to tackle the ball carrier to prevent the ball carrier from advancing the ball (i.e., gaining yards).
Defensive Linemen tend to be large yet quick so they can apply pressure to the Quarterback.

Linebackers are versatile in that, on a given play, they may attempt to a) "blitz" to sack the Quarterback, b) stop the Running Back, or c) prevent a completed pass.
Linebackers tend to be strong yet agile.

Defensive Backs are specialist pass defenders.
The main role of Cornerbacks is to cover the Wide Receivers.
Safeties serve as the last line of defense for longer passes.
Defensive Backs tend to be quick and agile.

#### Special Teams {#sec-footballRosterSpecialTeams}

The special teams involves specialist players who are on the field during all kicking plays including kickoffs, field goals, and punts.

Players on special teams include:

- Kicker (K)
- Punter (P)
- Holder
- Long Snapper
- Punt Returner
- Kick Returner
- and other players intended to block for or to tackle the ball carrier

On a field goal attempt, the Long Snapper snaps the ball to the Holder, who holds the ball for the Kicker.
The Kicker attempts field goals and, during kickoffs, kicks the ball to the opposing team.
During kickoffs, the Kick Returner catches the kicked ball and returns it for as many yards as possible.
During a punt play, the Long Snapper snaps the ball to the Punter who kicks (i.e., punts) the ball to the opposing team.
The Punt Returner catches the punted ball and returns it for as many yards as possible.

### The Field {#sec-footballField}

The football field is rectangular and is 120 yards long and 53 1/3 yards wide (109.73 m x 48.77 m).^[One yard is equal to three feet. A yard is just smaller than a meter (0.9144 meters).]
At each end of the 120-yard field is a team's end zone.
Each end zone is 10 yards long (9.14 m).
Thus, the distance from one end zone to the other end zone is 100 yards (91.44 m).
Behind each end zone is a field goal post.
A diagram of a football field is depicted in Figure @fig-footballField.

::: {#fig-footballField}
![](images/footballField.png){width=50%}

A Diagram of a Football Field. The yard markers depict the distance from the nearest end zone. The orange shaded area is called the "red zone", where chances of scoring points are highest. The original figure was modified to depict field goal posts. (Figure retrieved from <https://commons.wikimedia.org/wiki/File:American_football_field.svg>)
:::

### The Gameplay {#sec-footballGameplay}

At the beginning of the game, there is a coin flip to determine which teams receives the ball first and which team takes which side of the field.
During the kickoff, the kicking team kicks the ball to the receiving team, who has the option to return the kick.
The offense starts their possession at the 25 yard line—if there is no return (i.e., a touchback)—or wherever the kick returner is tackled or goes out of bounds.

The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball (i.e., gain) 10 yards.
A team can advance the ball either by running it or by throwing (i.e., passing) and catching it.
At the end of a rushing play, the ball advances to wherever the ball carrier is tackled or goes out of bounds (i.e., wherever the player is "down").
At the end of a passing play, if the thrown ball is caught (i.e., a completed pass), the ball advances to wherever the ball carrier is tackled or goes out of bounds.
If the thrown ball is not caught in bounds before the ball hits the ground (i.e., an incomplete pass), the ball does not advance.
Wherever the ball is advanced to dictates where the next play begins.
The yard position on the field where the next play takes place from is known as the "line of scrimmage".
Neither team can cross the line the line of scrimmage until the next play begins.
To begin the play, the ball is placed on the line of scrimmage and the Center gives (or "snaps") the ball to the Quarterback.

If the team advances the ball 10 or more yards within four downs, the team receives a "first down" and is awarded a new set of downs—four more downs to advance the ball 10 more yards.
If the team advances the ball all the way to the other team's end zone, they score a touchdown.
If the team fails to advance the ball 10 or more yards within four downs, the team loses the ball, and the other team takes possession at that spot on the field.
There are risks of giving the other team the ball with a short distance to score.
Thus, on fourth down, instead of trying to advance the ball for a first down, a team may choose to kick a field goal—to get points—or to punt.

A field goal involves a kicker kicking the ball with an intent to kick the ball through the field goal posts ("uprights").
To score points by making a field goal, the kicked ball must go between the uprights (extended vertically) and over the cross bar.

Punting involves a punter kicking the ball to the other team with an intent to give their opponent worse field position, thus making it harder for the other team to score.
The punting team tries to pin the opponent as close as possible to the opponent's end zone (i.e., as far as possible from the own team's end zone), so they have a longer distance to go to score a touchdown.

There are multiple ways that ball possession can switch from the offense to the other team.
After scoring a touchdown, field goal, or safety, there is a kickoff, in which the scoring team kicks the ball to the opponent.
Another way that the ball switches possession to the other team is if the team commits a turnover.
The defense can force a turnover by an interception, fumble recovery, or turnover on downs.
A turnover due to an interception occurs when a defensive player catches the quarterback's pass.
A turnover due to a fumble recovery occurs when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent.
A turnover on downs occurs when the team attempts on fourth down to achieve the remainder of the needed 10 yards to go but fails.

Other football-related situations include tackles for loss and sacks.
A tackle for loss occurs when a ball carrier is tackled behind the line of scrimmage.
A sack occurs when a Quarterback is tackled with the ball behind the line of scrimmage.
A pass defended occurs when a defensive player knocks down the ball in the air so that the indended receiver cannot catch the ball.

### The Scoring {#sec-footballScoring}

The goal of the team with the ball (i.e., the offense) is to score points.
It can do this by either advancing the ball into the other team's end zone (6 points) or by kicking a field goal (3 points).
Advancing the ball in the other team's end zone is called a touchdown.
After a touchdown, the offense chooses to attempt either a point-after-touchdown (PAT) or a two-point conversion.
A PAT is a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point.
A two-point conversion is a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone).
If the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points.

A team can kick a field goal from any distance as long as the kick goes through the goal posts.
The current record for the longest field goal is 66 yards (by Justin Tucker in 2021).

A safety occurs when the offense is tackled with the ball in their own end zone.
When a safety occurs, the opposing team (i.e., defense) is awarded two points and the ball.

### Glossary of Terms {#sec-footballGlossary}

- running play ("run") or rushing play (or "rush")—the attempt by an offensive player, typically the Running Back or Quarterback, to advance the ball "on the ground" by running it—not by passing it forward
- passing play (or "pass")—the attempt by an offensive player, typically the Quarterback, to advance the ball by throwing it forward to an offensive player
- passing attempt—the attempt to advance the ball by passing it (i.e., a thrown pass)
- rushing attempt—the attempt to advance the ball by running it
- passing completion—a thrown pass that is succesfully caught by an offensive player
- passing incompletion—a thrown pass that is not caught by an offensive player
- passing yards—the distance (in yards) the player advanced the ball by throwing it
- rushing yards—the distance (in yards) the player advanced the ball by running it
- receving yards—the distance (in yards) the player advanced the ball by catching thrown passes and then running with it further upfield
- kick/punt return yards—the distance (in yards) the player advanced the ball by returning kicks or punts
- turnover return yards—the distance (in yards) the player advanced the ball by returning turnovers
- reception—a pass that is caught by the offensive player
- touchdown—advancing the ball into the opponent's end zone either by a) throwing a completed pass that ends up in the end zone, b) running it into the end zone, c) catching it in the end zone, or d) catching it and then running it into the end zone
- passing touchdown—advancing the ball into the opponent's end zone either by throwing a completed pass that ends up in the end zone
- rushing touchdown—advancing the ball into the opponent's end zone either by running it into the end zone
- receiving touchdown—advancing the ball into the opponent's end zone either by catching it in the end zone or by catching it and then running it into the end zone
- kick/punt return touchdown—advancing the ball into the opponent's end zone when returning a kick or punt
- turnover return touchdown—advancing the ball into the opponent's end zone when returning a turnover (i.e., interception or fumble)
- two-point conversion—a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone) that is an option given to a team that scores a touchdown;
if the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points
- block—when the defense/special teams blocks a kick or field goal by hitting the ball just after it is kicked to prevent the ball from going far
- kickoff—the kicking team kicks the ball to the receiving team, who has the option to return the kick
- field goal—a kicker kicks the ball with an intent to kick the ball through the field goal posts ("uprights").
To score points by making a field goal, the kicked ball must go between the uprights (extended vertically) and over the cross bar.
If the field goal attempt is successful, the team gains 3 points.
- point after touchdown (PAT)—a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point
- extra point returned—if the defense/special teams returns the ball into the opponent's end zone during a point after touchdown (PAT) attempt, it is worth 2 points
- punt—a punter kicks the ball to the other team with an intent to give their opponent worse field position, thus making it harder for the other team to score
- fumble lost—when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent
- fumble forced—when a defensive player knocks the ball out of the hands of an offensive player, who had possession of the ball
- fumble recovery—when a defensive player recovers a fumble by the opponent
- interception—when a defensive player catches a pass from an offensive player
- tackle—when a player brings down the ball carrier
- tackle solo—when a player is the main tackler (i.e., the primary player to bring down the ball carrier)
- tackle assist—when a player is one of two or more players who, together, bring down the ball carrier
- tackle for loss—when an offensive player is tackled with the ball behind the line of scrimmage
- sack—when a Quarterback is tackled with the ball behind the line of scrimmage
- pass defended—when a defensive player knocks down the ball in the air so that the indended receiver cannot catch the ball
- safety—when the offense is tackled with the ball in their own end zone

## Fantasy Football {#sec-fantasyFootball}

### Overview of Fantasy Football {#sec-fantasyOverview}

Fantasy football is one of the most widely played games in the history of games.
It is estimated that around 62 million people play fantasy sports^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>)], of whom around 29 million play fantasy football.^[<https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]
The participants are in charge of managing and making strategic decisions for their imaginary team to have the best possible team that will score the most points.
Thus, the participants are called "managers".
Managers make decisions such as selecting which players to draft, selecting which players to play (i.e., "start") on a weekly basis, identifying players to pick up from the remaining pool of available players (i.e., waiver wire), and making trades with other teams.
Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.

::: {.content-visible when-format="html:js"}

Here is a video that provides a brief introduction to fantasy football:

![Video that Provides an Introduction to Fantasy Football.](images/fantasyFootball.mp4){width=100%}

:::

### The Fantasy League {#sec-league}

A fantasy football "league" is composed of various imaginary (i.e., "fantasy") teams—and their associated manager.
In the fantasy league, the managers' fantasy teams play against each other.
A fantasy league is commonly composed of 8, 10, or 12 fantasy teams, but leagues can have more or fewer teams.

### The Roster of a Fantasy Team {#sec-roster}

On a given roster, a manager has a "starting lineup" and a "bench".
Each week, the manager decides which players on their roster to put in the starting lineup, and which to keep on the bench.
In many leagues, a starting lineup is composed of offensive players, a kicker, and defense/special teams:

<u>Offensive players:</u>

| Position           | Typical Number of Players in Starting Lineup |
|--------------------|----------------------------------------------|
| Quarterback (QB)   | 1                                            |
| Running Back (RB)  | 2                                            |
| Wide Receiver (WR) | 2                                            |
| Tight End (TE)     | 1                                            |
| Flex Position      | 1                                            |

: Offensive Players in the Starting Lineup {#tbl-lineupOffense tbl-colwidths="[25,75]"}

A "flex position" is a flexible position that can involve a player from various positions: e.g., a Running Back, Wide Receiver, or Tight End.

<u>Kickers:</u>

- one Kicker (K)

<u>Defense/Special Teams:</u>

- one Team Defense (DST/D/DEF) or multiple Individual Defensive Players (IDP)

### Scoring {#sec-scoring}

#### Scoring Overview {#sec-scoringOverview}

In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
Managers receive points for only those players who are on their starting lineup (not players on their bench).
A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
Scoring settings can differ from league to league.

Below are common scoring settings for fantasy leagues.

#### Offensive Players {#sec-scoringOffense}

| Statistical category                              | Points               |
|---------------------------------------------------|----------------------|
| Rushing or receiving TD                           | 6                    |
| Returning a kick or punt for a TD                 | 6                    |
| Returning or recovering a fumble for a TD         | 6                    |
| Passing TD                                        | 4                    |
| Passing INT                                       | −2                   |
| Fumble lost                                       | −2                   |
| Rushing, passing, or receiving 2-point conversion | 2                    |
| Rushing or receiving yards                        | 1 point per 10 yards |
| Passing yards                                     | 1 point per 25 yards |

: Common Scoring Settings for Offensive Players {#tbl-scoringOffense tbl-colwidths="[50,50]"}

Note: "TD" = touchdown; "INT" = interception

Other common (but not necessarily standard) statistical categories include:

- receptions (called "point per reception" [PPR] leagues)
- return yards
- passing attempts
- rushing attempts

#### Kickers {#sec-scoringKickers}

| Statistical category                              | Points |
|---------------------------------------------------|--------|
| FG made: 50+ yards                                | 5      |
| FG made: 40–49 yards                              | 4      |
| FG made: 39 yards or less                         | 3      |
| Rushing, passing, or receiving 2-point conversion | 2      |
| Point after touchdown attempt made                | 1      |
| Point after touchdown attempt missed              | −1     |
| Missed FG: 0–39 yards                             | −2     |
| Missed FG: 40–49 yards                            | −1     |

: Common Scoring Settings for Kickers {#tbl-scoringKickers tbl-colwidths="[50,50]"}

Note: "FG" = field goal

#### Team Defense/Special Teams {#sec-scoringDST}

| Statistical category          | Points |
|-------------------------------|--------|
| Defensive or special teams TD | 3      |
| Interception                  | 2      |
| Fumble recovery               | 2      |
| Blocked punt, PAT, or FG      | 2      |
| Safety                        | 2      |
| Sack                          | 1      |

: Common Scoring Settings for Team Defense/Special Teams {#tbl-scoringDST tbl-colwidths="[50,50]"}

Note: "TD" = touchdown; "PAT" = point after touchdown; "FG" = field goal

#### Individual Defensive Players {#sec-scoringIDP}

| Statistical category  | Points |
|-----------------------|--------|
| Tackle solo           | 1      |
| Tackle assist         | 0.5    |
| Tackle for loss       | 1      |
| Sack                  | 2      |
| Interception          | 4      |
| Fumble forced         | 2      |
| Fumble recovery       | 2      |
| TD                    | 6      |
| Safety                | 2      |
| Pass defended         | 1      |
| Blocked kick          | 2      |
| Extra point returned  | 2      |

: Common Scoring Settings for Individual Defensive Players {#tbl-scoringIDP tbl-colwidths="[50,50]"}

Note: "TD" = touchdown

Other common (but not necessarily standard) statistical categories include:

- turnover return yards

#### Common Scoring Abbreviations {#sec-scoringAbbreviations}

- "TD" = touchdown
- "INT" = interception
- "yds" = yards
- "ATT" = attempts
- "2-pt conversion" = two-point conversion
- "FG" = field goal
- "PAT" = point after touchdown (i.e., extra point/point after attempt)