### 2024-06-02 player-evaluation.qmd
20240601 - athletic profile and historical performance

# Player Evaluation {#sec-playerEvaluation}

Evaluating players for fantasy football could be thought of as similar to the process of evaluating companies when picking stocks to buy.
You want to evaluate and compare various assets so that you get the assets with the best value.

There are various domains of criteria we can consider when evaluating a football player's fantasy prospects.
Potential domains to consider include:

- athletic profile
- historical performance
- health
- age and career stage
- situational factors
- matchups
- cognitive and motivational factors

# Athletic Profile

- body shape
    - height
    - weight
    - hand size
- body function
    - agility
    - strength
    - speed
    - acceleration/explosiveness
    - jumping ability

In terms of body shape, we might consider a player's height, weight, and hand size.
Height allows players to see over opponents and to reach balls higher in the air.
Thus, greater height is particularly valuable for Quarterbacks and Wide Receivers.
Heavier players are tougher to budge and to tackle.
Greater weight is particularly valuable for Linemen, Fullbacks, and Tight Ends, but it can also be valuable—to a deree—for Quarterbacks, Running Backs, and Wide Receivers.
Hand size is particularly valuable for people catching the ball; thus, a larger hand size is particularly valuable for Wide Receivers and Tight Ends.

In terms of body function, we can consider a player's agility, strength, speed, acceleration/explosiveness, and jumping ability.
For Wide Receivers, speed, explosiveness, and jumping ability are particularly valuable.
For Running Backs, agility, strength, speed, and explosiveness are particularly valuable.

# Historical Performance

> "The best predictor of future behavior is past behavior." – Unknown

> "Past performance does not guarantee future results." – A common disclaimer about investments.

- performance in college
    - draft position
- performance in the NFL
- efficiency
- consistency

It is important to consider a player's past performance.
However, the extent to which historical performance may predict future performance may depend on many factors such as (a) the similarity of the prior situation to the current situation, (b) how long ago the prior situation was, and (c) the extent to which the player (or situation) has changed in the interim.
For rookies, the player does not have prior seasons of performance in the National Football League (NFL) to draw upon.
Thus, when evaluating rookies, it can be helpful to consider their performance in college or in their prior leagues.
However, there are large differences between the situation in college and the situation in the NFL, so prior success in college may not portend future success in the NFL.
An indicator that intends to be prognostic of future performance, and that accounts for past performance, is a player's draft position—that is, how early (or late) was a player selected in the NFL Draft.
The earlier a player was selected in the NFL Draft, the greater likelihood that the player will perform well.

For players who have played in the NFL, past performance becomes more relevant because, presumably, the prior situation is more similar (than was their situation in college) to their current situation.
Nevertheless, lots of things change from game to game and season to season: injuries, coaches, coaching strategies, teammates, etc.
So just because a player performed well or poorly in a given game or season does not necessarily mean that they will perform similarly in subsequent games/seasons.
Nevertheless, historical performance is one of the best indicators we have.

In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency.
How efficient were they given the number of opportunities they had.
If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
If they were relativelly inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.

# Health

- current injury status
- injury history

# Age and Career Stage

- age
- touches

# Situational Factors

Situational factors are influenced by many factors, including skill of the player, skill of their teammates, coaching style, strategy of the opposing team, game scripts, etc.

- team quality
- role on team
- teammates
- opportunity and usage
    - snap count
    - touches/targets
    - red zone usage

# Matchups

- strength of schedule
- weekly matchup

# Cognitive and Motivational Factors

- reaction time
- intelligence
- work ethic and mental toughness
- incentives
    - contract performance incentives
    - whether they are in a contract year