### 2024-06-25 factor-analysis.qmd
Built site for gh-pages

# Factor Analysis {#sec-factorAnalysis}

## Getting Started {#sec-factorAnalysisGettingStarted}

### Load Packages {#sec-factorAnalysisLoadPackages}

```{r}

```

## Overview of Factor Analysis {#sec-factorAnalysisOverview}

Factor analysis involves the estimation of latent variables.
Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.