### 2024-06-27 actuarial.qmd
20240627 - sessionInfo() for reproducibility

# Human Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}

## Getting Started {#sec-judgmentVsActuarialGettingStarted}

### Load Packages {#sec-judgmentVsActuarialLoadPackages}

```{r}

```

::: {.content-visible when-format="html"}

## Session Info {#sec-judgmentVsActuarialSessionInfo}

```{r}
sessionInfo
```

:::