### 2024-06-27 bugs.qmd

# Bugs {#sec-bugs}

Bugs to report (`nflreadr`):

- `players` database has a season column but it is only populated for some players, and it is only populated with 2020 (<https://github.com/nflverse/nflreadr/issues/238>)
- `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)

<https://github.com/nflverse/nflreadr/issues/237>