### 2024-07-09 base-rates.qmd
20240708 - base rate fallacy

# Base Rates {#sec-baseRates}

## Getting Started {#sec-baseRatesGettingStarted}

### Load Packages {#sec-baseRatesLoadPackages}

```{r}

```

## Overview {#sec-baseRatesConclusion}

The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For instance, the base rate of HIV is the probability of developing HIV.
In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).

## Conclusion {#sec-baseRatesConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-baseRatesSessionInfo}

```{r}
sessionInfo()
```

:::