### 2024-07-13 base-rates.qmd
20240713 - clarifications

# Base Rates {#sec-baseRates}

## Getting Started {#sec-baseRatesGettingStarted}

### Load Packages {#sec-baseRatesLoadPackages}

```{r}
library("petersenlab")
```

## Overview {#sec-baseRatesOverview}

Predicting player performance is a complex prediction task.
Performance is probabilistically influenced by many processes, including processes internal to the player in addition to external processes.
Moreover, people's performance occurs in the context of a dynamic system with nonlinear, probabilistic, and cascading influences that change across time.
The ever-changing system makes behavior challenging to predict.
And, similar to chaos theory, one small change in the system can lead to large differences later on.
Moreover, there are important factors to keep in mind when making predictions.

Let's consider a prediction example, assuming the following probabilities:

- The probability of contracting HIV is .3%
- The probability of a positive test for HIV is 1%
- The probability of a positive test if you have HIV is 95%

What is the probability of HIV if you have a positive test?

As we will see, the probability is: $\frac{95\% \times .3\%}{1\%} = 28.5\%$.
So based on the above probabilities, if you have a positive test, the probability that you have HIV is 28.5%.
Most people tend to vastly overestimate the likelihood that the person has HIV in this example.
Why?
Because they do not pay enough attention to the base rate (in this example, the base rate of HIV is .3%).

## Issues Around Probability {#sec-probability}

### Types of Probabilities {#sec-probabilityTypes}

It is important to distinguish between different types of probabilities: marginal probabilities, joint probabilities, and conditional probabilities.

#### Base Rate (Marginal Probability) {#sec-baseRate}

The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For instance, the base rate of HIV is the probability of developing HIV.
In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).

For instance, we can consider the following marginal probabilities:

$P(C_i)$ is the probability (i.e., base rate) of a classification, $C$, independent of other things.
A base rate is often used as the "*prior probability*" in a Bayesian model.
In our example above, $P(C_i)$ is the base rate (i.e., prevalence) of HIV in the population: $P(\text{HIV}) = .3\%$.
$P(R_i)$ is the probability (base rate) of a response, $R$, independent of other things.
In the example above, $P(R_i)$ is the base rate of a positive test for HIV: $P(\text{positive test}) = 1\%$.
The base rate of a positive test is known as the *positivity rate* or *selection ratio*.

#### Joint Probability {#sec-jointProbability}

A *joint probability* is the probability of two (or more) events occurring simultaneously.
For instance, the probability of events $A$ and $B$ both occurring together is $P(A, B)$.
A joint probability can be calculated using the [marginal probability](#sec-baseRate) of each event, as in @eq-jointProbability:

$$
P(A, B) = P(A) \cdot P(B)
$$ {#eq-jointProbability}

Conversely (and rearranging the terms for the calculation of [conditional probability](#sec-conditionalProbability)), a [joint probability](#sec-jointProbability) can also be calculated using the [conditional probability](#sec-conditionalProbability) and [marginal probability](#sec-baseRate), as in @jointProbability2:

$$
P(A, B) = P(A | B) \cdot P(B)
$$ {#eq-jointProbability2}

#### Conditional Probability {#sec-conditionalProbability}

A *conditional probability* is the probability of one event occurring given the occurrence of another event.
Conditional probabilities are written as: $P(A | B)$.
This is read as the probability that event $A$ occurs given that event $B$ occurred.
For instance, we can consider the following conditional probabilities:

$P(C | R)$ is the probability of a classification, $C$, given a response, $R$.
In other words, $P(C | R)$ is the probability of having HIV given a positive test: $P(\text{HIV} | \text{positive test})$.
$P(R | C)$ is the probability of a response, $R$, given a classification, $C$.
In the example above, $P(R | C)$ is the probability of having a positive test given that a person has HIV: $P(\text{positive test} | \text{HIV}) = 95\%$.

A conditional probability can be calculated using the [joint probability](#sec-jointProbability) and [marginal probability](#sec-baseRate) (base rate), as in @eq-conditionalProbability:

$$
P(A, B) = P(A | B) \cdot P(B)
$$ {#eq-conditionalProbability}

### Confusion of the Inverse {#sec-inverseFallacy}

A [conditional probability](#sec-conditionalProbability) is not the same thing as its reverse (or inverse) [conditional probability](#sec-conditionalProbability).
Unless the [base rate](#sec-baseRate) of the two events ($C$ and $R$) are the same, $P(C | R) \neq P(R | C)$.
However, people frequently make the mistake of thinking that two inverse [conditional probabilities](#sec-conditionalProbability) are the same.
This mistake is known as the "confusion of the inverse", or the "inverse fallacy", or the "conditional probability fallacy".
The confusion of inverse probabilities is the logical error of representative thinking that leads people to assume that the probability of $C$ given $R$ is the same as the probability of $R$ given C, even though this is not true.
As a few examples to demonstrate the logical fallacy, if 93% of breast cancers occur in high-risk women, this does not mean that 93% of high-risk women will eventually get breast cancer.
As another example, if 77% of car accidents take place within 15 miles of a driver's home, this does not mean that you will get in an accident 77% of times you drive within 15 miles of your home.

Which car is the most frequently stolen?
It is often the Honda Accord or Honda Civic—probably because they are among the most popular/commonly available cars.
The probability that the car is a Honda Accord given that a car was stolen ($p(\text{Honda Accord } | \text{ Stolen})$) is what the media reports and what the police care about.
However, that is not what buyers and car insurance companies should care about.
Instead, they care about the probability that the car will be stolen given that it is a Honda Accord ($p(\text{Stolen } | \text{ Honda Accord})$).

Applied to fantasy football, the probability that a given player will be injured given that he is a Running Back ($p(\text{Injured } | \text{ RB})$) is not the same as the probability that a given player is a Running Back given that he is injured ($p(\text{RB } | \text{ Injured})$).

### Bayes' Theorem {#sec-bayesTheorem}

An alternative way of calculating a [conditional probability](#sec-conditionalProbability) is using the inverse [conditional probability](#sec-conditionalProbability) (instead of the [joint probability](#sec-jointProbability)).
This is known as Bayes' theorem.
Bayes' theorem can help us calculate a [conditional probability](#sec-conditionalProbability) of some classification, $C$, given some response, $R$, if we know the inverse [conditional probability](#sec-conditionalProbability) and the [base rate](#sec-baseRate) (marginal probability) of each.
Bayes' theorem is in @eq-bayes1:

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}
\end{aligned}
$$ {#eq-bayes1}

Or, equivalently (rearranging the terms):

$$
\begin{aligned}
  \frac{P(C | R)}{P(R | C)} = \frac{P(C_i)}{P(R_i)}
\end{aligned}
$$ {#eq-bayes2}

Or, equivalently (rearranging the terms):

$$
\begin{aligned}
  \frac{P(C | R)}{P(C_i)} = \frac{P(R | C)}{P(R_i)}
\end{aligned}
$$ {#eq-bayes3}

More generally, Bayes' theorem has been described as:

$$
\begin{aligned}
  P(H | E) &= \frac{P(E | H) \cdot P(H)}{P(E)} \\
  \text{posterior probability} &= \frac{\text{likelihood} \times \text{prior probability}}{\text{model evidence}}
\end{aligned}
$$ {#eq-bayes6}

where $H$ is the hypothesis, and $E$ is the evidence—the new information that was not used in computing the prior probability.

In Bayesian terms, the *posterior probability* is the conditional probability of one event occurring given another event—it is the updated probability after the evidence is considered.
In this case, the posterior probability is the probability of the classification occurring ($C$) given the response ($R$).
The *likelihood* is the inverse conditional probability—the probability of the response ($R$) occurring given the classification ($C$).
The *prior probability* is the marginal probability of the event (i.e., the classification) occurring, before we take into account any new information.
The *model evidence* is the marginal probability of the other event occurring—i.e., the marginal probability of seeing the evidence.

Bayes' theorem provides the foundation for a paradigm of statistics called Bayesian statistics, which (unlike frequentist statistics) does not use *p*-values.

In the HIV example above, we can calculate the [conditional probability](#sec-conditionalProbability) of HIV given a positive test using three terms: the [conditional probability](#sec-conditionalProbability) of a positive test given HIV (i.e., the sensitivity of the test), the [base rate](#sec-baseRate) of HIV, and the [base rate](#sec-baseRate) of a positive test for HIV.
The [conditional probability](#sec-conditionalProbability) of HIV given a positive test is in @eq-hivExample1:

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
  P(\text{HIV} | \text{positive test}) &= \frac{P(\text{positive test} | \text{HIV}) \cdot P(\text{HIV})}{P(\text{positive test})} \\
  &= \frac{\text{sensitivity of test} \times \text{base rate of HIV}}{\text{base rate of positive test}} \\
  &= \frac{95\% \times .3\%}{1\%} = \frac{.95 \times .003}{.01}\\
  &= 28.5\%
\end{aligned}
$$ {#eq-hivExample1}

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.

```{r}
petersenlab::pAgivenB(
  pBgivenA = .95,
  pA = .003,
  pB = .01)
```

Thus, assuming the probabilities in the example above, the [conditional probability](#conditionalProbability) of having HIV if a person has a positive test is 28.5%.
Given a positive test, chances are higher than not that the person does not have HIV.

Now let's see what happens if the person tests positive a second time.
We would revise our "[prior probability](#sec-baseRate)" for HIV from the general prevalence in the population (0.3%) to be the "posterior probability" of HIV given a first positive test (28.5%).
This is known as *Bayesian updating*.
We would also update the "evidence" to be the [marginal probability](#sec-baseRate) of getting a second positive test.

If we do not know a [marginal probability](#sec-baseRate) (i.e., base rate) of an event (e.g., getting a second positive test), we can calculate a [marginal probability](#sec-baseRate) with the *law of total probability* using [conditional probabilities](#sec-conditionalProbability) and the [marginal probability](#sec-baseRate) of another event (e.g., having HIV).
According to the law of total probability, the probability of getting a positive test is the probability that a person with HIV gets a positive test (i.e., sensitivity) times the base rate of HIV plus the probability that a person without HIV gets a positive test (i.e., false positive rate) times the [base rate](#sec-baseRate) of not having HIV, as in @eq-lawOfTotalProbability:

$$
\begin{aligned}
 P(\text{not } C_i) &= 1 - P(C_i) \\
  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
  1\% &= 95\% \times .3\% + P(R | \text{not } C) \times 99.7\% \\
\end{aligned}
$$ {#eq-lawOfTotalProbability}

In this case, we know the [marginal probability](#sec-baseRate) ($P(R_i)$), and we can use that to solve for the unknown [conditional probability](#sec-conditionalProbability) that reflects the false positive rate ($P(R | \text{not } C)$), as in @eq-conditionalProbabilityRevised:

$$
\scriptsize
\begin{aligned}
  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) && \\
  P(R_i) - [P(R | \text{not } C) \cdot P(\text{not } C_i)] &= P(R | C) \cdot P(C_i) && \text{Move } P(R | \text{not } C) \text{ to the left side} \\
  - [P(R | \text{not } C) \cdot P(\text{not } C_i)] &= P(R | C) \cdot P(C_i) - P(R_i) && \text{Move } P(R_i) \text{ to the right side} \\
  P(R | \text{not } C) \cdot P(\text{not } C_i) &= P(R_i) - [P(R | C) \cdot P(C_i)] && \text{Multiply by } -1 \\
  P(R | \text{not } C) &= \frac{P(R_i) - [P(R | C) \cdot P(C_i)]}{P(\text{not } C_i)} && \text{Divide by } P(R | \text{not } C) \\
  &= \frac{1\% - [95\% \times .3\%]}{99.7\%} = \frac{.01 - [.95 \times .003]}{.997}\\
  &= .7171515\% \\
\end{aligned}
$$ {#eq-conditionalProbabilityRevised}

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.

```{r}
petersenlab::pBgivenNotA(
  pBgivenA = .95,
  pA = .003,
  pB = .01)
```

With this [conditional probability](#sec-conditionalProbability) ($P(R | \text{not } C)$), the updated [marginal probability](#sec-baseRate) of having HIV ($P(C_i)$), and the updated marginal probability of not having HIV ($P(\text{not } C_i)$), we can now calculate an updated estimate of the [marginal probability](#sec-baseRate) of getting a second positive test.
The probability of getting a second positive test is the probability that a person with HIV gets a second positive test (i.e., sensitivity) times the updated probability of HIV plus the probability that a person without HIV gets a second positive test (i.e., false positive rate) times the updated probability of not having HIV, as in @eq-baseRateUpdated:

$$
\begin{aligned}
  P(R_{i}) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
  &= 95\% \times 28.5\% + .7171515\% \times 71.5\% = .95 \times .285 + .007171515 \times .715 \\
  &= 27.58776\%
\end{aligned}
$$ {#eq-baseRateUpdated}

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pB()` function that estimates the marginal probability of one event, $B$.

```{r}
petersenlab::pB(pBgivenA = .95, pA = .285, pBgivenNotA = .007171515)
```

We then substitute the updated [marginal probability](#sec-baseRate) of HIV ($P(C_i)$) and the updated [marginal probability](#sec-baseRate) of getting a second positive test ($P(R_i)$) into Bayes' theorem to get the probability that the person has HIV if they have a second positive test (assuming the errors of each test are independent, i.e., uncorrelated), as in @eq-baseRateUpdated2:

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
  P(\text{HIV} | \text{a second positive test}) &= \frac{P(\text{a second positive test} | \text{HIV}) \cdot P(\text{HIV})}{P(\text{a second positive test})} \\
  &= \frac{\text{sensitivity of test} \times \text{updated base rate of HIV}}{\text{updated base rate of positive test}} \\
  &= \frac{95\% \times 28.5\%}{27.58776\%} \\
  &= 98.14\%
\end{aligned}
$$ {#eq-baseRateUpdated2}

The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.

```{r}
petersenlab::pAgivenB(
  pBgivenA = .95,
  pA = .285,
  pB = .2758776)
```

Thus, a second positive test greatly increases the posterior probability that the person has HIV from 28.5% to over 98%.

As seen in the rearranged formula in @eq-bayes2, the ratio of the [conditional probabilities](#sec-conditionalProbability) is equal to the ratio of the [base rates](#sec-baseRate).
Thus, it is important to consider [base rates](#sec-baseRate).
People have a strong tendency to ignore (or give insufficient weight to) [base rates](#sec-baseRate) when making predictions.
The failure to consider the [base rate](#sec-baseRate) when making predictions when given specific information about a case is known as the [base rate fallacy](#sec-fallaciesBaseRate) or as [base rate neglect](#sec-fallaciesBaseRate).
For example, people tend to say that the probability of a rare event is more likely than it actually is given specific information.

As seen in the rearranged formula in @eq-bayes3, the inverse [conditional probabilities](#sec-conditionalProbability) ($P(C | R)$ and $P(R | C)$) are not equal unless the [base rates](#sec-baseRate) of $C$ and $R$ are the same.
If the [base rates](#sec-baseRate) are not equal, we are making at least some prediction errors.
If $P(C_i) > P(R_i)$, our predictions must include some false negatives.
If $P(R_i) > P(C_i)$, our predictions must include some false positives.

In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
Bayes' theorem provides a powerful tool to anchor predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).

## Base Rate of Rookie Performance {#sec-baseRateRookiePerformance}

### Quarterbacks {#sec-baseRateRookiePerformanceQBs}

### Running Backs {#sec-baseRateRookiePerformanceRBs}

## How to Account for Base Rates {#sec-accountForBaseRates}

There are various ways to account for [base rates](#sec-baseRate).
One approach is to use [actuarial formulas](#sec-actuarialPrediction) (rather than human judgment) to make the predictions.
[Actuarial formulas](#sec-actuarialPrediction) based on [multiple regression](#sec-multipleRegression) or [machine learning](#sec-machineLearning) can account for the [base rate](#sec-baseRate) of the event.

Another approach is to leverage Bayes' theorem, using Bayesian updating and the probability nomogram.

### Bayesian Updating {#sec-bayesianUpdating}

Bayesian updating involves comparing the relative probability of two outcomes, $P(C | R)$ versus $P(\text{not } C | R)$.
If we want to compare the relative probability of two outcomes, we can use the odds form of Bayes' theorem, as in @eq-bayes5:
$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
  P(\text{not } C | R) &= \frac{P(R | \text{not } C) \cdot P(\text{not } C_i)}{P(R_i)} \\
  \frac{P(C | R)}{P(\text{not } C | R)} &= \frac{\frac{P(R | C) \cdot P(C_i)}{P(R_i)}}{\frac{P(R | \text{not } C) \cdot P(\text{not } C_i)}{P(R_i)}} \\
  &= \frac{P(R | C) \cdot P(C_i)}{P(R | \text{not } C) \cdot P(\text{not } C_i)} \\
  &= \frac{P(C_i)}{P(\text{not } C_i)} \times \frac{P(R | C)}{P(R | \text{not } C)} \\
  \text{posterior odds} &= \text{prior odds} \times \text{likelihood ratio}
\end{aligned}
$$ {#eq-bayes5}

As presented in @eq-bayes5, the posttest (or posterior) odds are equal to the pretest odds multiplied by the [likelihood ratio](#sec-diagnosticLikelihoodRatio).
Below, we describe the [likelihood ratio](#sec-diagnosticLikelihoodRatio).

#### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio}

A likelihood ratio is the ratio of two probabilities.
It can be used to compare the likelihood of two possibilities.
The diagnostic likelihood ratio is an index of the predictive validity of an instrument: it is the ratio of the probability that a test result is correct to the probability that the test result is incorrect.
The diagnostic likelihood ratio is also called the risk ratio.
There are two types of diagnostic likelihood ratios: the [positive likelihood ratio](#sec-positiveLikelihoodRatio) and the [negative likelihood ratio](#sec-negativeLikelihoodRatio).

##### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio}

The positive likelihood ratio (LR+) compares the true positive rate to the false positive rate.
Positive likelihood ratio values range from 1 to infinity.
Higher values reflect greater accuracy, because it indicates the degree to which a true positive is more likely than a false positive.
The formula for calculating the positive likelihood ratio is in @eq-positiveLikelihoodRatio.

$$
\begin{aligned}
  \text{positive likelihood ratio (LR+)} &= \frac{\text{TPR}}{\text{FPR}} \\
  &= \frac{P(R|C)}{P(R|\text{not } C)} \\
  &= \frac{P(R|C)}{1 - P(\text{not } R|\text{not } C)} \\
  &= \frac{\text{sensitivity}}{1 - \text{specificity}}
\end{aligned}
$$ {#eq-positiveLikelihoodRatio}

##### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio}

The negative likelihood ratio (LR−) compares the false negative rate to the true negative rate.
Negative likelihood ratio values range from 0 to 1.
Smaller values reflect greater accuracy, because it indicates that a false negative is less likely than a true negative.
The formula for calculating the negative likelihood ratio is in @negativeLikelihoodRatio.

$$
\begin{aligned}
  \text{negative likelihood ratio } (\text{LR}-) &= \frac{\text{FNR}}{\text{TNR}} \\
  &= \frac{P(\text{not } R|C)}{P(\text{not } R|\text{not } C)} \\
  &= \frac{1 - P(R|C)}{P(\text{not } R|\text{not } C)} \\
  &= \frac{1 - \text{sensitivity}}{\text{specificity}}
\end{aligned}
$$ {#eq-positiveLikelihoodRatio}

#### Probability Nomogram {#sec-probabilityNomogram}

Using [Bayes's theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
The calculation of posttest probability is described in INSERT.
A *probability nomogram* is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
The updated (posttest) probability is where the connecting line crosses the third, right line.

::: {#fig-probabilityNomogram}
![](images/probabilityNomogram.png){width=50%}

Probability Nomogram. (Figure retrieved from [https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png](https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png)).
:::

For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the sensitivity (SN) and specificity (SP) of the test.

```{r}
petersenlab::posttestProbability(
  pretestProb = .005,
  likelihoodRatio = 10)

petersenlab::posttestProbability(
  pretestProb = .005,
  SN = .90,
  SP = .91)
```

The function can also estimate the posttest probability of an event given the number of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):

```{r}
petersenlab::posttestProbability(
  TP = 450,
  TN = 90545,
  FP = 8955,
  FN = 50)
```

We discuss true positives (TP), true negatives (TN), false positives (FP), false negatives (FN), sensitivity (SN), and specificity (SP) in INSERT.

If the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 0.11 from a negative test (i.e., [negative likelihood ratio](#sec-negativeLikelihoodRatio)), the updated probability is nearly indistinguishable from zero (0.05%).

```{r}
petersenlab::posttestProbability(
  pretestProb = .005,
  likelihoodRatio = 0.11)
```

::: {#fig-probabilityNomogramLine}
![](images/probabilityNomogramLine.png){width=50%}

Probability Nomogram Example. (Figure adapted from [https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png](https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png). Also provided in: @Petersen2024a and @PetersenPrinciplesPsychAssessment.)
:::

A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the sensitivity (SN) and specificity (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):

```{r}
petersenlab::nomogrammer(
  pretestProb = .005,
  SN = 0.90,
  SP = 0.91)

petersenlab::nomogrammer(
  pretestProb = .005,
  PLR = 10,
  NLR = 0.11)
```

The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.

The function can also create a nomogram plot using the true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):

```{r}
petersenlab::nomogrammer(
  TP = 450,
  TN = 90545,
  FP = 8955,
  FN = 50)
```

## Conclusion {#sec-baseRatesConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-baseRatesSessionInfo}

```{r}
sessionInfo()
```

:::