### 2024-07-28 mixed-models.qmd
20240728 - fit indices

# Mixed Models {#sec-mixedModels}

## Getting Started {#sec-mixedModelsGettingStarted}

### Load Packages {#sec-mixedModelsLoadPackages}

```{r}
library("petersenlab")
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("parallel")
library("plotly")
library("viridis")
library("tidyverse")
```

### Specify Package Options {#sec-mixedModelsPackageOptions}

```{r}
emm_options(lmerTest.limit = 100000)
emm_options(pbkrtest.limit = 100000)
```

### Load Data {#sec-mixedModelsLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_depthCharts.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

## Overview of Mixed Models {#sec-mixedModelsOverview}

We will discuss a modeling framework that goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.

### Ecological Fallacy {#sec-ecologicalFallacy}

### Simpson's Paradox {#sec-simpsonsParadox}

## Fantasy Points Per Season by Position, Age, and Experience {#sec-fantasyPointsByAgeExperience}

```{r}
player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
  filter(position_group %in% c("QB","RB","WR","TE"))

player_stats_seasonal_offense_subset$position[which(player_stats_seasonal_offense_subset$position == "HB")] <- "RB"

player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
  filter(position == "K")

player_stats_seasonal_offense_subset <- bind_rows(
  player_stats_seasonal_offense_subset,
  player_stats_seasonal_kicking_subset
)

player_stats_seasonal_offense_subset$player_idFactor <- factor(player_stats_seasonal_offense_subset$player_id)
player_stats_seasonal_offense_subset$positionFactor <- factor(player_stats_seasonal_offense_subset$position)
```

```{r}
#| eval: false
#| include: false

for(i in unique(nfl_depthCharts$season)){
  print(i)
  nfl_depthCharts %>%
    filter(season == i) %>% 
    select(week) %>% 
    arrange(week) %>% 
    unique() %>% 
    pull() %>% 
    print()
}
```

```{r}
seasons17week <- 2001:2020
seasons18week <- 2021:max(nfl_depthCharts$season, na.rm = TRUE)

endOfSeasonDepthCharts <- nfl_depthCharts %>% 
  filter((season %in% seasons17week & week == 18) | (season %in% seasons18week & week == 19)) # get end-of-season depth charts

qb1s <- endOfSeasonDepthCharts %>% 
  filter(position == "QB", depth_team == 1)

fb1s <- endOfSeasonDepthCharts %>% 
  filter(position == "FB", depth_team == 1)

k1s <- endOfSeasonDepthCharts %>% 
  filter(position == "K", depth_team == 1)

rb1s <- endOfSeasonDepthCharts %>% 
  filter(position == "RB", depth_team == 1)

wr1s <- endOfSeasonDepthCharts %>% 
  filter(position == "WR", depth_team == 1)

te1s <- endOfSeasonDepthCharts %>% 
  filter(position == "TE", depth_team == 1)

player_stats_seasonal_offense_subsetDepth <- player_stats_seasonal_offense_subset %>% 
  filter(player_id %in% c(
    qb1s$gsis_id,
    fb1s$gsis_id,
    k1s$gsis_id,
    rb1s$gsis_id,
    wr1s$gsis_id,
    te1s$gsis_id
    ))
```

### Scatterplots of Fantasy Points by Age and Position {#sec-fantasyPointsByAgeScatterplot}

[Scatterplots](#sec-scatterplot) are a helpful tool for quickly examining the association between two variables.
However, [scatterplots](#sec-scatterplot)—as well as correlation and multiple regression—can hide meaningful associations that differ across units of analysis.

#### Quarterbacks {#sec-fantasyPointsByAgeScatterplotQB}

A [scatterplot](#sec-scatterplot) of Quarterbacks' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotQB.

```{r}
#| label: fig-fantasyPointsByAgeScatterplotQB
#| fig-cap: "Scatterplot of Fantasy Points by Age for Quarterbacks."
#| code-fold: true

plot_scatterplotFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeQB,
  tooltip = c("age","fantasy_points","text","label"))
```

Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to increase with age.

```{r}
cor.test(
  formula = ~ age + fantasy_points,
  data = player_stats_seasonal_offense_subset %>% filter(position == "QB")
)
```

#### Fullbacks {#sec-fantasyPointsByAgeScatterplotFB}

A [scatterplot](#sec-scatterplot) of Fullbacks' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotFB.

```{r}
#| label: fig-fantasyPointsByAgeScatterplotFB
#| fig-cap: "Scatterplot of Fantasy Points by Age for Fullbacks."
#| code-fold: true

plot_scatterplotFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeFB,
  tooltip = c("age","fantasy_points","text","label"))
```

Based on the [scatterplot](#sec-scatterplot), Fullbacks' fantasy points appear to be relatively stable across ages.

#### Running Backs {#sec-fantasyPointsByAgeScatterplotRB}

A [scatterplot](#sec-scatterplot) of Running Backs' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotRB.

```{r}
#| label: fig-fantasyPointsByAgeScatterplotRB
#| fig-cap: "Scatterplot of Fantasy Points by Age for Running Backs."
#| code-fold: true

plot_scatterplotFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeRB,
  tooltip = c("age","fantasy_points","text","label"))
```

Based on the [scatterplot](#sec-scatterplot), Running Backs' fantasy points appear to be relatively stable across ages.

#### Wide Receivers {#sec-fantasyPointsByAgeScatterplotWR}

A [scatterplot](#sec-scatterplot) of Wide Receivers' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotWR.

```{r}
#| label: fig-fantasyPointsByAgeScatterplotWR
#| fig-cap: "Scatterplot of Fantasy Points by Age for Wide Receivers."
#| code-fold: true

plot_scatterplotFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeWR,
  tooltip = c("age","fantasy_points","text","label"))
```

Based on the [scatterplot](#sec-scatterplot), Wide Receivers' fantasy points appear to be relatively stable across ages.

#### Tight Ends {#sec-fantasyPointsByAgeScatterplotTE}

A [scatterplot](#sec-scatterplot) of Tight Ends' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotTE.

```{r}
#| label: fig-fantasyPointsByAgeScatterplotTE
#| fig-cap: "Scatterplot of Fantasy Points by Age for Tight Ends."
#| code-fold: true

plot_scatterplotFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  geom_smooth(
    mapping = aes(
    x = age,
    y = fantasy_points),
    inherit.aes = FALSE
  ) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_scatterplotFantasyPointsByAgeTE,
  tooltip = c("age","fantasy_points","text","label"))
```

Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Tight Ends' fantasy points appear to increase with age.

```{r}
cor.test(
  formula = ~ age + fantasy_points,
  data = player_stats_seasonal_offense_subset %>% filter(position == "TE")
)
```

### Plots of Raw Trajectories of Fantasy Points By Age and Player {#sec-fantasyPointsByAgeExperienceRawData}

[Scatterplots](#sec-scatterplot) can be helpful for quickly visualizing the association between two variables.
However, as mentioned earlier, [scatterplots](#sec-scatterplot) can hide the association between variables at different units of analysis.
For instance, consider that we are trying to predict how a player will perform based on their age.
We are interested not only in what the association is between age and fantasy points between players (i.e., a between-person association).
We are also interested in what the association is between age and fantasy points *within* a given player (and within each player; i.e., a within-individual association).
Arguably, the within-individual association between age and fantasy points is more relevant to the prediction of performance than the association between age and fantasy points between players.
Assuming that the between-player association between age and fantasy points is the same as the within-player association when it is not is an example of the [ecological fallacy](#sec-fallaciesEcological).

Below, we depict players' raw trajectories of fantasy points as a function of age.
These are known as spaghetti plots.
By examining the trajectory for each player, we can get a better understanding of hor performance changes (within an individual) as a function of age.

#### Quarterbacks {#sec-fantasyPointsByAgeExperienceRawDataQB}

A plot of Quarterbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeQB.

```{r}
#| label: fig-rawFantasyPointsByAgeQB
#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Quarterbacks."
#| code-fold: true

plot_rawFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "QB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeQB,
  tooltip = c("age","fantasy_points","text","label"))
```

#### Fullbacks {#sec-fantasyPointsByAgeExperienceRawDataFB}

A plot of Fullbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeFB.

```{r}
#| label: fig-rawFantasyPointsByAgeFB
#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Fullbacks."
#| code-fold: true

plot_rawFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "FB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeFB,
  tooltip = c("age","fantasy_points","text","label"))
```

#### Running Backs {#sec-fantasyPointsByAgeExperienceRawDataRB}

A plot of Running Backs' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeRB.

```{r}
#| label: fig-rawFantasyPointsByAgeRB
#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Running Backs."
#| code-fold: true

plot_rawFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "RB") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeRB,
  tooltip = c("age","fantasy_points","text","label"))
```

#### Wide Receivers {#sec-fantasyPointsByAgeExperienceRawDataWR}

A plot of Wide Receivers' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeWR.

```{r}
#| label: fig-rawFantasyPointsByAgeWR
#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Wide Receivers."
#| code-fold: true

plot_rawFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "WR") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeWR,
  tooltip = c("age","fantasy_points","text","label"))
```

#### Tight Ends {#sec-fantasyPointsByAgeExperienceRawDataTE}

A plot of Tight Ends' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeTE.

```{r}
#| label: fig-rawFantasyPointsByAgeTE
#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Tight Ends."
#| code-fold: true

plot_rawFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subset %>% 
    filter(position == "TE") %>% 
    mutate(
      age = round(age, 2),
      fantasy_points = round(fantasy_points, 2)
    ),
  mapping = aes(
    x = age,
    y = fantasy_points,
    color = player_id)) +
  geom_line(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
  )) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(
  plot_rawFantasyPointsByAgeTE,
  tooltip = c("age","fantasy_points","text","label"))
```

### Linear Regression Models {#sec-fantasyPointsByAgeExperienceModelsRegression}

#### Null Model {#sec-fantasyPointsByAgeExperienceModelsRegressionNull}

```{r}
pointsPerSeason_nullModel <- lm(
  fantasy_points ~ 1,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

summary(pointsPerSeason_nullModel)
summary(pointsPerSeason_nullModel)$r.squared
AIC(pointsPerSeason_nullModel)
MuMIn::AICc(pointsPerSeason_nullModel)
```

#### Linear Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}

```{r}
pointsPerSeason_linearRegression <- lm(
  fantasy_points ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

summary(pointsPerSeason_linearRegression)
summary(pointsPerSeason_linearRegression)$r.squared
AIC(pointsPerSeason_linearRegression)
MuMIn::AICc(pointsPerSeason_linearRegression)
```

#### Quadratic Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}

```{r}
pointsPerSeason_quadraticRegression <- lm(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
  data = player_stats_seasonal_offense_subset,
  na.action = "na.exclude"
)

summary(pointsPerSeason_quadraticRegression)
summary(pointsPerSeason_quadraticRegression)$r.squared
AIC(pointsPerSeason_quadraticRegression)
MuMIn::AICc(pointsPerSeason_quadraticRegression)
```

#### Compare Models {#sec-fantasyPointsByAgeExperienceModelsRegressionCompare}

```{r}
anova(
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression
)

AIC(
  pointsPerSeason_nullModel,
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression
  )

lmModels <- list(
  "nullModel" = pointsPerSeason_nullModel,
  "linearRegression" = pointsPerSeason_linearRegression,
  "quadraticRegression" = pointsPerSeason_quadraticRegression
)

bbmle::AICtab(lmModels)

summary(pointsPerSeason_nullModel)$r.squared
summary(pointsPerSeason_linearRegression)$r.squared
summary(pointsPerSeason_quadraticRegression)$r.squared

deviance(pointsPerSeason_nullModel)
deviance(pointsPerSeason_linearRegression)
deviance(pointsPerSeason_quadraticRegression)

logLik(pointsPerSeason_nullModel)
logLik(pointsPerSeason_linearRegression)
logLik(pointsPerSeason_quadraticRegression)
```

### Mixed Models {#sec-fantasyPointsByAgeExperienceModels}

By accounting for which player each observation comes from using mixed models, we can examine the association between age and fantasy points in a more meaningful way, without violating the [assumption in multiple regression](#sec-assumptionsRegression) that the observations are independent (i.e., that the residuals are uncorrelated).

#### Random Intercepts Model {#sec-fantasyPointsByAgeExperienceModelsIntercepts}

```{r}
pointsPerSeason_randomIntercepts <- lmerTest::lmer(
  fantasy_points ~ 1 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_randomIntercepts)
MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
performance::icc(pointsPerSeason_randomIntercepts)
AIC(pointsPerSeason_randomIntercepts)
AICc(pointsPerSeason_randomIntercepts)
```

#### Random Intercepts Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}

```{r}
pointsPerSeason_position <- lmerTest::lmer(
  fantasy_points ~ positionFactor + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_position)
MuMIn::r.squaredGLMM(pointsPerSeason_position)
emmeans::emmeans(pointsPerSeason_position, "positionFactor")
performance::icc(pointsPerSeason_position)
AIC(pointsPerSeason_position)
```

#### Identify the Best-Fitting Functional Form of Age {#sec-fantasyPointsByAgeExperienceModelsAge}

##### Linear Models {#sec-fantasyPointsByAgeExperienceModelsAgeLinear}

```{r}
pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
AIC(pointsPerSeason_positionAgeFixedLinearSlopes)

pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeRandomLinearSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
```

##### Quadratic Models {#sec-fantasyPointsByAgeExperienceModelsAgeQuadratic}

```{r}
pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)

pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)

pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
```

##### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeCompare}

After fitting several models, we now must compare their fit to determine which model fits "best" while also considering parsimony.
Parsimonious models are more likely to be true and more likely to generalize to other samples, because more complex models are more likely to overfit the data.
Thus, more complex models will almost always fit better than simpler models.
Thus, we are not just interested in whether a more complex model fits better than the simpler model; we also care about whether the more complex model fits *significantly* better than the simpler model given its additional complexity.
For evaluating and comparing models, we examine the [likelihood ratio test](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT), the [Akaike Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC) (AIC), the [corrected AIC](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc) (AICc), the [Bayesian Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC) (BIC), [$R^2$](#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared), [deviance](#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance), and [log likelihood](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLogLik).

The BIC penalizes model complexity more than the AIC does.
The BIC is preferable when there is a "true" model, and one intends to identify the true model.
The AIC is preferable when we are concerned more about predictive accuracy and when overfitting is less of a concern.
Because we are more concerned about predictive accuracy and we do not believe one of these models is the "true" model per se of age-related changes in fantasy performance, we will give more weight to AIC than BIC.

Below, we specify various groups of models for the model fit comparisons:

```{r}
lmVsMixedModel <- list(
  "nullModel" = pointsPerSeason_nullModel,
  "randomIntercepts" = pointsPerSeason_randomIntercepts
)

lmAndMixedModels <- list(
  "nullModel" = pointsPerSeason_nullModel,
  "linearRegression" = pointsPerSeason_linearRegression,
  "quadraticRegression" = pointsPerSeason_quadraticRegression,
  "randomIntercepts" = pointsPerSeason_randomIntercepts,
  "position" = pointsPerSeason_position,
  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)

mixedModels <- list(
  "randomIntercepts" = pointsPerSeason_randomIntercepts,
  "position" = pointsPerSeason_position,
  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)

mixedModels1 <- list(
  "randomIntercepts" = pointsPerSeason_randomIntercepts,
  "position" = pointsPerSeason_position
)

mixedModels2 <- list(
  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)
```

###### Likelihood Ratio Test {#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT}

```{r}
anova(
  pointsPerSeason_randomIntercepts,
  pointsPerSeason_position
)

anova(
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)
```

###### Akaike Information Criterion (AIC) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC}

```{r}
AIC(
  pointsPerSeason_nullModel,
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression,
  pointsPerSeason_randomIntercepts,
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
  )

bbmle::AICtab(lmAndMixedModels)
```

###### Corrected Akaike Information Criterion (AICc) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc}

```{r}
#AICcmodavg::aictab(lmVsMixedModel) # throws error (can't mix lm with lmer)
bbmle::AICctab(lmVsMixedModel)

AICcmodavg::aictab(mixedModels) # throws error (can't mix lm with lmer)
#bbmle::AICctab(mixedModels) # throws error (different numbers of observations)

AICcmodavg::aictab(mixedModels1)
bbmle::AICctab(mixedModels1)

AICcmodavg::aictab(mixedModels2)
bbmle::AICctab(mixedModels2)
```

###### Bayesian Information Criterion (BIC) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC}

```{r}
BIC(
  pointsPerSeason_nullModel,
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression,
  pointsPerSeason_randomIntercepts,
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
  )

#AICcmodavg::bictab(lmAndMixedModels) # throws error (can't mix lm with lmer)
bbmle::BICtab(lmVsMixedModel)

AICcmodavg::bictab(mixedModels)
#bbmle::AICctab(mixedModels) # throws error (different numbers of observations)

AICcmodavg::bictab(mixedModels1)
bbmle::BICtab(mixedModels1)

AICcmodavg::bictab(mixedModels2)
bbmle::BICtab(mixedModels2)
```

###### $R^2$ {#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared}

```{r}
summary(pointsPerSeason_nullModel)$r.squared
summary(pointsPerSeason_linearRegression)$r.squared
summary(pointsPerSeason_quadraticRegression)$r.squared
MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

###### Deviance {#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance}

```{r}
deviance(pointsPerSeason_nullModel)
deviance(pointsPerSeason_linearRegression)
deviance(pointsPerSeason_quadraticRegression)
deviance(pointsPerSeason_randomIntercepts)
deviance(pointsPerSeason_positionAgeFixedLinearSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

###### Log Likelihood {#sec-fantasyPointsByAgeExperienceModelsAgeCompareLogLik}

```{r}
logLik(pointsPerSeason_nullModel)
logLik(pointsPerSeason_linearRegression)
logLik(pointsPerSeason_quadraticRegression)
logLik(pointsPerSeason_randomIntercepts)
logLik(pointsPerSeason_positionAgeFixedLinearSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

##### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}

```{r}
num_cores <- detectCores()
num_cores

pointsPerSeason_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subset,
  nthreads = num_cores
)

pointsPerSeason_gamSummary <- summary(pointsPerSeason_gam)

pointsPerSeason_gamSummary
pointsPerSeason_gamSummary$r.sq
MuMIn::r.squaredGLMM(pointsPerSeason_gam)
AIC(pointsPerSeason_gam)
```

###### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}

```{r}
linearMixedModelsVsGAM <- list(
  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
  "gam" = pointsPerSeason_gam
)

AIC(
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
  pointsPerSeason_gam
)

#AICcmodavg::aictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
bbmle::AICctab(linearMixedModelsVsGAM)

BIC(
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
  pointsPerSeason_gam
)

#AICcmodavg::bictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
bbmle::BICtab(linearMixedModelsVsGAM)

summary(pointsPerSeason_nullModel)$r.squared
summary(pointsPerSeason_linearRegression)$r.squared
summary(pointsPerSeason_quadraticRegression)$r.squared
MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
MuMIn::r.squaredGLMM(pointsPerSeason_gam)

deviance(pointsPerSeason_nullModel)
deviance(pointsPerSeason_linearRegression)
deviance(pointsPerSeason_quadraticRegression)
deviance(pointsPerSeason_randomIntercepts)
deviance(pointsPerSeason_positionAgeFixedLinearSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
deviance(pointsPerSeason_gam)

logLik(pointsPerSeason_nullModel)
logLik(pointsPerSeason_linearRegression)
logLik(pointsPerSeason_quadraticRegression)
logLik(pointsPerSeason_randomIntercepts)
logLik(pointsPerSeason_positionAgeFixedLinearSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
logLik(pointsPerSeason_gam)
```

##### Players Who Were (at Least Once) at the Top of the End-of-Season Depth Chart {#sec-fantasyPointsByAgeExperienceModelsAgeDepth}

```{r}
pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
  data = player_stats_seasonal_offense_subset,
  control = lmerControl(optimizer = "bobyqa")
)

summary(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

pointsPerSeasonDepth_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
  data = player_stats_seasonal_offense_subsetDepth,
  nthreads = num_cores
)

pointsPerSeasonDepth_gamSummary <- summary(pointsPerSeasonDepth_gam)

pointsPerSeasonDepth_gamSummary
pointsPerSeasonDepth_gamSummary$r.sq
MuMIn::r.squaredGLMM(pointsPerSeasonDepth_gam)
AIC(pointsPerSeasonDepth_gam)
```

### Plots of Model-Implied Fantasy Points by Position and Age {#sec-fantasyPointsByAgePlots}

```{r}
# Create newdata object
pointsPerSeason_positionAge_newData <- expand.grid(
  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
  age = seq(from = 20, to = 40, length.out = 10000)
)

pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0

# From Quadratic Model: All Players
pointsPerSeason_positionAge_newData$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

# From Quadratic Model: Players at Top of End-of-Season Depth Chart
pointsPerSeason_positionAge_newData$fantasyPoints_depthQuadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = pointsPerSeason_positionAge_newData,
  re.form = NA
)

# From GAM Model: All Players
pointsPerSeason_positionAge_newData$fantasyPoints_gam <- predict(
  object = pointsPerSeason_gam,
  newdata = pointsPerSeason_positionAge_newData,
  newdata.guaranteed = TRUE,
  exclude = "s(player_idFactor,ageCentered20)"
)

# From GAM Model: Players at Top of End-of-Season Depth Chart
pointsPerSeason_positionAge_newData$fantasyPoints_depthGAM <- predict(
  object = pointsPerSeasonDepth_gam,
  newdata = pointsPerSeason_positionAge_newData,
  newdata.guaranteed = TRUE,
  exclude = "s(player_idFactor,ageCentered20)"
)
```

Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAMdepth].

#### Quadratic Model {#sec-fantasyPointsByAgePlotsQuadratic}

```{r}
#| label: fig-fantasyPointsByAgeQuadratic
#| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age."
#| code-fold: true

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_quadratic,
    color = positionFactor
  )
) + 
  geom_smooth() +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Quadratic Model with All Players"
  ) +
  theme_classic()
```

#### Quadratic Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsQuadraticDepth}

```{r}
#| label: fig-fantasyPointsByAgeQuadraticDepth
#| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age For Players Who Were Once at the Top of the End-of-Season Depth Chart."
#| code-fold: true

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_depthQuadratic,
    color = positionFactor
  )
) + 
  geom_smooth() +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Quadratic Model with Players Who Were Once at Top of End-of-Season Depth Chart"
  ) +
  theme_classic()
```

#### Generalized Additive Model {#sec-fantasyPointsByAgePlotsGAM}

```{r}
#| label: fig-fantasyPointsByAgeGAM
#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Generalized Additive Model."
#| code-fold: true

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    color = positionFactor
  )
) + 
  geom_smooth() +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Generalized Additive Model with All Players"
  ) +
  theme_classic()
```

#### Generalized Additive Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsGAMdepth}

```{r}
#| label: fig-fantasyPointsByAgeGAMdepth
#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."
#| code-fold: true

ggplot2::ggplot(
  data = pointsPerSeason_positionAge_newData,
  mapping = aes(
    x = age,
    y = fantasyPoints_depthGAM,
    color = positionFactor
  )
) + 
  geom_smooth() +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age and Position",
    subtitle = "Generalized Additive Model with Players Who Were Once at Top of End-of-Season Depth Chart"
  ) +
  theme_classic()
```

### Plots of Individuals' Model-Implied Fantasy Points by Age {#sec-fantasyPointsIndividualByAgePlots}

```{r}
player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))

player_stats_seasonal_offense_subsetCC$fantasyPoints_quadratic <- predict(
  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
  newdata = player_stats_seasonal_offense_subsetCC
)

player_stats_seasonal_offense_subsetCC$fantasyPoints_gam <- predict(
  object = pointsPerSeason_gam,
  newdata = player_stats_seasonal_offense_subsetCC
)
```

```{r}
#| code-fold: true

zeroAge <- pointsPerSeason_positionAge_newData %>% 
  group_by(positionFactor) %>% 
  filter(fantasyPoints_gam < 0) %>% 
  slice(which.min(age))

peakAge <- pointsPerSeason_positionAge_newData %>% 
  group_by(positionFactor) %>% 
  slice(which.max(fantasyPoints_gam))

peakAge2 <- pointsPerSeason_positionAge_newData %>% 
  filter(age > 22) %>% 
  group_by(positionFactor) %>% 
  slice(which.max(fantasyPoints_gam))

qbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "QB")], 0)
fbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "FB")], 0)
rbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "RB")], 0)
wrPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "WR")], 0)
wrPeakAge2 <- round(peakAge2$age[which(peakAge$positionFactor == "WR")], 0)
tePeakAge <- round(peakAge$age[which(peakAge$positionFactor == "TE")], 0)

qbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "QB")], 0)
fbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "FB")], 0)
rbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "RB")], 0)
wrZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "WR")], 0)
teZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "TE")], 0)
```

#### Quarterbacks {#sec-fantasyPointsIndividualByAgePlotsQB}

A plot of Quarterbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeQB.
The model-implied peak of Quarterbacks' fantasy points is at age `{r} qbPeakAge`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} qbZeroAge`.

```{r}
#| label: fig-individualFantasyPointsByAgeQB
#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks. Overlaid with the Model-Implied Trajectory for Quarterbacks in Blue."
#| code-fold: true

plot_individualFantasyPointsByAgeQB <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "QB"),
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    group = player_id)) +
  geom_smooth(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    se = FALSE,
    linewidth = 0.5,
    color = "black") +
  geom_smooth(
    mapping = aes(
      x = age,
      y = fantasyPoints_gam
    ),
    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "QB"),
    inherit.aes = FALSE,
    se = TRUE,
    linewidth = 2
  ) +
  geom_point(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    size = 1,
    color = "transparent" # make points invisible but keep tooltips
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
  ) +
  theme_classic()

ggplotly(
  plot_individualFantasyPointsByAgeQB,
  tooltip = c("age","fantasyPoints_gam","text","label")
)
```

#### Fullbacks {#sec-fantasyPointsIndividualByAgePlotsFB}

A plot of Fullbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeFB.
The model-implied peak of Fullbacks' fantasy points is at age `{r} fbPeakAge`.
The model-predicted value of zero fantasy points for Fullbacks is at `{r} fbZeroAge`.

```{r}
#| label: fig-individualFantasyPointsByAgeFB
#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks in Blue."
#| code-fold: true

plot_individualFantasyPointsByAgeFB <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "FB"),
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    group = player_id)) +
  geom_smooth(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    se = FALSE,
    linewidth = 0.5,
    color = "black") +
  geom_smooth(
    mapping = aes(
      x = age,
      y = fantasyPoints_gam
    ),
    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "FB"),
    inherit.aes = FALSE,
    se = TRUE,
    linewidth = 2
  ) +
  geom_point(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    size = 1,
    color = "transparent" # make points invisible but keep tooltips
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Fullbacks"
  ) +
  theme_classic()

ggplotly(
  plot_individualFantasyPointsByAgeFB,
  tooltip = c("age","fantasyPoints_gam","text","label")
)
```

#### Running Backs {#sec-fantasyPointsIndividualByAgePlotsRB}

A plot of Running Backs' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeRB.
The model-implied peak of Running Backs' fantasy points is at age `{r} rbPeakAge`.
The model-predicted value of zero fantasy points for Running Backs is at `{r} rbZeroAge`.

```{r}
#| label: fig-individualFantasyPointsByAgeRB
#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs. Overlaid with the Model-Implied Trajectory for Running Backs in Blue."
#| code-fold: true

plot_individualFantasyPointsByAgeRB <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "RB"),
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    group = player_id)) +
  geom_smooth(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    se = FALSE,
    linewidth = 0.5,
    color = "black") +
  geom_smooth(
    mapping = aes(
      x = age,
      y = fantasyPoints_gam
    ),
    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "RB"),
    inherit.aes = FALSE,
    se = TRUE,
    linewidth = 2
  ) +
  geom_point(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    size = 1,
    color = "transparent" # make points invisible but keep tooltips
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Running Backs"
  ) +
  theme_classic()

ggplotly(
  plot_individualFantasyPointsByAgeRB,
  tooltip = c("age","fantasyPoints_gam","text","label")
)
```

#### Wide Receivers {#sec-fantasyPointsIndividualByAgePlotsWR}

A plot of Wide Receivers' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeWR.
The model-implied peaks of Wide Receivers' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
The model-predicted value of zero fantasy points for Wide Receivers is at `{r} wrZeroAge`.

```{r}
#| label: fig-individualFantasyPointsByAgeWR
#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers. Overlaid with the Model-Implied Trajectory for Wide Receivers in Blue."
#| code-fold: true

plot_individualFantasyPointsByAgeWR <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "WR"),
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    group = player_id)) +
  geom_smooth(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    se = FALSE,
    linewidth = 0.5,
    color = "black") +
  geom_smooth(
    mapping = aes(
      x = age,
      y = fantasyPoints_gam
    ),
    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "WR"),
    inherit.aes = FALSE,
    se = TRUE,
    linewidth = 2
  ) +
  geom_point(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    size = 1,
    color = "transparent" # make points invisible but keep tooltips
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
  ) +
  theme_classic()

ggplotly(
  plot_individualFantasyPointsByAgeWR,
  tooltip = c("age","fantasyPoints_gam","text","label")
)
```

#### Tight Ends {#sec-fantasyPointsIndividualByAgePlotsTE}

A plot of Tight Ends' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeTE.
The model-implied peak of Tight Ends' fantasy points is at age `{r} tePeakAge`.
The model-predicted value of zero fantasy points for Tight Ends is at `{r} teZeroAge`.

```{r}
#| label: fig-individualFantasyPointsByAgeTE
#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends. Overlaid with the Model-Implied Trajectory for Wide Tight Ends in Blue."
#| code-fold: true

plot_individualFantasyPointsByAgeTE <- ggplot(
  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "TE"),
  mapping = aes(
    x = age,
    y = fantasyPoints_gam,
    group = player_id)) +
  geom_smooth(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    se = FALSE,
    linewidth = 0.5,
    color = "black") +
  geom_smooth(
    mapping = aes(
      x = age,
      y = fantasyPoints_gam
    ),
    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "TE"),
    inherit.aes = FALSE,
    se = TRUE,
    linewidth = 2
  ) +
  geom_point(
    aes(
      x = age,
      y = fantasyPoints_gam,
      text = player_display_name, # add player name for mouse over tooltip
      label = season # add season for mouse over tooltip
    ),
    size = 1,
    color = "transparent" # make points invisible but keep tooltips
  ) +
  labs(
    x = "Player Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age: Tight Ends"
  ) +
  theme_classic()

ggplotly(
  plot_individualFantasyPointsByAgeTE,
  tooltip = c("age","fantasyPoints_gam","text","label")
)
```

### Summary of Findings {#sec-fantasyPointsByAgeExperienceSummary}

We applied mixed models with random intercepts and random slopes to allow our model estimates to account for the fact that different players have different starting points (intercepts) and different changes over time (slopes) in fantasy points.
A quadratic, inverted-U-shaped form as a function of age fit better than a linear form as a function of age in predicting players' fantasy points.
A generalized additive model that allowed further nonlinearity fit even better than the quadratic model.

Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
However, this conclusion would be wrong.
When we account for the longitudinal data (i.e., multiple observations over time for the same player) using mixed models, we observe that fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
In other words, the association between age and fantasy points differs at the person level versus the group level.
This is an example of [Simpson's paradox](#sec-simpsonsParadox).

The discrepancy between the positive or null association between age and players' fantasy points at the group level, and the negative association at the person level may be due, in part, to the selective attrition of players with age.
The players who play the longest will tend to be the highest performing players, whereas the poorest performing players will retire or get dropped from the team at younger ages.
Thus, the selective attrition of weaker players may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position), until the player eventually retires or is dropped from the team.
Selective attrition is common in [longitudinal studies](#sec-longitudinal) (such as this one) and [intervention studies](#sec-intervention).
For instance, attrition may be more likely for individuals from lower socioeconomic status backgrounds because they may face more challenges in continuing in longitudinal studies such as fewer financial resources, greater life stressors, etc.
In addition, people who experience side effects or lack of improvement may be more likely to drop out of [intervention studies](#sec-intervention).
Examining only those who completed treatment (an example of [selection bias](#sec-causalDiagramSelectionBias)) would make the intervention look more effective than it actually was because the people who stay in the study are those who experience the greatest improvement.
Thus, it is important to use approaches such as mixed models or other approaches that account for the multiple observations from the same person, that use all available information, and that do not exclude people who do not complete all portions of the study.

## Conclusion {#sec-mixedModelsConclusion}

Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
Inclusion of random effects allows the association between the predictor variables (the intercept and age) and the outcome variable (fantasy points) to differ for each individual in the grouping level (in this case, each player).
This allows for more accurately predicting phenomena.
Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
However, this conclusion would be wrong.
When we account for the longitudinal data using mixed models, we observe that players' fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
In other words, the association between age and fantasy points differs at the person level versus the group level, which is an example of [Simpson's paradox](#sec-simpsonsParadox).
In sum, mixed models are valuable for examining associations between variables when there are multiple levels of data (i.e., multiple observations within the same unit, known as clustering or nesting).
It is important not to confuse the association at one level (e.g., group level) with the association at another level (e.g., person level), which is an example of the [ecological fallacy](#sec-fallaciesEcological).

::: {.content-visible when-format="html"}

## Session Info {#sec-mixedModelsSessionInfo}

```{r}
sessionInfo()
```

:::