### 2024-08-11 time-series-analysis.qmd
20240811 - time series analysis

# Time Series Analysis {#sec-timeSeries}

## Getting Started {#sec-timeSeriesGettingStarted}

### Load Packages {#sec-timeSeriesLoadPackages}

```{r}
library("xts")
library("zoo")
library("forecast")
library("tidyverse")
```

### Load Data {#sec-timeSeriesLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

## Overview of Time Series Analysis {#sec-timeSeriesOverview}

Time series analysis is useful when trying to generate forecasts from longitudinal data.
There many different types of time series analyses.
For simplicity, in this chapter, we use autoregressive integrated moving average (ARIMA) models to demonstrate one approach to time series analysis.
ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values (i.e., autoregressive effects and moving average effects).
This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
Nevertheless, ARIMA models are widely used in forecasting financial markets and economic indicators.
Thus, it is a useful technique to learn.

Adapted from: <https://rc2e.com/timeseriesanalysis> (archived at <https://perma.cc/U5P6-2VWC>).

## Create the Time Series Objects {#sec-createTimeSeriesObjects}

```{r}
weeklyFantasyPoints_tomBrady <- player_stats_weekly_offense %>% 
  filter(
    player_id == "00-0019596" | player_display_name == "Tom Brady")

weeklyFantasyPoints_peytonManning <- player_stats_weekly_offense %>% 
  filter(
    player_id == "00-0010346" | player_display_name == "Peyton Manning")

ts_tomBrady <- xts::xts(
  x = weeklyFantasyPoints_tomBrady["fantasy_points"],
  order.by = weeklyFantasyPoints_tomBrady$gameday)

ts_peytonManning <- xts::xts(
  x = weeklyFantasyPoints_peytonManning["fantasy_points"],
  order.by = weeklyFantasyPoints_peytonManning$gameday)

ts_tomBrady
ts_peytonManning

ts_combined <- merge(
  ts_tomBrady,
  ts_peytonManning
)

names(ts_combined) <- c("Tom Brady","Peyton Manning")
```

## Plot the Time Series {#sec-plotTimeSeries}

```{r}
plot(
  ts_tomBrady,
  main = "Tom Brady's Fantasy Points by Game")
```

```{r}
plot(
  ts_combined,
  legend,
  legend.loc = "topright",
  main = "Fantasy Points by Game")
```

## Rolling Mean/Median {#sec-rollingMeanMedian}

```{r}
zoo::rollmean(
  x = ts_tomBrady,
  k = 5)

zoo::rollmedian(
  x = ts_tomBrady,
  k = 5)
```

## Autocorrelation {#sec-autoCorrelation}

```{r}
acf(ts_tomBrady)
```

```{r}
Box.test(ts_tomBrady)
```

## Fit an Autoregressive Integrated Moving Average Model {#sec-arimaModel}

```{r}
forecast::auto.arima(ts_tomBrady)
forecast::auto.arima(ts_peytonManning)
```

```{r}
arima_tomBrady <- arima(
  ts_tomBrady,
  order = c(5, 1, 4))

summary(arima_tomBrady)
confint(arima_tomBrady)
forecast::checkresiduals(arima_tomBrady)

arima_tomBrady_removeNonSigTerms <- arima(
  ts_tomBrady,
  order = c(5, 1, 4),
  fixed = c(NA, NA, 0, NA, NA, NA, NA, NA, NA))

summary(arima_tomBrady_removeNonSigTerms)
confint(arima_tomBrady_removeNonSigTerms)
forecast::checkresiduals(arima_tomBrady_removeNonSigTerms)

arima_peytonManning <- arima(
  ts_peytonManning,
  order = c(1, 0, 1))

summary(arima_peytonManning)
confint(arima_peytonManning)
forecast::checkresiduals(arima_peytonManning)
```

## Generate the Model Forecasts {#sec-arimaForecasts}

```{r}
forecast_tomBrady <- forecast::forecast(
  arima_tomBrady,
  level = c(80, 95)) # 80% and 95% confidence intervals

forecast_peytonManning <- forecast::forecast(
  arima_peytonManning,
  level = c(80, 95)) # 80% and 95% confidence intervals

forecast_tomBrady
forecast_peytonManning
```

## Plot the Model Forecasts {#sec-plotARIMAforecast}

```{r}
forecast::autoplot(forecast_tomBrady) + 
  labs(
    x = "Game Number",
    y = "Fantasy Points",
    title = "Tom Brady's Historical and Projected Fantasy Points by Game"
  ) +
  theme_classic()
```

## Conclusion {#sec-timeSeriesConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-timeSeriesSessionInfo}

```{r}
sessionInfo()
```

:::