### 2024-11-27 bugs.qmd
20241126 - calculate_stats() function from nflfastR, TODOs, and bugs

# Bugs {#sec-bugs}

Bugs to report (`nflreadr` or `nflfastR`):

- `players` database has a season column but it is only populated for some players, and it is only populated with 2020 (<https://github.com/nflverse/nflreadr/issues/238>)
- `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)
- common naming convention for key variables, for purposes of merging (<https://github.com/nflverse/nflreadr/issues/237>)
- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (<https://github.com/nflverse/nflfastR/issues/480>)
- calculate career stats using `calculate_stats()` function