### 2025-04-07 correlation.qmd
20250407 - correlation

# Correlation Analysis {#sec-correlation}

## Getting Started {#sec-correlationGettingStarted}

### Load Packages {#sec-correlationLoadPackages}

```{r}
library("petersenlab")
library("XICOR")
library("psych")
library("DescTools")
library("tidyverse")
```

### Load Data {#sec-correlationLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

## Overview of Correlation {#sec-correlationOverview}

Correlation is an index of the association between variables.
Covariance is the association between variables and in an unstandardized metric that differs for variables with different scales.
By contrast, correlation is in a standardized metric that does not differ for variables with different scales.
When examining the association between variables that are [interval](#sec-interval) or [ratio](#sec-ratio) levels of measurement, Pearson correlation is used.
When examining the association between variables that are [ordinal](#sec-ordinal) in level of measurement, Spearman correlation is used.
Pearson correlation is an index of the *linear* association between variables.
If a nonlinear association is present, other indices like xi [$\xi$; @Chatterjee2021] and distance correlation coefficients are better suited to detect the association.

## The Correlation Coefficient ($r$)

The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
As noted in @sec-statisticalTestsPartitionedVariance, the correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance, as in @eq-correlationPartitionedVariance.

The correlation coefficient ranges from −1.0 to +1.0.\index{correlation}
The correlation coefficient ($r$) tells you two things: (1) the direction (sign) of the association (positive or negative) and (2) the magnitude of the association.\index{correlation}
If the correlation coefficient is positive, the association is positive.\index{correlation}
If the correlation coefficient is negative, the association is negative.\index{correlation}
If the association is positive, as `X` increases, `Y` increases (or conversely, as `X` decreases, `Y` decreases).\index{correlation}
If the association is negative, as `X` increases, `Y` decreases (or conversely, as `X` decreases, `Y` increases).\index{correlation}
The smaller the absolute value of the correlation coefficient (i.e., the closer the $r$ value is to zero), the weaker the association and the flatter the slope of the best-fit line in a scatterplot.\index{correlation}
The larger the absolute value of the correlation coefficient (i.e., the closer the absolute value of the $r$ value is to one), the stronger the association and the steeper the slope of the best-fit line in a scatterplot.\index{correlation}
See @fig-rangeOfCorrelations for a range of different correlation coefficients and what some example data may look like for each direction and strength of association.\index{correlation}

```{r}
#| label: fig-rangeOfCorrelations
#| fig-cap: "Correlation Coefficients."
#| fig-alt: "Correlation Coefficients."
#| fig-height: 12
#| code-fold: true
#| message: false

set.seed(52242)
correlations <- data.frame(criterion = rnorm(1000))

correlations$v1 <- complement(correlations$criterion, -1)
correlations$v2 <- complement(correlations$criterion, -.9)
correlations$v3 <- complement(correlations$criterion, -.8)
correlations$v4 <- complement(correlations$criterion, -.7)
correlations$v5 <- complement(correlations$criterion, -.6)
correlations$v6 <- complement(correlations$criterion, -.5)
correlations$v7 <- complement(correlations$criterion, -.4)
correlations$v8 <- complement(correlations$criterion, -.3)
correlations$v9 <- complement(correlations$criterion, -.2)
correlations$v10 <-complement(correlations$criterion, -.1)
correlations$v11 <-complement(correlations$criterion, 0)
correlations$v12 <-complement(correlations$criterion, .1)
correlations$v13 <-complement(correlations$criterion, .2)
correlations$v14 <-complement(correlations$criterion, .3)
correlations$v15 <-complement(correlations$criterion, .4)
correlations$v16 <-complement(correlations$criterion, .5)
correlations$v17 <-complement(correlations$criterion, .6)
correlations$v18 <-complement(correlations$criterion, .7)
correlations$v19 <-complement(correlations$criterion, .8)
correlations$v20 <-complement(correlations$criterion, .9)
correlations$v21 <-complement(correlations$criterion, 1)

par(mfrow = c(7,3), mar = c(1, 0, 1, 0))

# -1.0
plot(correlations$criterion, correlations$v1, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v1)$estimate, 2))))
abline(lm(v1 ~ criterion, data = correlations), col = "black")

# -.9
plot(correlations$criterion, correlations$v2, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v2)$estimate, 2))))
abline(lm(v2 ~ criterion, data = correlations), col = "black")

# -.8
plot(correlations$criterion, correlations$v3, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v3)$estimate, 2))))
abline(lm(v3 ~ criterion, data = correlations), col = "black")

# -.7
plot(correlations$criterion, correlations$v4, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v4)$estimate, 2))))
abline(lm(v4 ~ criterion, data = correlations), col = "black")

# -.6
plot(correlations$criterion, correlations$v5, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v5)$estimate, 2))))
abline(lm(v5 ~ criterion, data = correlations), col = "black")

# -.5
plot(correlations$criterion, correlations$v6, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v6)$estimate, 2))))
abline(lm(v6 ~ criterion, data = correlations), col = "black")

# -.4
plot(correlations$criterion, correlations$v7, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v7)$estimate, 2))))
abline(lm(v7 ~ criterion, data = correlations), col = "black")

# -.3
plot(correlations$criterion, correlations$v8, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v8)$estimate, 2))))
abline(lm(v8 ~ criterion, data = correlations), col = "black")

# -.2
plot(correlations$criterion, correlations$v9, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v9)$estimate, 2))))
abline(lm(v9 ~ criterion, data = correlations), col = "black")

# -.1
plot(correlations$criterion, correlations$v10, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v10)$estimate, 2))))
abline(lm(v10 ~ criterion, data = correlations), col = "black")

# 0.0
plot(correlations$criterion, correlations$v11, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v11)$estimate, 2))))
abline(lm(v11 ~ criterion, data = correlations), col = "black")

# 0.1
plot(correlations$criterion, correlations$v12, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v12)$estimate, 2))))
abline(lm(v12 ~ criterion, data = correlations), col = "black")

# 0.2
plot(correlations$criterion, correlations$v13, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v13)$estimate, 2))))
abline(lm(v13 ~ criterion, data = correlations), col = "black")

# 0.3
plot(correlations$criterion, correlations$v14, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v14)$estimate, 2))))
abline(lm(v14 ~ criterion, data = correlations), col = "black")

# 0.4
plot(correlations$criterion, correlations$v15, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v15)$estimate, 2))))
abline(lm(v15 ~ criterion, data = correlations), col = "black")

# 0.5
plot(correlations$criterion, correlations$v16, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v16)$estimate, 2))))
abline(lm(v16 ~ criterion, data = correlations), col = "black")

# 0.6
plot(correlations$criterion, correlations$v17, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v17)$estimate, 2))))
abline(lm(v17 ~ criterion, data = correlations), col = "black")

# 0.7
plot(correlations$criterion, correlations$v18, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v18)$estimate, 2))))
abline(lm(v18 ~ criterion, data = correlations), col = "black")

# 0.8
plot(correlations$criterion, correlations$v19, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v19)$estimate, 2))))
abline(lm(v19 ~ criterion, data = correlations), col = "black")

# 0.9
plot(correlations$criterion, correlations$v20, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v20)$estimate, 2))))
abline(lm(v20 ~ criterion, data = correlations), col = "black")

# 1.0
plot(correlations$criterion, correlations$v21, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v21)$estimate, 2))))
abline(lm(v21 ~ criterion, data = correlations), col = "black")

invisible(dev.off()) #par(mfrow = c(1,1))
```

See @fig-interepretingCorrelationCoefficients for the interpretation of the magnitude and direction (sign) of various correlation coefficients.

```{r}
#| label: fig-interepretingCorrelationCoefficients
#| fig-cap: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
#| fig-alt: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
#| fig-height: 9
#| fig-width: 9
#| code-fold: true
#| message: false

library("patchwork")

set.seed(52242)
correlations2 <- data.frame(criterion = rnorm(15))

correlations2$v1 <- complement(correlations2$criterion, -1)
correlations2$v2 <- complement(correlations2$criterion, -.9)
correlations2$v3 <- complement(correlations2$criterion, -.8)
correlations2$v4 <- complement(correlations2$criterion, -.7)
correlations2$v5 <- complement(correlations2$criterion, -.6)
correlations2$v6 <- complement(correlations2$criterion, -.5)
correlations2$v7 <- complement(correlations2$criterion, -.4)
correlations2$v8 <- complement(correlations2$criterion, -.3)
correlations2$v9 <- complement(correlations2$criterion, -.2)
correlations2$v10 <-complement(correlations2$criterion, -.1)
correlations2$v11 <-complement(correlations2$criterion, 0)
correlations2$v12 <-complement(correlations2$criterion, .1)
correlations2$v13 <-complement(correlations2$criterion, .2)
correlations2$v14 <-complement(correlations2$criterion, .3)
correlations2$v15 <-complement(correlations2$criterion, .4)
correlations2$v16 <-complement(correlations2$criterion, .5)
correlations2$v17 <-complement(correlations2$criterion, .6)
correlations2$v18 <-complement(correlations2$criterion, .7)
correlations2$v19 <-complement(correlations2$criterion, .8)
correlations2$v20 <-complement(correlations2$criterion, .9)
correlations2$v21 <-complement(correlations2$criterion, 1)

# -1.0
p1 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v1
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Perfect Negative Association",
    subtitle = expression(paste(italic("r"), " = ", "−1.0"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# -0.9
p2 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v2
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Strong Negative Association",
    subtitle = expression(paste(italic("r"), " = ", "−.9"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# -0.5
p3 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v6
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Moderate Negative Association",
    subtitle = expression(paste(italic("r"), " = ", "−.5"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# -0.2
p4 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v9
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Weak Negative Association",
    subtitle = expression(paste(italic("r"), " = ", "−.2"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# 0.0
p5 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v11
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "No Association",
    subtitle = expression(paste(italic("r"), " = ", ".0"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# 0.2
p6 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v13
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Weak Positive Association",
    subtitle = expression(paste(italic("r"), " = ", ".2"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# 0.5
p7 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v16
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Moderate Positive Association",
    subtitle = expression(paste(italic("r"), " = ", ".5"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# 0.9
p8 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v20
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Strong Positive Association",
    subtitle = expression(paste(italic("r"), " = ", ".9"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

# 1.0
p9 <- ggplot(
  data = correlations2,
  mapping = aes(
    x = criterion,
    y = v21
  )
) + 
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE) +
  labs(
    title = "Perfect Positive Association",
    subtitle = expression(paste(italic("r"), " = ", "1.0"))
  ) +
  theme_classic(
    base_size = 12) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())

p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 +
  plot_layout(
    ncol = 3,
    heights = 1,
    widths = 1)
```

An interactive visualization by @Magnusson2023 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)

## Examples {#sec-correlationExamples}

Is there an association between a player's height and the number of fantasy points they score?
We can examine this possibility separately by position.

First, let's examine descriptive statistics (height is measured in inches):

```{r}
player_stats_seasonal %>% 
  dplyr::select(height, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

### Covariance {#sec-correlationExamplesCovariance}

The following syntax shows the variance-covariance of a set of variables, where the covariance of a variable with itself is the variable's variance.

```{r}
cov(
  player_stats_seasonal[,c("height","weight","fantasyPoints")],
  use = "pairwise.complete.obs")
```

If you just want the covariance between two variables, you can use the following syntax.

```{r}
cov(
  player_stats_seasonal$height,
  player_stats_seasonal$fantasyPoints,
  use = "pairwise.complete.obs")
```

Thus, it appears there is a negative covariance between height and fantasy points in the whole population of National Football League (NFL) players.

As shown below, the negative covariance between height and fantasy points holds when examining just Quarterbacks, Running Backs, Wide Receivers, and Tight Ends.

```{r}
cov(
  player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")) %>% select(height,fantasyPoints),
  use = "pairwise.complete.obs")
```

We can also examine the association separately by position.
Here is the association for Quarterbacks:

```{r}
cov(
  player_stats_seasonal %>% filter(position == "QB") %>% select(height,fantasyPoints),
  use = "pairwise.complete.obs")
```

Here is the association for Running Backs:

```{r}
cov(
  player_stats_seasonal %>% filter(position == "RB") %>% select(height,fantasyPoints),
  use = "pairwise.complete.obs")
```

Here is the association for Wide Receivers:

```{r}
cov(
  player_stats_seasonal %>% filter(position == "WR") %>% select(height,fantasyPoints),
  use = "pairwise.complete.obs")
```

Here is the association for Tight Ends:

```{r}
cov(
  player_stats_seasonal %>% filter(position == "TE") %>% select(height,fantasyPoints),
  use = "pairwise.complete.obs")
```

Interestingly, there is a *positive* covariance between height and fantasy points when examining each position separately.
This is an example of [Simpson's paradox](#sec-simpsonsParadox), where the sign of an association differs at different levels of analysis, as described in @sec-simpsonsParadox.
There is a *negative* covariance between height and fantasy points when examining Quarterbacks, Running Backs, Wide Receivers, and Tight Ends altogether, but there is a *positive* covariance between height and fantasy points when examining the association between each position separately.
That is, if you were to examine all positions together, you might assume that being taller might be disadvantageous to scoring fantasy fantasy points; however, once we examine the association within a given position, it emerges that height appears to be somewhat advantageous, if anything, for scoring fantasy points.

The covariance is an unstandardized index of the association between two variables.
However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
The [Pearson correlation coefficient](Pearson Correlation), $r$, is an example of a standardized index of the covariance between two variables.

### Pearson Correlation {#sec-correlationExamplesPearson}

Pearson correlation assumes that the variables are [interval](#sec-interval) or [ratio](#sec-interval) level of measurement.
If the data are [ordinal](#sec-ordinal), [Spearman correlation](#sec-correlationExamplesSpearman) is more appropriate.

The following syntax shows the Pearson correlation matrix of a set of variables.

```{r}
cor(
  player_stats_seasonal[,c("height","weight","fantasyPoints")],
  use = "pairwise.complete.obs")
```

If you just want the Pearson correlation between two variables, you can use the following syntax.

```{r}
cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal
)
```

```{r}
#| include: false

pearson_test <- cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal
)

pearson_r <- pearson_test$estimate
pearson_df <- pearson_test$parameter
pearson_p <- pearson_test$p.value
```

The $r$ value (`{r} pearson_r`) is slightly negative.
The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
The $p$-value is less than .05, so it is a statistically significant correlation.
We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$`{r} pearson_df`$] =$ `{r} pearson_r`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.

Now let's examine the association separately by position.
Here is the association for Quarterbacks:

```{r}
cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal %>% 
    filter(position == "QB")
)
```

Here is the association for Running Backs:

```{r}
cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal %>% 
    filter(position == "RB")
)
```

Here is the association for Wide Receivers:

```{r}
cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal %>% 
    filter(position == "WR")
)
```

Here is the association for Tight Ends:

```{r}
cor.test(
  ~ height + fantasyPoints,
  data = player_stats_seasonal %>% 
    filter(position == "TE")
)
```

The sign of the association was positive for each position, suggesting that, for a given position (among Quarterbacks, Running Backs, Wide Receivers, and Tight Ends), taller players tend to score more fantasy points.
However, the [effect size](#sec-practicalSignificance) is small.
Moreover, as described in @sec-correlation-correlationAndCausation, just because there is an association between variables does not mean that the association reflects a causal effect.

Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance test of a correlation.
We could address this assumption by analyzing only one season from each player.
However, for simplicity, we present results from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression).

### Spearman Correlation {#sec-correlationExamplesSpearman}

### Nonlinear Correlation {#sec-correlationExamplesNonlinear}

A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2021]:

### Correlation Matrix {#sec-correlationExamplesMatrix}

### Correlogram {#sec-correlationExamplesCorrelogram}

## Impact of Outliers {#sec-correlationOutliers}

## Impact of Restricted Range {#sec-correlationRestrictedRange}

## Correlation Does Not Imply Causation {#sec-correlation-correlationAndCausation}

As described in @sec-correlationCausation, correlation does not imply causation.
There are several reasons (described in @sec-correlationCausation) that, just because `X` is correlated with `Y` does not necessarily mean that `X` causes `Y`.
However, correlation can still be useful.
In order for two processes to be causally related, they must be associated.
That is, association is necessary but insufficient for causality.

## Conclusion {#sec-correlationConclusion}

Correlation is an index of the association between variables.
The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
Although correlation does not imply causation, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.

## Session Info {#sec-correlationSessionInfo}

::: {.content-visible when-format="html"}

```{r}
sessionInfo()
```

:::