### 2025-06-19 simulation.qmd
20250619 - Monte Carlo simulation

# Simulation: Bootstrapping and the Monte Carlo Method {#sec-simulation}

## Getting Started {#sec-simulationGettingStarted}

### Load Packages {#sec-simulationLoadPackages}

```{r}
library("ffanalytics")
library("data.table")
library("future")
library("future.apply")
library("progressr")
library("SimDesign")
library("fitdistrplus")
library("sn")
library("tidyverse")
```

### Load Data {#sec-simulationLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
```

```{r}
load(file = "./data/players_projectedPoints_seasonal.RData")
load(file = "./data/nfl_playerIDs.RData")
```

## Overview {#sec-simulationOverview}

A simulation is an "imitative representation" of a phenomenon that could exist the real world.
In statistics, simulations are computer-driven investigations to better understand a phenomenon by studying its behavior under different conditions.
For instance, we might want to determine the likely range of outcomes for a player, in terms of the range of fantasy points that a player might score over the course of a season.
Simulations can be conducted in various ways.
Two common ways of conducting simulations are via [bootstrapping](#sec-simulationOverviewBootstrapping) and via [Monte Carlo simulation](#sec-simulationOverviewMonteCarlo).

### Bootstrapping {#sec-simulationOverviewBootstrapping}

Bootstrapping involves repeated resampling (with replacement) from observed data.
For instance, if we have 100 sources provide projections for a player, we could estimate the most likely range of fantasy points for the player by repeatedly sampling from the 100 projections.

### Monte Carlo Simulation {#sec-simulationOverviewMonteCarlo}

Monte Carlo simulation involves repeated random sampling from a known distribution [@Sigal2016].
For instance, if we know the population distribution for the likely outcomes for a player—e.g., a normal distribution with a known mean (e.g., 150 points) and standard deviation (e.g., 20 points)—we can repeatedly sample randomly from this distribution.
The distribution could be, as examples, a normal distribution, a log-normal distribution, a binomial distribution, a chi-square distribution, etc. [@Sigal2016].
The distribution provides a probability density function, which indicates the probability that any particular value would be observed if the data arose from that distribution.

## Simulation of Projected Statistics and Points {#sec-simulationProjectedPoints}

### Bootstrapping {#sec-simulationProjectedPointsBootstrapping}

#### Prepare Data {#sec-simulationProjectedPointsBootstrappingPrepareData}

```{r}
all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
```

```{r}
#| label: vars-to-bootstrap-by-position

vars_by_pos <- list(
  QB = c(
    "games",
    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "fumbles_lost", "fumbles_total", "two_pts",
    "sacks",
    "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",
    "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds"
    ),
  RB = c(
    "games",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
    "fumbles_lost", "fumbles_total", "two_pts",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
  ),
  WR = c(
    "games",
    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
    "fumbles_lost", "fumbles_total", "two_pts",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
  ),
  TE = c(
    "games",
    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
    "fumbles_lost", "fumbles_total", "two_pts",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
  ),
  K = c(
    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
    "fg", "fg_att", "fg_miss", "xp", "xp_att"
  ),
  D = c(
    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
  ),
  DL = c(
    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
  ),
  LB = c(
    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
  ),
  DB = c(
    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
  ),
  DST = c(
    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
  )
)
```

#### Bootstrapping Function {#sec-simulationProjectedPointsBootstrappingFunction}

For performing the bootstrapping, we leverage the `data.table` [@R-data.table], `future` [@R-future], and `future.apply` [@R-future.apply] packages for speed (by using parallel processing) and memory efficiency.
We use the `progressr` [@R-progressr] package to create a progress bar.

```{r}
#| label: function-for-bootstrap-simulation

bootstrapSimulation <- function(
    projectedStats,
    vars_by_pos,
    n_iter = 10000,
    seed = NULL,
    progress = TRUE) {

  dt <- data.table::as.data.table(projectedStats) # use data.table for speed
  all_ids <- unique(dt$id)

  if (!is.null(seed)) set.seed(seed)

  future::plan(future::multisession) # parallelize tasks across multiple background R sessions using multiple cores to speed up simulation

  if (progress) progressr::handlers("txtprogressbar") # specify progress-bar style

  results <- progressr::with_progress({ # wrap in with_progress for progress bar
    p <- if (progress) progressr::progressor(along = all_ids) else NULL # create progressor for progress bar

    future.apply::future_lapply(
      all_ids, # apply the function below to each player using a parallelized loop
      function(player_id) {
        if (!is.null(p)) p() # advance progress bar

        player_data <- dt[id == player_id]
        player_pos  <- unique(player_data$pos)

        if (length(player_pos) != 1 || !player_pos %in% names(vars_by_pos)) return(NULL)

        stat_vars <- vars_by_pos[[player_pos]] # pull the relevant stat variables to simulate for this player's position
        out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)

        for (var in stat_vars) { # loop over each stat variable that should be simulated for the player's position
          if (var %in% names(player_data)) {
            non_na_values <- player_data[[var]][!is.na(player_data[[var]])] # pull non-missing values of the stat for the player (from all projection sources)

            if (length(non_na_values) > 0) {
              out[[var]] <- sample(non_na_values, n_iter, replace = TRUE) # if there are valid values, sample with replacement to simulate n_iter values
            } else {
              out[[var]] <- NA_real_ # specify a numeric missing value (if all values were missing)
            }
          } else {
            out[[var]] <- NA_real_ # specify a numeric missing value (if the stat variable doesn't exist)
          }
        }

        return(out)
      },
      future.seed = TRUE # ensures that each parallel process gets a reproducible random seed
    )
  })

  data.table::rbindlist(results, use.names = TRUE, fill = TRUE) # combines all the individual player results into one large data table, aligning columns by name; fill = TRUE ensures that missing columns are filled with NA where necessary
}
```

#### Run the Bootstrapping Simulation {#sec-simulationProjectedPointsBootstrappingRun}

```{r}
#| label: code-to-run-bootstrap-simulation
#| eval: false

bootstappedStats <- bootstrapSimulation(
  projectedStats = all_proj,
  vars_by_pos = vars_by_pos,
  n_iter = 5000,
  seed = 52242)
```

```{r}
#| label: run-bootstrap-simulation
#| include: false

bootstappedStats <- bootstrapSimulation(
  projectedStats = all_proj,
  vars_by_pos = vars_by_pos,
  n_iter = 5000,
  seed = 52242,
  progress = FALSE)
```

```{r}
#| label: free-up-memory1
#| include: false

rm(all_proj)
gc()
```

#### Score Fantasy Points from the Simulation {#sec-simulationProjectedPointsBootstrappingScore}

```{r}
#| label: put-simulated-data-in-format-for-scoring1

data.table::setnames(bootstappedStats, "iteration", "data_src") # data.table equivalent to: bootstappedStats$data_src <- bootstappedStats$iteration

bootstappedStatsByPosition <- split(
  bootstappedStats,
  by = "pos",
  keep.by = TRUE)
```

```{r}
#| label: free-up-memory2
#| include: false

rm(bootstappedStats)
gc()
```

```{r}
#| label: put-simulated-data-in-format-for-scoring2

bootstappedStatsByPosition <- lapply(
  bootstappedStatsByPosition,
  setDF)

attr(bootstappedStatsByPosition, "season") <- 2024
attr(bootstappedStatsByPosition, "week") <- 0
```

```{r}
#| label: free-up-memory3
#| include: false

gc()
```

```{r}
#| label: score-bootstrap-simulation

bootstrappedFantasyPoints <- ffanalytics:::source_points(
  data_result = bootstappedStatsByPosition,
  scoring_rules = ffanalytics::scoring)
```

```{r}
#| label: free-up-memory4
#| include: false

rm(bootstappedStatsByPosition)
gc()
```

```{r}
#| label: postprocessing-bootstrap-simulation

bootstrappedFantasyPoints$iteration <- bootstrappedFantasyPoints$data_src
bootstrappedFantasyPoints$data_src <- NULL

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
    by = c("id" = "mfl_id")
  )

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  rename(projectedPoints = raw_points)
```

```{r}
#| label: free-up-memory5
#| include: false

gc()
```

```{r}
#| label: compute-fantasy-points-example
#| eval: false
#| include: false

# https://github.com/FantasyFootballAnalytics/ffanalytics/blob/08cab3cbe6eb645b853bb814e671e186b78a0837/R/calc_projections.R#L218-L253

# This gives an example for the data format accepted by the source_points() function
l = ffanalytics:::scrape_espn(season = 2024, week = 0)

# The two important things for the "data_result" object going into the source_points()
# function are the format of the list (with positions as names) and the attributes:
# season and week

# Before sending the data_result to the source_points() function make sure it has these attributes set 
# (e.g., this is for 2024 week 0)

attr(data_result, "season") <- 2024
attr(data_result, "week") <- 0

df <- ffanalytics:::source_points(data_result = l, scoring_rules = ffanalytics::scoring)
```

#### Summarize Players' Distribution of Projected Fantasy Points {#sec-simulationProjectedPointsBootstrappingSummarize}

```{r}
#| label: summarize-bootstrap-simulation-scores

bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
  group_by(id) %>% 
  summarise(
    mean = mean(projectedPoints, na.rm = TRUE),
    SD = sd(projectedPoints, na.rm = TRUE),
    min = min(projectedPoints, na.rm = TRUE),
    max = max(projectedPoints, na.rm = TRUE),
    q10 = quantile(projectedPoints, .10, na.rm = TRUE), # 10th quantile
    q90 = quantile(projectedPoints, .90, na.rm = TRUE), # 90th quantile
    range = max(projectedPoints, na.rm = TRUE) - min(projectedPoints, na.rm = TRUE),
    IQR = IQR(projectedPoints, na.rm = TRUE),
    MAD = mad(projectedPoints, na.rm = TRUE),
    CV = SD/mean,
    median = median(projectedPoints, na.rm = TRUE),
    pseudomedian = DescTools::HodgesLehmann(projectedPoints, na.rm = TRUE),
    mode = petersenlab::Mode(projectedPoints, multipleModes = "mean"),
    skewness = psych::skew(projectedPoints, na.rm = TRUE),
    kurtosis = psych::kurtosi(projectedPoints, na.rm = TRUE)
  )
```

```{r}
#| label: free-up-memory6
#| include: false

gc()
```

#### View Players' Distribution of Projected Fantasy Points {#sec-simulationProjectedPointsBootstrappingView}

```{r}
#| label: merge-player-info

bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints_summary %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team","position")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, mean:kurtosis, everything()) %>% 
  arrange(-mean)
```

```{r}
#| label: free-up-memory7
#| include: false

gc()
```

```{r}
#| label: display-bootstrap-simulation-summaries-qb

bootstrappedFantasyPoints_summary %>% 
  filter(position == "QB") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-rb

bootstrappedFantasyPoints_summary %>% 
  filter(position == "RB") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-wr

bootstrappedFantasyPoints_summary %>% 
  filter(position == "WR") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-te

bootstrappedFantasyPoints_summary %>% 
  filter(position == "TE") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-k

bootstrappedFantasyPoints_summary %>% 
  filter(position == c("K","PK")) %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-dl

bootstrappedFantasyPoints_summary %>% 
  filter(position %in% c("DL","DT","DE")) %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-lb

bootstrappedFantasyPoints_summary %>% 
  filter(position %in% c("LB","MLB","OLB")) %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: display-bootstrap-simulation-summaries-db

bootstrappedFantasyPoints_summary %>% 
  filter(position %in% c("DB","S","CB")) %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```

```{r}
#| label: identify-players-with-lo-or-hi-cv
#| eval: false
#| include: false

bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(CV) %>% head
bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(-CV) %>% head
```

An example distribution of projected fantasy points is in @fig-bootstrapping1.

```{r}
#| label: fig-bootstrapping1
#| fig-cap: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."
#| fig-alt: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."

ggplot2::ggplot(
  data = bootstrappedFantasyPoints %>%
    filter(pos == "QB" & name == "Patrick Mahomes"),
  mapping = aes(
    x = projectedPoints)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  #coord_cartesian(
  #  xlim = c(0,400)) +
  labs(
    x = "Fantasy Points",
    y = "Density",
    title = "Distribution of Projected Fantasy Points for Patrick Mahomes"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

Projections of two players—one with relatively narrow uncertainty and one with relatively wide uncertainty—are depicted in @fig-bootstrapping2.

```{r}
#| label: fig-bootstrapping2
#| fig-cap: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping. There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."
#| fig-alt: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping. There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."

ggplot2::ggplot(
  data = bootstrappedFantasyPoints %>%
    filter(pos == "QB" & (name %in% c("Dak Prescott", "Drake Maye"))),
  mapping = aes(
    x = projectedPoints,
    group = name,
    #color = name,
    fill = name)
) +
  geom_histogram(
    aes(y = after_stat(density))
  ) +
  geom_density(
    alpha = 0.6 # add transparency
  ) +
  coord_cartesian(
    xlim = c(0,NA),
    expand = FALSE) +
  #geom_rug() +
  labs(
    x = "Fantasy Points",
    y = "Density",
    fill = "",
    color = "",
    title = "Distribution of Projected Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
#| label: free-up-memory8
#| include: false

rm(bootstrappedFantasyPoints)
gc()
```

### Monte Carlo Simulation {#sec-simulationProjectedPointsMonteCarlo}

#### `SimDesign` Package {#sec-simulationProjectedPointsMonteCarloSimDesign}

You can generate a template for Monte Carlo simulations in the `SimDesign` [@R-SimDesign; @Chalmers2020] package using the following code:

```{r}
SimDesign::SimFunctions()
```

```{r}
#| eval: false
#| include: false

#-------------------------------------------------------------------

Design <- SimDesign::createDesign(
  factor1 = NA,
  factor2 = NA)

#-------------------------------------------------------------------

Generate <- function(condition, fixed_objects) {
  dat <- data.frame()
  dat
}

Analyse <- function(condition, dat, fixed_objects) {
  ret <- nc(stat1 = NaN, stat2 = NaN)
  ret
}

Summarise <- function(condition, results, fixed_objects) {
  ret <- c(bias = NaN, RMSE = NaN)
  ret
}

#-------------------------------------------------------------------

res <- SimDesign::runSimulation(
  design = Design,
  replications = 2,
  generate = Generate,
  analyse = Analyse,
  summarise = Summarise)

res
```

#### Prepare Data {#sec-simulationProjectedPointsMonteCarloPrepareData}

```{r}
#| include: false

all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
```

```{r}
all_proj <- all_proj %>% 
  rename(projectedPoints = raw_points)
```

#### Optimal Distribution for Each Player {#sec-simulationProjectedPointsMonteCarloOptimalDistribution}

```{r}
# Function to identify the "best" distribution (Normal vs Skew‑Normal) for every player (tries both families and picks by AIC; uses empirical distribution if fewer than 2 unique scores)
fit_best <- function(x) {
  # Basic checks
  if (length(unique(x)) < 2 || all(is.na(x))) { # Use empirical distribution if there are fewer than 2 unique scores
    return(list(type = "empirical", empirical = x))
  }

  # Try Normal Distribution
  fit_norm <- tryCatch(
    fitdistrplus::fitdist(x, distr = "norm"),
    error = function(e) NULL
  )

  # Try Skew-Normal Distribution
  fit_skew <- tryCatch(
    sn::selm(x ~ 1),
    error = function(e) NULL
  )

  # Handle bad fits: sd = NA, etc.
  if (!is.null(fit_norm) && any(is.na(fit_norm$estimate))) {
    fit_norm <- NULL
  }

  if (!is.null(fit_skew)) {
    pars <- tryCatch(sn::coef(fit_skew, param.type = "dp"), error = function(e) NULL)
    if (is.null(pars) || any(is.na(pars))) {
      fit_skew <- NULL
    }
  }

  # Choose best available
  if (!is.null(fit_norm) && !is.null(fit_skew)) {
    aic_norm <- AIC(fit_norm)
    aic_skew <- AIC(fit_skew)
    if (aic_skew + 2 < aic_norm) { # skew-normal is more complex (has more parameters) than normal distribution, so only select a skew-normal distribution if it fits substantially better than a normal distribution
      pars <- sn::coef(fit_skew, param.type = "dp")
      return(list(
        type  = "skewnorm",
        xi    = pars["dp.location"],
        omega = pars["dp.scale"],
        alpha = pars["dp.shape"]))
    } else {
      return(list(
        type = "norm",
        mean = fit_norm$estimate["mean"],
        sd   = fit_norm$estimate["sd"]))
    }
  } else if (!is.null(fit_norm)) {
    return(list(
      type = "norm",
      mean = fit_norm$estimate["mean"],
      sd   = fit_norm$estimate["sd"]))
  } else {
    return(list(type = "empirical", empirical = x))
  }
}
```

```{r}
proj_dists_tbl <- all_proj |>
  filter(!is.na(id) & id != "") |>
  group_by(id) |>
  summarise(
    dist_info = list(fit_best(projectedPoints)),
    n_proj = n(), # record of how many sources they have
    .groups = "drop"
  )

proj_dists <- proj_dists_tbl |>
  filter(!is.na(id) & id != "") |>
  distinct(id, .keep_all = TRUE) |>
  (\(x) setNames(x$dist_info, x$id))()
```

```{r}
proj_dists_tbl |>
  dplyr::mutate(
    dist_type = purrr::map_chr(dist_info, ~ .x$type)
  ) |>
  dplyr::count(dist_type)
```

#### `SimDesign` Step 1: Design Grid {#sec-simulationProjectedPointsMonteCarloDesign}

Now we build the `SimDesign` design grid based on the number of projections that each player had.

```{r}
Design <- proj_dists_tbl |>
  transmute(
    id,
    n_sources = n_proj # how many projections the player had
  )
```

```{r}
#| eval: false

missing_ids <- setdiff(Design$id, names(proj_dists))
length(missing_ids) # should be 0

any(is.na(proj_dists_tbl$id)) # should be FALSE
any(is.na(Design$id))         # should be FALSE
any(is.na(names(proj_dists))) # should be FALSE
```

#### `SimDesign` Step 2: Generate {#sec-simulationProjectedPointsMonteCarloGenerate}

```{r}
Generate <- function(condition, fixed_objects = NULL) {

  dist_info <- fixed_objects$proj_dists[[as.character(condition$id)]]
  n_sources <- condition$n_sources

  sim_points <- switch(
    dist_info$type,
    empirical = sample(
      dist_info$empirical,
      n_sources,
      replace = TRUE),

    norm = rnorm(
      n_sources,
      mean = dist_info$mean,
      sd = dist_info$sd),

    skewnorm = sn::rsn(
      n_sources,
      xi = dist_info$xi,
      omega = dist_info$omega,
      alpha = dist_info$alpha),

    stop("Unknown distribution type: ", dist_info$type)
  )

  data.frame(
    id = condition$id,
    sim_points = sim_points)
}
```

#### `SimDesign` Step 3: Analyze {#sec-simulationProjectedPointsMonteCarloAnalyze}

```{r}
Analyse <- function(condition, dat, fixed_objects = NULL) {
  tibble::tibble(
    id        = condition$id,
    mean_pts  = mean(dat$sim_points, na.rm = TRUE),
    sd_pts    = sd(dat$sim_points, na.rm = TRUE),
    q10       = quantile(dat$sim_points, 0.10, na.rm = TRUE),
    q90       = quantile(dat$sim_points, 0.90, na.rm = TRUE),
    p100      = mean(dat$sim_points >= 100, na.rm = TRUE),
    p150      = mean(dat$sim_points >= 150, na.rm = TRUE),
    p200      = mean(dat$sim_points >= 200, na.rm = TRUE),
    p250      = mean(dat$sim_points >= 250, na.rm = TRUE),
    p300      = mean(dat$sim_points >= 300, na.rm = TRUE),
    p350      = mean(dat$sim_points >= 300, na.rm = TRUE)
  )
}
```

#### `SimDesign` Step 4: Summarize {#sec-simulationProjectedPointsMonteCarloSummarize}

```{r}
Summarise <- function(condition, results, fixed_objects = NULL) {
  dplyr::summarise(
    results,
    across(
      where(is.numeric),
      list(
        mean = ~mean(.x, na.rm = TRUE),
        sd   = ~sd(.x,  na.rm = TRUE)),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )
}
```

#### `SimDesign` Step 5: Run the Simulation {#sec-simulationProjectedPointsMonteCarloRun}

Now, we can run the model.

::: {#nte-monteCarloSimulation .callout-note title="Monte Carlo Simulation"}
Note: the following code that runs the simulation takes a while.
If you just want to save time and load the results object instead of running the simulation, you can load the results object of the simulation (which has already been run) using this code:

```{r}
load(url("https://osf.io/download/ues7n/"))
```
:::

```{r}
#| eval: false

monteCarloSim_results <- SimDesign::runSimulation(
  design = Design,
  replications = 1000,
  generate = Generate,
  analyse = Analyse,
  summarise = Summarise,
  fixed_objects = list(proj_dists = proj_dists),
  seed = genSeeds(Design, iseed = 52242), # for reproducibility
  parallel = TRUE # for faster (parallel) processing
)
```

#### Simulation Results {#sec-simulationProjectedPointsMonteCarloResults}

```{r}
monteCarloSim_results <- monteCarloSim_results %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","position","team")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, everything()) %>% 
  arrange(-mean_pts_mean)
```

```{r}
#| eval: false
#| include: false

save(
  monteCarloSim_results,
  file = "./data/monteCarloSim_results.RData")
```

```{r}
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/monteCarloSim_results.RData", fsep = ""))
```

```{r}
monteCarloSim_results
```

```{r}
monteCarloSim_results %>% 
  filter(position == "QB")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "RB")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "WR")
```

```{r}
monteCarloSim_results %>% 
  filter(position == "TE")
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("K","PK"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("DL","DT","DE"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("LB","MLB","OLB"))
```

```{r}
monteCarloSim_results %>% 
  filter(position %in% c("DB","S","CB"))
```

## Conclusion {#sec-simulationConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-simulationSessionInfo}

```{r}
sessionInfo()
```

:::