### 2025-07-02 machine-learning.qmd
20250702 - describe model types

# Machine Learning {#sec-machineLearning}

## Getting Started {#sec-machineLearningGettingStarted}

### Load Packages {#sec-machineLearningLoadPackages}

```{r}
library("petersenlab")
library("parallel")
library("future")
library("missRanger")
library("powerjoin")
library("tidymodels")
library("LongituRF")
library("gpboost")
library("effectsize")
library("tidyverse")
```

### Load Data {#sec-machineLearningLoadData}

```{r}
#| eval: false
#| include: false

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
#load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/nfl_pbp.RData")
load(file = "./data/nfl_4thdown.RData")
load(file = "./data/nfl_participation.RData")
#load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

```{r}
#| label: free-up-memory1-ml
#| include: false

gc()
```

### Specify Options {#machineLearningSpecifyOptions}

```{r}
options(scipen = 999) # prevent scientific notation
```

## Overview of Machine Learning {#sec-machineLearningOverview}

Machine learning takes us away from focusing on [causal inference](#sec-causalInference).
Machine learning does not care about which processes are causal—i.e., which processes influence the outcome.
Instead, machine learning cares about prediction—it cares about a predictor variable to the extent that it increases predictive accuracy regardless of whether it is causally related to the outcome.

Machine learning can be useful for leveraging big data and lots of predictor variable to develop predictive models with greater accuracy.
However, many machine learning techniques are black boxes—it is often unclear how or why certain predictions are made, which can make it difficult to interpret the model's decisions and understand the underlying relationships between variables.
Machine learning tends to be a data-driven, atheoretical technique.
This can result in overfitting.
Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in cross-validation to evaluate the extent of shrinkage of model coefficients.
The data that the model is trained on is known as the "training data".
The data that the model was not trained on but is then is independently tested on—i.e., the hold-out sample—is the "test data".
Shrinkage occurs when predictor variables explain some random error variance in the original model.
When the model is applied to an independent sample (i.e., the test data), the predictive model will likely not perform quite as well, and the regressions coefficients will tend to get smaller (i.e., shrink).

If the test data were collected as part of the same processes as the original data and were merely held out for purposes of analysis, this is called internal cross-validation.
If the test data were collected separately from the original data used to train the model, this is called external cross-validation.

Most machine learning methods were developed with cross-sectional data in mind.
That is, they assume that each person has only one observation on the outcome variable.
However, with longitudinal data, each person has multiple observations on the outcome variable.

When performing machine learning, various approaches may help address this:

- transform data from [long to wide](#sec-longToWide) form, so that each person has only one row
- when designing the training and test sets, keep all measurements from the same person in the same data object (either the training or test set); do not have some measurements from a given person in the training set and other measurements from the same person in the test set
- use a machine learning approach that accounts for the clustered/nested nature of the data

## Types of Machine Learning {#sec-machineLearningTypes}

There are many approaches to machine learning.
This chapter discusses several key ones:

- supervised learning
    - continuous outcome (i.e., regression)
      - [linear regression](#sec-fittingModels-regression)
      - [lasso regression](#sec-lasso)
      - [ridge regression](#sec-ridgeRegression)
      - [elastic net regression](#sec-elasticNet)
      - [random forest](#sec-randomForest)
    - categorical outcome (i.e., classification)
      - [logistic regression](#sec-multipleRegressionTypes)
      - support vector machine
      - [random forest](#sec-randomForest)
      - boosting
        - [tree boosting](#sec-treeBoosting)
- unsupervised learning
    - [cluster analysis](#sec-clusterAnalysis)
    - [principal component analysis](#sec-pca)
    - [factor analysis](#sec-factorAnalysis)
- semi-supervised learning
- reinforcement learning
    - deep learning
- ensemble

*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions than any one method might be able to achieve on its own.

### Supervised Learning {#sec-machineLearningTypesSupervised}

Supervised learning involves learning from data where the correct classification or outcome is known.
For instance, predicting how many points a player will score is a supervised learning task, because there is a ground truth—the actual number of points scored—that can be used to train and evaluate the model.

Unlike linear and logistic regression, various machine learning techniques can handle multicollinearity, including lasso regression, ridge regression, and elastic net regression.
Least absolute shrinkage and selection operator (lasso) regression performs selection of which predictor variables to keep in the model by shrinking some coefficients to zero, effectively removing them from the model.
Ridge regression shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to yield stable estimates for multiple correlated predictor variables in the context of multicollinearity.
Elastic net involves a combination of lasso and ridge regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero (like lasso, for variable selection), and it shrinks the coefficients of some predictor variables toward zero (like ridge, for handling multicollinearity among correlated predictors).

Unless interactions or nonlinear terms are specified, linear, logistic, lasso, ridge, and elastic net regression assume additive and linear associations between the predictors and outcome.
That is, they do not automatically account for interactions among the predictor variables or for nonlinear associations between the predictor variables and the outcome variable (unless interaction terms or nonlinear transformations are explicitly included).
By contrast, random forests and tree boosting methods automatically account for interactions and nonlinear associations between predictors and the outcome variable.
These models recursively partition the data in ways that capture complex patterns without the need to manually specify interaction or polynomial terms.

### Unsupervised Learning {#sec-machineLearningTypesUnsupervised}

Unsupervised learning involves learning from data without known classifications.
Unsupervised learning is used to discover hidden patterns, groupings, or structures in the data.
For instance, if we want to identify different subtypes of Wide Receivers based on their playing style or performance metrics, or uncover underlying dimensions in a large dataset, we would use an unsupervised learning approach.

We describe [cluster analysis](#sec-clusterAnalysis) in @sec-clusterAnalysis.
We describe [principal component analysis](#sec-pca) in @sec-pca.

### Semi-supervised Learning {#sec-machineLearningTypesSemisupervised}

Semi-supervised learning combines supervised learning and unsupervised learning by training the model on some data for which the classification is known and some data for which the classification is not known.

### Reinforcement Learning {#sec-machineLearningTypesReinforcement}

Reinforcement learning involves an agent learning to make decisions by interacting with the environment.
Through trial and error, the agent receives feedback in the form of rewards or penalties and learns a strategy that maximizes the cumulative reward over time.

## Data Processing {#sec-machineLearningDataProcessing}

```{r}
#| eval: false
#| include: false

varNames <- c(
  names(nfl_players),
  names(nfl_teams),
  names(nfl_rosters),
  names(nfl_rosters_weekly),
  names(nfl_schedules),
  names(nfl_combine),
  names(nfl_draftPicks),
  names(nfl_depthCharts),
  names(nfl_pbp),
  names(nfl_4thdown),
  names(nfl_participation),
  #names(nfl_actualFantasyPoints_player_weekly),
  names(nfl_injuries),
  names(nfl_snapCounts),
  names(nfl_espnQBR_seasonal),
  names(nfl_espnQBR_weekly),
  names(nfl_nextGenStats_weekly),
  names(nfl_advancedStatsPFR_seasonal),
  names(nfl_advancedStatsPFR_weekly),
  names(nfl_playerContracts),
  names(nfl_ftnCharting),
  names(nfl_playerIDs),
  names(nfl_rankings_draft),
  names(nfl_rankings_weekly),
  names(nfl_expectedFantasyPoints_weekly),
  names(nfl_expectedFantasyPoints_pbp)
)

varNames <- unique(varNames)

write.csv(
  varNames,
  file = "./data/varNames.csv",
  row.names = FALSE
)

nfl_players$gsis_id
nfl_rosters$gsis_id
nfl_rosters_weekly$gsis_id
nfl_draftPicks$gsis_id
nfl_depthCharts$gsis_id
nfl_advancedStatsPFR_seasonal$gsis_id

nfl_actualStats_offense_weekly$player_id
nfl_expectedFantasyPoints_weekly$player_id

nfl_rankings$id

nfl_combine$pfr_id
nfl_advancedStatsPFR_seasonal$pfr_id
#nfl_advancedStatsPFR_seasonal$pfr_player_id

nfl_playerIDs$gsis_id
nfl_playerIDs$pfr_id
```

### Prepare Data for Merging {#sec-machineLearningPrepareDataForMerging}

```{r}
#| label: prepare-data-merging-machine-learning

# Prepare data for merging
#-todo: calculate years_of_experience
## Use common name for the same (gsis_id) ID variable

#nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
#  rename(gsis_id = player_id)
#
#nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal %>% 
#  rename(gsis_id = player_id)

player_stats_seasonal_offense <- player_stats_seasonal %>% 
  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  rename(gsis_id = player_id)

player_stats_weekly_offense <- player_stats_weekly %>% 
  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  rename(gsis_id = player_id)

nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
  rename(gsis_id = player_id)

## Rename other variables to ensure common names

## Ensure variables with the same name have the same type
nfl_players <- nfl_players %>% 
  mutate(
    birth_date = as.Date(birth_date),
    jersey_number = as.character(jersey_number),
    gsis_it_id = as.character(gsis_it_id),
    years_of_experience = as.integer(years_of_experience))

player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  mutate(
    birth_date = as.Date(birth_date),
    jersey_number = as.character(jersey_number),
    gsis_it_id = as.character(gsis_it_id))

nfl_rosters <- nfl_rosters %>% 
  mutate(
    draft_number = as.integer(draft_number))

nfl_rosters_weekly <- nfl_rosters_weekly %>% 
  mutate(
    draft_number = as.integer(draft_number))

nfl_depthCharts <- nfl_depthCharts %>% 
  mutate(
    season = as.integer(season))

nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
  mutate(
    season = as.integer(season),
    receptions = as.integer(receptions)) %>% 
  distinct(gsis_id, season, week, .keep_all = TRUE) # drop duplicated rows

## Rename variables
nfl_draftPicks <- nfl_draftPicks %>%
  rename(
    games_career = games,
    pass_completions_career = pass_completions,
    pass_attempts_career = pass_attempts,
    pass_yards_career = pass_yards,
    pass_tds_career = pass_tds,
    pass_ints_career = pass_ints,
    rush_atts_career = rush_atts,
    rush_yards_career = rush_yards,
    rush_tds_career = rush_tds,
    receptions_career = receptions,
    rec_yards_career = rec_yards,
    rec_tds_career = rec_tds,
    def_solo_tackles_career = def_solo_tackles,
    def_ints_career = def_ints,
    def_sacks_career = def_sacks
  )

## Subset variables
nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
  select(gsis_id:position, contains("_exp"), contains("_diff"), contains("_team")) #drop "raw stats" variables (e.g., rec_yards_gained) so they don't get coalesced with actual stats

# Check duplicate ids
player_stats_seasonal_offense %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_advancedStatsPFR_seasonal %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1, !is.na(gsis_id)) %>% 
  select(gsis_id, pfr_id, season, team, everything()) %>% 
  head()
```

```{r}
#| label: free-up-memory2-ml
#| include: false

gc()
```

Identify objects with shared variable names:

```{r}
dplyr::intersect(
  names(nfl_players),
  names(nfl_draftPicks))

length(na.omit(nfl_players$position)) # use by default (more cases)
length(na.omit(nfl_draftPicks$position))

dplyr::intersect(
  names(player_stats_seasonal_offense),
  names(nfl_advancedStatsPFR_seasonal))

length(na.omit(player_stats_seasonal_offense$season)) # use by default (more cases)
length(na.omit(nfl_advancedStatsPFR_seasonal$season))

length(na.omit(player_stats_seasonal_offense$team)) # use by default (more cases)
length(na.omit(nfl_advancedStatsPFR_seasonal$team))

length(na.omit(player_stats_seasonal_offense$age)) # use by default (more cases)
length(na.omit(nfl_advancedStatsPFR_seasonal$age))

dplyr::intersect(
  names(nfl_rosters_weekly),
  names(nfl_expectedFantasyPoints_weekly))

length(na.omit(nfl_rosters_weekly$season)) # use by default (more cases)
length(na.omit(nfl_expectedFantasyPoints_weekly$season))

length(na.omit(nfl_rosters_weekly$week)) # use by default (more cases)
length(na.omit(nfl_expectedFantasyPoints_weekly$week))

length(na.omit(nfl_rosters_weekly$position)) # use by default (more cases)
length(na.omit(nfl_expectedFantasyPoints_weekly$position))

length(na.omit(nfl_rosters_weekly$full_name)) # use by default (more cases)
length(na.omit(nfl_expectedFantasyPoints_weekly$full_name))
```

### Merge Data {#sec-machineLearningMergeData}

To merge data, we use the `powerjoin` package [@R-powerjoin]:

```{r}
#| label: merge-data1-machine-learning

# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
#-todo: remove redundant variables
playerListToMerge <- list(
  nfl_players %>% filter(!is.na(gsis_id)),
  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
)

playerSeasonListToMerge <- list(
  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season)),
  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season))
)

playerSeasonWeekListToMerge <- list(
  nfl_rosters_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
  #nfl_actualStats_offense_weekly,
  nfl_expectedFantasyPoints_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week))
  #nfl_advancedStatsPFR_weekly,
)

playerSeasonWeekPositionListToMerge <- list(
  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week))
)

# Merge data
playerMerged <- playerListToMerge %>% 
  reduce(
    powerjoin::power_full_join,
    by = c("gsis_id"),
    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2

playerSeasonMerged <- playerSeasonListToMerge %>% 
  reduce(
    powerjoin::power_full_join,
    by = c("gsis_id","season"),
    conflict = coalesce_xy) # where the objects have the same variable name (e.g., team), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2

playerSeasonWeekMerged <- playerSeasonWeekListToMerge %>% 
  reduce(
    powerjoin::power_full_join,
    by = c("gsis_id","season","week"),
    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
```

```{r}
#| label: free-up-memory3-ml
#| include: false

gc()
```

Identify objects with shared variable names:

```{r}
dplyr::intersect(
  names(playerSeasonMerged),
  names(playerMerged))
```

```{r}
#| label: merge-data2-machine-learning

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged %>% select(-age, -years_of_experience, -team, -team_abbr, -team_seq, -current_team_id), # drop variables from id objects that change from year to year (and thus are not necessarily accurate for a given season)
  by = "gsis_id",
  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
) %>% 
  filter(!is.na(season)) %>% 
  select(gsis_id, season, player_display_name, position, team, games, everything())
```

```{r}
#| label: free-up-memory4-ml
#| include: false

gc()
```

```{r}
dplyr::intersect(
  names(playerSeasonWeekMerged),
  names(seasonalData))
```

```{r}
#| label: merge-data3-machine-learning

seasonalAndWeeklyData <- powerjoin::power_full_join(
  playerSeasonWeekMerged,
  seasonalData,
  by = c("gsis_id","season"),
  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
) %>% 
  filter(!is.na(week)) %>% 
  select(gsis_id, season, week, full_name, position, team, everything())
```

```{r}
#| label: free-up-memory5-ml
#| include: false

gc()
```

```{r}
# Duplicate cases
seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()

seasonalAndWeeklyData %>% 
  group_by(gsis_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

### Additional Processing {#sec-mlAdditionalProcessing}

```{r}
#| label: additional-processing-machine-learning

# Convert character and logical variables to factors
seasonalData <- seasonalData %>% 
  mutate(
    across(
      where(is.character),
      as.factor
    ),
    across(
      where(is.logical),
      as.factor
    )
  )
```

```{r}
#| label: free-up-memory6-ml
#| include: false

gc()
```

### Fill in Missing Data for Static Variables {#sec-fillMissingData}

```{r}
#| label: fill-in-missing-data-machine-learning

seasonalData <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  fill(
    player_name, player_display_name, pos, position, position_group,
    .direction = "downup") %>% 
  ungroup()
```

```{r}
#| label: free-up-memory7-ml
#| include: false

gc()
```

### Create New Data Object for Merging with Later Predictions {#sec-newDataObject}

```{r}
#| label: create-new-data-machine-learning

newData_seasonal <- seasonalData %>% 
  filter(season == max(season, na.rm = TRUE))
```

### Lag Fantasy Points {#sec-lagFantasyPoints}

```{r}
#| label: lag-fantasy-points-machine-learning

seasonalData_lag <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasyPoints)
  ) %>% 
  ungroup()

seasonalData_lag %>% 
  select(gsis_id, player_display_name, season, fantasyPoints, fantasyPoints_lag) # verify that lagging worked as expected
```

```{r}
#| label: free-up-memory8-ml
#| include: false

gc()
```

### Subset to Predictor Variables and Outcome Variable {#sec-subsetToPredictorsAndOutcome}

```{r}
#| label: subset-data-machine-learning

seasonalData_lag %>% select_if(~class(.) == "Date")
seasonalData_lag %>% select_if(is.character)
seasonalData_lag %>% select_if(is.factor)
seasonalData_lag %>% select_if(is.logical)

dropVars <- c(
  "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
  "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
  "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
  "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
  "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
  "football_name", "team")

seasonalData_lag_subset <- seasonalData_lag %>% 
  dplyr::select(-any_of(dropVars))
```

```{r}
#| label: free-up-memory9-ml
#| include: false

gc()
```

### Separate by Position {#sec-separateByPosition}

```{r}
#| label: separate-by-position-machine-learning

seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
  filter(position == "QB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    completions:rushing_2pt_conversions, special_teams_tds, contains(".pass"), contains(".rush"))

seasonalData_lag_subsetRB <- seasonalData_lag_subset %>% 
  filter(position == "RB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetWR <- seasonalData_lag_subset %>% 
  filter(position == "WR") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetTE <- seasonalData_lag_subset %>% 
  filter(position == "TE") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))
```

```{r}
#| label: free-up-memory10-ml
#| include: false

gc()
```

### Split into Test and Training Data {#sec-splitTestTraining}

```{r}
#| eval: false
#| include: false

# Notes to self:
# - Using only retired players as the holdout sample, so that we can include all current/active players in the training model

# CURRENTLY (WILL CHANGE):
# 
# - seasonalData: 1999-2024: i.e. all data currently available
# - seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
# - newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)

# to create:
# 
# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players; 2023 rather than 2024 because don't have fantasy points for retired players in most recent year
# - (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2024, thus predicting fantasy points in 2025, and removing the fantasy_points_lag variable)

# To impute:
# 
# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players

# IMPUTATION AND MODEL PROCESS:
# 
# IMPUTATION:
# - training data
# - test data
# - all data (training and test data), used for generating next year predictions
# 
# MODEL:
# - training data (imputed version of seasonalData_lag_all): all seasons (except 2024 predicting 2025), most players
# 
# EVALUATE MODEL PREDICTIONS:
# - test data (imputed version of seasonalData_lag_test): all seasons (except 2024 predicting 2025), (some of the) retired players
# 
# GENERATE MODEL PREDICTIONS
# - next year data (newData_seasonal): 2024 (predicting 2025)
```

```{r}
#| label: split-test-training-machine-learning

seasonalData_lag_qb_all <- seasonalData_lag_subsetQB
seasonalData_lag_rb_all <- seasonalData_lag_subsetRB
seasonalData_lag_wr_all <- seasonalData_lag_subsetWR
seasonalData_lag_te_all <- seasonalData_lag_subsetTE

set.seed(52242) # for reproducibility (to keep the same train/holdout players)

activeQBs <- unique(seasonalData_lag_qb_all$gsis_id[which(seasonalData_lag_qb_all$season == max(seasonalData_lag_qb_all$season, na.rm = TRUE))])
retiredQBs <- unique(seasonalData_lag_qb_all$gsis_id[which(seasonalData_lag_qb_all$gsis_id %ni% activeQBs)])
numQBs <- length(unique(seasonalData_lag_qb_all$gsis_id))
qbHoldoutIDs <- sample(retiredQBs, size = ceiling(.2 * numQBs)) # holdout 20% of players

activeRBs <- unique(seasonalData_lag_rb_all$gsis_id[which(seasonalData_lag_rb_all$season == max(seasonalData_lag_rb_all$season, na.rm = TRUE))])
retiredRBs <- unique(seasonalData_lag_rb_all$gsis_id[which(seasonalData_lag_rb_all$gsis_id %ni% activeRBs)])
numRBs <- length(unique(seasonalData_lag_rb_all$gsis_id))
rbHoldoutIDs <- sample(retiredRBs, size = ceiling(.2 * numRBs)) # holdout 20% of players

set.seed(52242) # for reproducibility (to keep the same train/holdout players); added here to prevent a downstream error with predict.missRanger() due to missingness; this suggests that an error can arise from including a player in the holdout sample who has missingness in particular variables; would be good to identify which player(s) in the holdout sample evoke that error to identify the kinds of missingness that yield the error

activeWRs <- unique(seasonalData_lag_wr_all$gsis_id[which(seasonalData_lag_wr_all$season == max(seasonalData_lag_wr_all$season, na.rm = TRUE))])
retiredWRs <- unique(seasonalData_lag_wr_all$gsis_id[which(seasonalData_lag_wr_all$gsis_id %ni% activeWRs)])
numWRs <- length(unique(seasonalData_lag_wr_all$gsis_id))
wrHoldoutIDs <- sample(retiredWRs, size = ceiling(.2 * numWRs)) # holdout 20% of players

activeTEs <- unique(seasonalData_lag_te_all$gsis_id[which(seasonalData_lag_te_all$season == max(seasonalData_lag_te_all$season, na.rm = TRUE))])
retiredTEs <- unique(seasonalData_lag_te_all$gsis_id[which(seasonalData_lag_te_all$gsis_id %ni% activeTEs)])
numTEs <- length(unique(seasonalData_lag_te_all$gsis_id))
teHoldoutIDs <- sample(retiredTEs, size = ceiling(.2 * numTEs)) # holdout 20% of players

seasonalData_lag_qb_train <- seasonalData_lag_qb_all %>% 
  filter(gsis_id %ni% qbHoldoutIDs)
seasonalData_lag_qb_test <- seasonalData_lag_qb_all %>% 
  filter(gsis_id %in% qbHoldoutIDs)

seasonalData_lag_rb_train <- seasonalData_lag_rb_all %>% 
  filter(gsis_id %ni% rbHoldoutIDs)
seasonalData_lag_rb_test <- seasonalData_lag_rb_all %>% 
  filter(gsis_id %in% rbHoldoutIDs)

seasonalData_lag_wr_train <- seasonalData_lag_wr_all %>% 
  filter(gsis_id %ni% wrHoldoutIDs)
seasonalData_lag_wr_test <- seasonalData_lag_wr_all %>% 
  filter(gsis_id %in% wrHoldoutIDs)

seasonalData_lag_te_train <- seasonalData_lag_te_all %>% 
  filter(gsis_id %ni% teHoldoutIDs)
seasonalData_lag_te_test <- seasonalData_lag_te_all %>% 
  filter(gsis_id %in% teHoldoutIDs)
```

```{r}
#| label: free-up-memory11-ml
#| include: false

rm(nfl_players)
rm(nfl_teams)
rm(nfl_rosters)
rm(nfl_rosters_weekly)
rm(nfl_schedules)
rm(nfl_combine)
rm(nfl_draftPicks)
rm(nfl_depthCharts)
rm(nfl_pbp)
rm(nfl_4thdown)
rm(nfl_participation)
rm(nfl_actualFantasyPoints_weekly)
rm(nfl_injuries)
rm(nfl_snapCounts)
rm(nfl_espnQBR_seasonal)
rm(nfl_espnQBR_weekly)
rm(nfl_nextGenStats_weekly)
rm(nfl_advancedStatsPFR_seasonal)
rm(nfl_advancedStatsPFR_weekly)
rm(nfl_playerContracts)
rm(nfl_ftnCharting)
#rm(nfl_playerIDs)
rm(nfl_rankings_draft)
rm(nfl_rankings_weekly)
rm(nfl_expectedFantasyPoints_weekly)
rm(nfl_expectedFantasyPoints_pbp)
rm(nfl_actualStats_career)
rm(nfl_actualStats_career_player)
rm(nfl_actualStats_career_player_inclPost)
rm(nfl_actualStats_seasonal)
rm(nfl_actualStats_seasonal_player)
rm(nfl_actualStats_seasonal_player_inclPost)
rm(nfl_actualStats_seasonal_team)
rm(nfl_actualStats_seasonal_team_inclPost)
rm(player_stats_weekly)
rm(player_stats_weekly_offense)
rm(player_stats_seasonal)
rm(player_stats_seasonal_offense)

rm(playerListToMerge)
rm(playerMerged)
rm(playerSeasonListToMerge)
rm(playerSeasonMerged)
rm(playerSeasonWeekListToMerge)
rm(playerSeasonWeekMerged)
rm(playerSeasonWeekPositionListToMerge)
rm(seasonalAndWeeklyData)

gc()
```

### Impute the Missing Data {#sec-missingDataImputation}

```{r}
#| eval: false
#| include: false

# Notes to self

# - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
# - CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION
```

Many of the machine learning approaches described in this chapter require no missing observations in order for a case to be included in the analysis.
In this section, we demonstrate one approach to imputing missing data.
Here is a vignette demonstrating how to impute missing data using `missForest()`: <https://rpubs.com/lmorgan95/MissForest> (archived at: <https://perma.cc/6GB4-2E22>).
Below, we impute the training data (and all data) separately by position.
We then use the imputed training data to make out-of-sample predictions to fill in the missing data for the testing data.
We do not want to impute the training and testing data together so that we can keep them separate for the purposes of cross-validation.
However, we impute all data (training and test data together) for purposes of making out-of-sample predictions from the machine learning models to predict players' performance next season (when actuals are not yet available for evaluating their accuracy).
To impute data, we use the `missRanger` package [@R-missRanger].

::: {#nte-machineLearningImputeMissingData .callout-note title="Impute missing data for machine learning"}
Note: the following code takes a while to run.
:::

```{r}
#| label: impute-qb-machine-learning

# QBs
seasonalData_lag_qb_all_imp <- missRanger::missRanger(
  seasonalData_lag_qb_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_qb_all_imp

data_all_qb <- seasonalData_lag_qb_all_imp$data
data_all_qb$fantasyPointsMC_lag <- scale(data_all_qb$fantasyPoints_lag, scale = FALSE) # mean-centered
data_all_qb_matrix <- data_all_qb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
newData_qb <- data_all_qb %>% 
  filter(season == max(season, na.rm = TRUE)) %>% 
  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
newData_qb_matrix <- data_all_qb_matrix[
  data_all_qb_matrix[, "season"] == max(data_all_qb_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
  , # all columns
  drop = FALSE]

dropCol_qb <- which(colnames(newData_qb_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
newData_qb_matrix <- newData_qb_matrix[, -dropCol_qb, drop = FALSE]

seasonalData_lag_qb_train_imp <- missRanger::missRanger(
  seasonalData_lag_qb_train,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_qb_train_imp

data_train_qb <- seasonalData_lag_qb_train_imp$data
data_train_qb$fantasyPointsMC_lag <- scale(data_train_qb$fantasyPoints_lag, scale = FALSE) # mean-centered
data_train_qb_matrix <- data_train_qb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

seasonalData_lag_qb_test_imp <- predict(
  object = seasonalData_lag_qb_train_imp,
  newdata = seasonalData_lag_qb_test,
  seed = 52242)

data_test_qb <- seasonalData_lag_qb_test_imp
data_test_qb_matrix <- data_test_qb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
#| label: free-up-memory12-ml
#| include: false

gc()
```

```{r}
#| label: impute-rb-machine-learning

# RBs
seasonalData_lag_rb_all_imp <- missRanger::missRanger(
  seasonalData_lag_rb_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_rb_all_imp

data_all_rb <- seasonalData_lag_rb_all_imp$data
data_all_rb$fantasyPointsMC_lag <- scale(data_all_rb$fantasyPoints_lag, scale = FALSE) # mean-centered
data_all_rb_matrix <- data_all_rb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
newData_rb <- data_all_rb %>% 
  filter(season == max(season, na.rm = TRUE)) %>% 
  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
newData_rb_matrix <- data_all_rb_matrix[
  data_all_rb_matrix[, "season"] == max(data_all_rb_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
  , # all columns
  drop = FALSE]

dropCol_rb <- which(colnames(newData_rb_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
newData_rb_matrix <- newData_rb_matrix[, -dropCol_rb, drop = FALSE]

seasonalData_lag_rb_train_imp <- missRanger::missRanger(
  seasonalData_lag_rb_train,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_rb_train_imp

data_train_rb <- seasonalData_lag_rb_train_imp$data
data_train_rb$fantasyPointsMC_lag <- scale(data_train_rb$fantasyPoints_lag, scale = FALSE) # mean-centered
data_train_rb_matrix <- data_train_rb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

seasonalData_lag_rb_test_imp <- predict(
  object = seasonalData_lag_rb_train_imp,
  newdata = seasonalData_lag_rb_test,
  seed = 52242)

data_test_rb <- seasonalData_lag_rb_test_imp
data_test_rb_matrix <- data_test_rb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
#| label: free-up-memory13-ml
#| include: false

gc()
```

```{r}
#| label: impute-wr-machine-learning

# WRs
seasonalData_lag_wr_all_imp <- missRanger::missRanger(
  seasonalData_lag_wr_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_wr_all_imp

data_all_wr <- seasonalData_lag_wr_all_imp$data
data_all_wr$fantasyPointsMC_lag <- scale(data_all_wr$fantasyPoints_lag, scale = FALSE) # mean-centered
data_all_wr_matrix <- data_all_wr %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
newData_wr <- data_all_wr %>% 
  filter(season == max(season, na.rm = TRUE)) %>% 
  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
newData_wr_matrix <- data_all_wr_matrix[
  data_all_wr_matrix[, "season"] == max(data_all_wr_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
  , # all columns
  drop = FALSE]

dropCol_wr <- which(colnames(newData_wr_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
newData_wr_matrix <- newData_wr_matrix[, -dropCol_wr, drop = FALSE]

seasonalData_lag_wr_train_imp <- missRanger::missRanger(
  seasonalData_lag_wr_train,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_wr_train_imp

data_train_wr <- seasonalData_lag_wr_train_imp$data
data_train_wr$fantasyPointsMC_lag <- scale(data_train_wr$fantasyPoints_lag, scale = FALSE) # mean-centered
data_train_wr_matrix <- data_train_wr %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

seasonalData_lag_wr_test_imp <- predict(
  object = seasonalData_lag_wr_train_imp,
  newdata = seasonalData_lag_wr_test,
  seed = 52242)

data_test_wr <- seasonalData_lag_wr_test_imp
data_test_wr_matrix <- data_test_wr %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
#| label: free-up-memory14-ml
#| include: false

gc()
```

```{r}
#| label: impute-te-machine-learning

# TEs
seasonalData_lag_te_all_imp <- missRanger::missRanger(
  seasonalData_lag_te_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_te_all_imp

data_all_te <- seasonalData_lag_te_all_imp$data
data_all_te$fantasyPointsMC_lag <- scale(data_all_te$fantasyPoints_lag, scale = FALSE) # mean-centered
data_all_te_matrix <- data_all_te %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
newData_te <- data_all_te %>% 
  filter(season == max(season, na.rm = TRUE)) %>% 
  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
newData_te_matrix <- data_all_te_matrix[
  data_all_te_matrix[, "season"] == max(data_all_te_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
  , # all columns
  drop = FALSE]

dropCol_te <- which(colnames(newData_te_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
newData_te_matrix <- newData_te_matrix[, -dropCol_te, drop = FALSE]

seasonalData_lag_te_train_imp <- missRanger::missRanger(
  seasonalData_lag_te_train,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

seasonalData_lag_te_train_imp

data_train_te <- seasonalData_lag_te_train_imp$data
data_train_te$fantasyPointsMC_lag <- scale(data_train_te$fantasyPoints_lag, scale = FALSE) # mean-centered
data_train_te_matrix <- data_train_te %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

seasonalData_lag_te_test_imp <- predict(
  object = seasonalData_lag_te_train_imp,
  newdata = seasonalData_lag_te_test,
  seed = 52242)

data_test_te <- seasonalData_lag_te_test_imp
data_test_te_matrix <- data_test_te %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
#| label: free-up-memory15-ml
#| include: false

gc()
```

## Identify Cores for Parallel Processing {#sec-coresParallel}

```{r}
#| label: parallel-cores-machine-learning
#| include: false

num_cores <- parallel::detectCores()
num_true_cores <- parallel::detectCores(logical = FALSE)
```

```{r}
#| eval: false

num_cores <- parallel::detectCores() - 1
num_true_cores <- parallel::detectCores(logical = FALSE) - 1
```

```{r}
num_cores
```

We use the `future` package [@R-future] for parallel (faster) processing.

```{r}
future::plan(future::multisession, workers = num_cores)
```

## Set up the Cross-Validation Folds {#sec-folds}

### *k*-Fold Cross-Validation {#sec-kFoldCrossValidation}

*k*-fold cross-validation partitions the data into *k* folds (subsets).
In each of the *k* iterations, the model is trained on $k - 1$ folds and is evaluated on the remaining fold.
For example, in a 10-fold cross-validation (i.e., $k = 10$), as used below, the model is trained 10 times, each time leaving out a different 10% of the data for validation.
*k*-fold cross-validation is widely used because it tends to yield stable estimates of model performance, by balancing bias and variance.
It is also computationally efficient, requiring only *k* model fits to evaluate model performance.

```{r}
set.seed(52242) # for reproducibility

folds_kFold <- rsample::group_vfold_cv(
  data_train_qb,
  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
  v = 10) # 10-fold cross-validation
```

### Leave-One-Out (LOO) Cross-Validation {#sec-looCrossValidation}

Leave-one-out (LOO) cross-validation partitions the data into *n* folds, where *n* is the sample size.
In each of the *n* iterations, the model is trained on $n - 1$ observations and is evaluated on the one left out.
For example, in a LOO cross-validation with 100 players, the model is trained 100 times, each time leaving out a different player for validation.
LOO cross-validation is especially useful when the dataset is small—too small to form reliable training sets in [*k*-fold cross-validation](#sec-kFoldCrossValidation) (e.g., with $k = 5$ or $k = 10$, which divide the sample into 5 or 10 folds, respectively).
However, LOO tends to be less computationally efficient because it requires more model fits than [*k*-fold cross-validation](#sec-kFoldCrossValidation).
LOO tends to have low bias, producing performance estimates closer to those obtained when fitting the model to the full dataset, because each model is trained on nearly all the data.
However, LOO also tends to have high variance in its error estimates, because each validation fold contains only a single observation, making those estimates more sensitive to individual data points.

```{r}
set.seed(52242) # for reproducibility

folds_loo <- rsample::loo_cv(data_train_qb)
```

## Fitting the Traditional Linear Regression Models {#sec-fittingModels-regression}

We describe linear regression in @sec-multipleRegression.

### Regression with One Predictor {#sec-regressionOnePredictor}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ageCentered20,
  data = data_train_qb)

# Define Model
lm_spec <- parsnip::linear_reg() %>%
  parsnip::set_engine("lm") %>%
  parsnip::set_mode("regression")

# Workflow
lm_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(lm_spec)

# Fit Model with Cross-Validation
cv_results <- tune::fit_resamples(
  lm_wf,
  resamples = folds,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_resamples(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  lm_wf,
  data = data_train_qb)

# View Coefficients
final_model %>% 
  workflows::extract_fit_parsnip() %>% 
  broom::tidy()

final_model %>%
  workflows::extract_fit_parsnip() %>% 
  effectsize::standardize_parameters()

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-regressionOnePredictor
#| fig-cap: "Predicted Versus Actual Fantasy Points for Regression Model with One Predictor (Player Age)."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Regression Model with One Predictor (Player Age)."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

### Regression with Multiple Predictors {#sec-regressionMultiplePredictors}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
lm_spec <- parsnip::linear_reg() %>%
  parsnip::set_engine("lm") %>%
  parsnip::set_mode("regression")

# Workflow
lm_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(lm_spec)

# Fit Model with Cross-Validation
cv_results <- tune::fit_resamples(
  lm_wf,
  resamples = folds,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_resamples(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  lm_wf,
  data = data_train_qb)

# View Coefficients
final_model %>% 
  workflows::extract_fit_parsnip() %>% 
  broom::tidy()

final_model %>%
  workflows::extract_fit_parsnip() %>% 
  effectsize::standardize_parameters()

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-regressionMultiplePredictors
#| fig-cap: "Predicted Versus Actual Fantasy Points for Regression Model with Multiple Predictors."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Regression Model with Multiple Predictors."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

## Fitting the Machine Learning Models {#sec-fittingModels-machineLearning}

### Least Absolute Shrinkage and Selection Option (Lasso) {#sec-lasso}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
lasso_spec <- 
  parsnip::linear_reg(
    penalty = tune::tune(),
    mixture = 1) %>%
  parsnip::set_engine("glmnet")

# Workflow
lasso_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(lasso_spec)

# Define grid of penalties to try (log scale is typical)
penalty_grid <- dials::grid_regular(
  dials::penalty(range = c(-4, -1)),
  levels = 20)

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  lasso_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  lasso_wf,
  best_penalty)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  final_wf,
  data = data_train_qb)

# View Coefficients
final_model %>% 
  workflows::extract_fit_parsnip() %>% 
  broom::tidy()

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-lassoModel
#| fig-cap: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (Lasso) Model."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (Lasso) Model."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

### Ridge Regression {#sec-ridgeRegression}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
ridge_spec <- 
  parsnip::linear_reg(
    penalty = tune::tune(),
    mixture = 0) %>%
  parsnip::set_engine("glmnet")

# Workflow
ridge_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(ridge_spec)

# Define grid of penalties to try (log scale is typical)
penalty_grid <- dials::grid_regular(
  dials::penalty(range = c(-4, -1)),
  levels = 20)

# Tune the Penalty Parameter
cv_results <- tune::tune_grid(
  ridge_wf,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  ridge_wf,
  best_penalty)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  final_wf,
  data = data_train_qb)

# View Coefficients
final_model %>% 
  workflows::extract_fit_parsnip() %>% 
  broom::tidy()

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-ridgeRegressionModel
#| fig-cap: "Predicted Versus Actual Fantasy Points for Ridge Regression Model."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Ridge Regression Model."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

### Elastic Net {#sec-elasticNet}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
enet_spec <- 
  parsnip::linear_reg(
    penalty = tune::tune(),
    mixture = tune::tune()) %>%
  parsnip::set_engine("glmnet")

# Workflow
enet_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(enet_spec)

# Define a regular grid for both penalty and mixture
grid_enet <- dials::grid_regular(
  dials::penalty(range = c(-4, -1)),
  dials::mixture(range = c(0, 1)),
  levels = c(20, 5) # 20 penalty values × 5 mixture values
)

# Tune the Grid
cv_results <- tune::tune_grid(
  enet_wf,
  resamples = folds,
  grid = grid_enet,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  enet_wf,
  best_penalty)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  final_wf,
  data = data_train_qb)

# View Coefficients
final_model %>% 
  workflows::extract_fit_parsnip() %>% 
  broom::tidy()

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-elasticNetModel
#| fig-cap: "Predicted Versus Actual Fantasy Points for Elastic Net Model."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Elastic Net Model."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

### Random Forest Machine Learning {#sec-randomForest}

#### Cross-Sectional Data {#sec-randomForestCrossSectionalData}

```{r}
# Set seed for reproducibility
set.seed(52242)

# Set up Cross-Validation
folds <- folds_kFold

# Define Recipe (Formula)
rec <- recipes::recipe(
  fantasyPoints_lag ~ ., # use all predictors
  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))

# Define Model
rf_spec <- 
  parsnip::rand_forest(
    mtry = tune::tune(),
    min_n = tune::tune(),
    trees = 500) %>%
  parsnip::set_mode("regression") %>%
  parsnip::set_engine(
    "ranger",
    importance = "impurity")

# Workflow
rf_wf <- workflows::workflow() %>%
  workflows::add_recipe(rec) %>%
  workflows::add_model(rf_spec)

# Create Grid
n_predictors <- recipes::prep(rec) %>%
  recipes::juice() %>%
  dplyr::select(-fantasyPoints_lag) %>%
  ncol()

# Dynamically define ranges based on data
rf_params <- hardhat::extract_parameter_set_dials(rf_spec) %>%
  dials:::update.parameters(
    mtry = dials::mtry(range = c(1L, n_predictors)),
    min_n = dials::min_n(range = c(2L, 10L))
  )

rf_grid <- dials::grid_random(rf_params, size = 15) #dials::grid_regular(rf_params, levels = 5)

# Tune the Grid
cv_results <- tune::tune_grid(
  rf_wf,
  resamples = folds,
  grid = rf_grid,
  metrics = yardstick::metric_set(rmse, mae, rsq),
  control = tune::control_grid(save_pred = TRUE)
)

# View Cross-Validation metrics
tune::collect_metrics(cv_results)

# Identify best penalty
tune::select_best(cv_results, metric = "rmse")
tune::select_best(cv_results, metric = "mae")
tune::select_best(cv_results, metric = "rsq")

best_penalty <- tune::select_best(cv_results, metric = "mae")

# Finalize Workflow with Best Penalty
final_wf <- tune::finalize_workflow(
  rf_wf,
  best_penalty)

# Fit Final Model on Training Data
final_model <- workflows::fit(
  final_wf,
  data = data_train_qb)

# View Feature Importance
rf_fit <- final_model %>% 
  workflows::extract_fit_parsnip()

rf_fit

ranger_obj <- rf_fit$fit

ranger_obj$variable.importance

# Predict on Test Data
df <- data_test_qb %>%
  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)

# Evaluate Accuracy of Predictions
petersenlab::accuracyOverall(
  predicted = df$pred,
  actual = df$fantasyPoints_lag,
  dropUndefined = TRUE
)
```

```{r}
#| label: fig-randomForestModel
#| fig-cap: "Predicted Versus Actual Fantasy Points for Random Forest Model."
#| fig-alt: "Predicted Versus Actual Fantasy Points for Random Forest Model."

# Calculate combined range for axes
axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)

ggplot(
  df,
  aes(
    x = pred,
    y = fantasyPoints_lag)) +
  geom_point(
    size = 2,
    alpha = 0.6) +
  geom_abline(
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "dashed") +
  coord_equal(
    xlim = axis_limits,
    ylim = axis_limits) +
  labs(
    title = "Predicted vs Actual Fantasy Points (Test Data)",
    x = "Predicted Fantasy Points",
    y = "Actual Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

```{r}
newData_qb %>%
  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
  left_join(
    .,
    nfl_playerIDs %>% select(gsis_id, name),
    by = "gsis_id"
  ) %>% 
  select(name, fantasyPoints_lag) %>% 
  arrange(-fantasyPoints_lag)
```

Now we can stop the parallel backend:

```{r}
future::plan(future::sequential)
```

```{r}
#| include: false

gc()
```

#### Longitudinal Data {#sec-randomForestLongitudinalData}

Approaches to estimating random forest models with longitudinal data are described by @@Hu2023.
Below, we fit longitudinal random forest models using the `MERF()` function of the `LongituRF` package [@R-LongituRF].

```{r}
smerf <- LongituRF::MERF(
  X = data_train_qb %>% dplyr::select(season:att_br.rush) %>% as.matrix(), # predictors of the fixed effects
  Y = data_train_qb[,c("fantasyPoints_lag")] %>% as.matrix(), # outcome variable
  Z = data_train_qb %>% dplyr::mutate(constant = 1) %>% dplyr::select(constant, passing_yards, passing_tds, passing_interceptions, passing_epa, pacr) %>% as.matrix(), # predictors of the random effects
  id = data_train_qb[,c("gsis_id")] %>% as.matrix(), # player ID (for nesting)
  time = data_train_qb[,c("ageCentered20")] %>% as.matrix(), # time variable
  ntree = 500,
  sto = "BM")

smerf$forest # the fitted random forest (obtained at the last iteration)
smerf$random_effects %>% data.frame() # the predicted random effects for each player
smerf$omega %>% data.frame() # the predicted stochastic processes
smerf$OOB # OOB error at each iteration

# Predict on Test Data
predict(
  smerf,
  X = data_test_qb %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
  Z = data_train_qb %>% dplyr::mutate(constant = 1) %>% dplyr::select(constant, passing_yards, passing_tds, passing_interceptions, passing_epa, pacr) %>% as.matrix(),
  id = data_test_qb[,c("gsis_id")] %>% as.matrix(),
  time = data_test_qb[,c("ageCentered20")] %>% as.matrix())
```

```{r}
#| label: fig-randomForestLongitudinalLogLikelihood
#| fig-cap: "Evolution of the Log-Likelihood."
#| fig-alt: "Evolution of the Log-Likelihood."

plot(smerf$Vraisemblance)
```

### Combining Tree-Boosting with Mixed Models {#sec-treeBoosting}

To combine tree-boosting with mixed models, we use the `gpboost` package [@R-gpboost].

Adapted from here:
<https://towardsdatascience.com/mixed-effects-machine-learning-for-longitudinal-panel-data-with-gpboost-part-iii-523bb38effc>

#### Process Data {#sec-treeBoostingProcessData}

If using a gamma distribution, it requires positive-only values:

```{r}
data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
```

#### Specify Predictor Variables {#sec-treeBoostingSpecifyPredictors}

```{r}
#| label: specify-predictors-machine-learning

pred_vars_qb <- data_train_qb_matrix %>% 
  as_tibble() %>% 
  select(-fantasyPoints_lag, -fantasyPointsMC_lag, -ageCentered20, ageCentered20Quadratic) %>% # -gsis_id
  names()

pred_vars_qb_categorical <- "gsis_id" # to specify categorical predictors
```

#### Specify General Model Options {#sec-treeBoostingSpecifyGeneralModelOptions}

```{r}
#| label: specify-general-model-options-machine-learning

model_likelihood <- "gamma" # gaussian
nrounds <- 2000 # maximum number of boosting iterations (i.e., number of trees built sequentially); more rounds = potentially better learning, but also greater risk of overfitting
```

#### Identify Optimal Tuning Parameters {#sec-treeBoostingOptimalTuningParameters}

For identifying the optimal tuning parameters for boosting, we partition the training data into inner training data and validation data.
We randomly split the training data into 80% inner training data and 20% held-out validation data.
We then use the [mean absolute error](#sec-meanAbsoluteError) as our index of prediction accuracy on the held-out validation data.

```{r}
#| label: tuning-parameters-machine-learning

# Partition training data into inner training data and validation data
ntrain_qb <- dim(data_train_qb_matrix)[1]

set.seed(52242)
valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb)) #

folds_qb <- list(valid_tune_idx_qb)

# Specify parameter grid, gp_model, and gpb.Dataset
param_grid_qb <- list(
  "learning_rate" = c(0.2, 0.1, 0.05, 0.01), # the step size used when updating predictions after each boosting round (high values make big updates, which can speed up learning but risk overshooting; low values are usually more accurate but require more rounds)
  "max_depth" = c(3, 5, 7), # maximum depth (levels) of each decision tree; deeper trees capture more complex patterns and interactions but risk overfitting; shallower trees tend to generalize better
  "min_data_in_leaf" = c(10, 50, 100), # minimum number of training examples in a leaf node; higher values = more regularization (simpler trees)
  "lambda_l2" = c(0, 1, 5)) # L2 regularization penalty for large weights in tree splits; adds a "cost" for complexity; helps prevent overfitting by shrinking the contribution of each tree

other_params_qb <- list(
  num_leaves = 2^6) # maximum number of leaves per tree; controls the maximum complexity of each tree (along with max_depth); more leaves = more expressive models, but can overfit if min_data_in_leaf is too small; num_leaves must be consistent with max_depth, because deeper trees naturally support more leaves; max is: 2^n, where n is the largest max_depth

gp_model_qb <- gpboost::GPModel(
  group_data = data_train_qb_matrix[,"gsis_id"],
  likelihood = model_likelihood,
  group_rand_coef_data = cbind(
    data_train_qb_matrix[,"ageCentered20"],
    data_train_qb_matrix[,"ageCentered20Quadratic"]),
  ind_effect_group_rand_coef = c(1,1))

gp_data_qb <- gpboost::gpb.Dataset(
  data = data_train_qb_matrix[,pred_vars_qb],
  categorical_feature = pred_vars_qb_categorical,
  label = data_train_qb_matrix[,"fantasyPoints_lag"]) # could instead use mean-centered variable (fantasyPointsMC_lag) and add mean back afterward

# Find optimal tuning parameters
opt_params_qb <- gpboost::gpb.grid.search.tune.parameters(
  param_grid = param_grid_qb,
  params = other_params_qb,
  num_try_random = NULL,
  folds = folds_qb,
  data = gp_data_qb,
  gp_model = gp_model_qb,
  nrounds = nrounds,
  early_stopping_rounds = 50, # stops training early if the model hasn’t improved on the validation set in 50 rounds; prevents overfitting and saves time
  verbose_eval = 1,
  metric = "mae")

opt_params_qb
```

```{r}
#| label: free-up-memory16-ml
#| include: false

gc()
```

A learning rate of 1 is very high for boosting.
Even if a learning rate of 1 did well in tuning, I use a lower learning rate (0.1) to avoid overfitting.
I also added some light regularization (`lambda_l2`) for better generalization.
I also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
I set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.

#### Specify Model and Tuning Parameters {#sec-treeBoostingSpecifyModel}

```{r}
#| label: model-parameters-machine-learning

gp_model_qb <- gpboost::GPModel(
  group_data = data_train_qb_matrix[,"gsis_id"],
  likelihood = model_likelihood,
  group_rand_coef_data = cbind(
    data_train_qb_matrix[,"ageCentered20"],
    data_train_qb_matrix[,"ageCentered20Quadratic"]),
  ind_effect_group_rand_coef = c(1,1))

gp_data_qb <- gpboost::gpb.Dataset(
  data = data_train_qb_matrix[,pred_vars_qb],
  categorical_feature = pred_vars_qb_categorical,
  label = data_train_qb_matrix[,"fantasyPoints_lag"])

params_qb <- list(
  learning_rate = 0.1,
  max_depth = 5,
  min_data_in_leaf = 10,
  lambda_l2 = 1,
  num_leaves = 2^5,
  num_threads = num_cores)

nrounds_qb <- 123 # identify optimal number of trees through iteration and cross-validation

#gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead")) # to speed up model estimation
```

#### Fit Model {#sec-treeBoostingFitModel}

```{r}
#| label: fit-model-machine-learning

gp_model_fit_qb <- gpboost::gpb.train(
  data = gp_data_qb,
  gp_model = gp_model_qb,
  nrounds = nrounds_qb,
  params = params_qb) # verbose = 0
```

```{r}
#| label: free-up-memory17-ml
#| include: false

gc()
```

#### Model Results {#sec-treeBoostingModelResults}

```{r}
summary(gp_model_qb) # estimated random effects model

gp_model_qb_importance <- gpboost::gpb.importance(gp_model_fit_qb)
gp_model_qb_importance
```

```{r}
#| label: fig-featureImportanceTreeBoosting
#| fig-cap: "Importance of Features (Predictors) in Tree Boosting Machine Learning Model."
#| fig-alt: "Importance of Features (Predictors) in Tree Boosting Machine Learning Model."

gpboost::gpb.plot.importance(gp_model_qb_importance)
```

#### Evaluate Accuracy of Model on Test Data {#sec-treeBoostingModelAccuracy}

```{r}
#| label: evaluate-model-machine-learning

# Test Model on Test Data
pred_test_qb <- predict(
  gp_model_fit_qb,
  data = data_test_qb_matrix[,pred_vars_qb],
  group_data_pred = data_test_qb_matrix[,"gsis_id"],
  group_rand_coef_data_pred = cbind(
    data_test_qb_matrix[,"ageCentered20"],
    data_test_qb_matrix[,"ageCentered20Quadratic"]),
  predict_var = FALSE,
  pred_latent = FALSE)

y_pred_test_qb <- pred_test_qb[["response_mean"]] # if outcome is mean-centered, add mean(data_train_qb_matrix[,"fantasyPoints_lag"])

predictedVsActual <- data.frame(
  predictedPoints = y_pred_test_qb,
  actualPoints = data_test_qb_matrix[,"fantasyPoints_lag"]
)

predictedVsActual

petersenlab::accuracyOverall(
  predicted = predictedVsActual$predictedPoints,
  actual = predictedVsActual$actualPoints,
  dropUndefined = TRUE
)
```

#### Generate Predictions for Next Season {#sec-treeBoostingModelPredictions}

```{r}
#| label: next-season-predictions-machine-learning

# Generate model predictions for next season
pred_nextYear_qb <- predict(
  gp_model_fit_qb,
  data = newData_qb_matrix[,pred_vars_qb],
  group_data_pred = newData_qb_matrix[,"gsis_id"],
  group_rand_coef_data_pred = cbind(
    newData_qb_matrix[,"ageCentered20"],
    newData_qb_matrix[,"ageCentered20Quadratic"]),
  predict_var = FALSE,
  pred_latent = FALSE)

newData_qb$fantasyPoints_lag <- pred_nextYear_qb$response_mean

# Merge with player names
newData_qb <- left_join(
  newData_qb,
  nfl_playerIDs %>% select(gsis_id, name),
  by = "gsis_id"
)

newData_qb %>% 
  arrange(-fantasyPoints_lag) %>% 
  select(name, fantasyPoints_lag, fantasyPoints)
```

## Conclusion {#sec-machineLearningConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-machineLearningSessionInfo}

```{r}
sessionInfo()
```

:::