### 2025-07-08 _chunk-timing.qmd
20250708 - track chunk execution times

```{r}
#| label: setup-timing
#| include: false

# Set knitr options globally for all chunks
knitr::opts_chunk$set(time_it = TRUE)

# Get chapter-specific log file name
current_file <- knitr::current_input()
log_file <- sprintf("chunk-times-%s.txt", tools::file_path_sans_ext(basename(current_file)))

# Clear previous log
if (file.exists(log_file)) file.remove(log_file)

# Create timing hook
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    label <- if (!is.null(options$label)) options$label else "<unnamed>"
    if (before) {
      now <<- Sys.time()
    } else {
      elapsed <- difftime(Sys.time(), now, units = "secs")
      msg <- sprintf("This chunk (`%s`) ran in %.2f seconds.", label, as.numeric(elapsed))

      # Print to GitHub Actions log
      message(msg)

      # Save to chapter-specific log file
      cat(msg, file = log_file, append = TRUE, sep = "\n")
    }
  }
}))
```