### 2025-08-10 getting-started.qmd
20250809 - label functions

{{< include _chunk-timing.qmd >}}

# Getting Started with `R` for Data Analysis {#sec-gettingStarted}

The book uses the software `R` [@R-base] for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
This chapter provides an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.

## Learning Statistics {#sec-learningStats}

Here are resources for learning statistics:

- OpenIntro Statistics: <https://www.openintro.org/book/os/>
- Introductory Statistics: <https://openstax.org/details/books/introductory-statistics-2e>

## Learning `R` {#sec-learningR}

Here are a various resources for learning `R`:

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- `R` for Data Science: <https://r4ds.hadley.nz>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Learning Statistics with `R`: <https://learningstatisticswithr.com>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package [@R-learnr]: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse` [@R-tidyverse; @Wickham2019_packages], which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

## Getting Help with `R` {#sec-gettingHelpR}

If you have `R` questions, you can ask them in a number of places:

- Forums:
    - `Posit`: <https://forum.posit.co>
    - `StackOverflow`: <https://stackoverflow.com/questions/tagged/r>
    - `Reddit`: <https://www.reddit.com/r/rstats/>
- The `R` mailing list: <https://stat.ethz.ch/mailman/listinfo/r-help>

@Salmon2018 provides additional resources and good guidance for getting help with `R`: <https://masalmon.eu/2018/07/22/wheretogethelp/> (archived at <https://perma.cc/4RRE-KL33>).

When posting a question on forums or mailing lists, keep a few things in mind:

- Read the posting guidelines before posting!
- Be respectful of other people and their time.
`R` is free software.
People are offering their free time to help.
They are under no obligation to help you.
If you are disrespectful or act like they owe you anything, you will rub people the wrong way and will be less likely to get help.
- Provide a minimal, reproducible example.
Providing a minimal, reproducible example can be crucial for getting a helpful response.
By going to the trouble of creating a minimal, reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
@StackOverflow2025 offers guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
@StackOverflow2018 offers a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
@Bryan2025 provide `reprex` do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.

## Initial Setup {#sec-initialSetup}

To get started, follow the following steps:

1. Install `R`: <https://cran.r-project.org>
1. Install `RStudio Desktop`: <https://posit.co/download/rstudio-desktop>
1. After installing `RStudio`, open `RStudio` and run the following code in the console to install the `R` packages used in this book (note: this will take a while):
   ```{r}
   #| eval: false

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","grid","patchwork",
     "pROC", "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats",
     "AICcmodavg","rstan","brms","tidybayes","bbmle","fitdistrplus","sn",
     "mclust","magrittr","viridis","viridisLite","msir","plotly","webshot2",
     "quantmod","fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
     "LongituRF","gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
     "mice","miceadds","interactions","robustbase","ordinal","MASS",
     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk",
     "TTR"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

   remotes::install_github("DevPsyLab/petersenlab")
   remotes::install_github("FantasyFootballAnalytics/ffanalytics")
   remotes::install_github("stan-dev/cmdstanr")
   ```

[^petersenlabPackageGitHub]: Although the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] is hosted on CRAN, installing from GitHub will ensure you have the latest version.

::: {#nte-gettingStarted .callout-note title="If you are in Professor Petersen's class"}
If you are in Professor Petersen's class, also perform the following steps:

1. Download and install `git`: <https://git-scm.com/downloads>
1. Set up a free account on [GitHub.com](https://github.com).
1. Download and install `GitHub Desktop`: <https://desktop.github.com>
1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
1. Go to the following GitHub repository: <https://github.com/isaactpetersen/QuartoBlogFantasyFootball> and complete the following steps:
    1. Click "Use this Template" (in the top right of the screen) > "Create a new repository"
    1. Make sure the checkbox is selected for the following option: "Include all branches"
    1. Make sure your Owner account is selected
    1. Specify the repository name to whatever you want, such as `FantasyFootballBlog`
    1. Type a brief description, such as `Files for my fantasy football blog`
    1. Keep the repository public (this is necessary for generating your blog)
    1. Select "Create repository"
1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
    1. Click "Settings" (in the top of the screen)
    1. Click "Actions" (in the left sidebar) > "General"
    1. Make sure the following are selected:
        - "Read and write permissions" (under "Workflow permissions")
        - "Allow GitHub Actions to create and approve pull requests"
        - then click "Save"
    1. Click "Pages" (in the left sidebar)
    1. Make sure the following are selected:
        - "Deploy from a branch" (under "Source")
        - "gh-pages/(root)" (under "Branch")
        - then click "Save"
1. Clone the repository to your local computer by clicking "Code" > "Open with GitHub Desktop", select the folder where you want the repository to be saved on your local computer, and click "Clone"
:::

## General Workflow {#sec-generalWorkflowAnalysis}

Here is a general workflow for statistical analysis:

1. Load the packages you need
1. Load the data files
1. Do any data processing needed to prepare the data files to be merged.
    - For example, make sure each data file has unique values on the keys.
1. Merge the data files
1. Do any data processing needed to prepare the data files for analysis.
For example:
    - calculate/recode variables
    - transform wide to long or long to wide
    - filter to the relevant rows
    - examine descriptive statistics to ensure data are plausible
1. Do the analysis!
    - evaluate analysis assumptions
    - conduct sensitivity analyses to see extent to which findings differ using different subsets of the data, different modeling approaches/assumptions, etc.

## Install Packages {#sec-installingPackages}

You can install `R` packages using the `utils::install.packages()` function with the following syntax:

```{r}
#| eval: false

install.packages("INSERT_PACKAGE_NAME_HERE")
```

For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse]:

```{r}
#| eval: false

install.packages("tidyverse")
```

You can also install multiple packages with one line of code (replacing each package with its respective name):

```{r}
#| eval: false

install.packages(c("PACKAGE1","PACKAGE2","PACKAGE3"))
```

## Load Packages {#sec-loadPackages}

You can load a given library using the `base::library()` function with the following syntax:

```{r}
#| eval: false

library("INSERT_PACKAGE_NAME_HERE")
```

For instance, you can use the following code to load the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] and `tidyverse` [@R-tidyverse] packages:

```{r}
library("petersenlab")
library("tidyverse")
```

## Using Functions and Arguments {#sec-functionsArguments}

A function often takes particular input(s) and produces some form of output.
The name of the function is followed by parentheses; the inputs go in between the parentheses.
The possible inputs that a function can accept are called "arguments".
You can learn about a particular function and its arguments by entering a question mark before the name of the function:

```{r}
#| eval: false

?NAME_OF_FUNCTION()
```

Below, we provide examples for how to learn about and use functions and arguments, by using the `base::seq()` function as an example.
The `base::seq()` function creates a sequence of numbers.
To learn about the `base::seq()` function, which creates a sequence of numbers, you can execute the following command:

```{r}
#| eval: false

?seq()
```

This is what the documentation shows for the `base::seq()` function in the `Usage` section:

```{r}
#| eval: false

seq(
  from = 1,
  to = 1,
  by = ((to - from)/(length.out - 1)),
  length.out = NULL,
  along.with = NULL,
  ...)
```

Based on this information, we know that the `base::seq()` function takes the following arguments:

- `from`
- `to`
- `by`
- `length.out`
- `along.with`
- `...`

The arguments have default values that are used if the user does not specify values for the arguments.
The default values are provided in the `Usage` section and are in @tbl-seqFunction:

| Argument     | Default Value for Argument       |
|:-------------|:---------------------------------|
| `from`       | `1`                              |
| `to`         | `1`                              |
| `by`         | `((to - from)/(length.out - 1))` |
| `length.out` | `NULL`                           |
| `along.with` | `NULL`                           |

: Arguments and defaults for the `seq()` function. Arguments with a default of `NULL` are not used unless a value is provided by the user. {#tbl-seqFunction}

What each argument represents (i.e., the meaning of `from`, `to`, `by`, etc.) is provided in the `Arguments` section of the documentation.
You can specify a function and its arguments either by providing values for each argument in the order indicated by the function, or by naming its arguments.
Naming arguments explicitly (rather than merely relying on order) is considered best practice because it is safer—doing so prevents you from accidentally assigning an input to the wrong argument.

Here is an example of providing values to the arguments in the order indicated by the function, to create a sequence of numbers from 1 to 9:

```{r}
seq(1, 9)
```

Here is an example of providing values to the arguments by naming its arguments:

```{r}
seq(
  from = 1,
  to = 9,
  by = 1)
```

If you provide values to arguments by naming the arguments, you can reorder the arguments and get the same answer:

```{r}
seq(
  by = 1,
  to = 9,
  from = 1)
```

There are various combinations of arguments that one could use to obtain the same result.
For instance, here is code to generate a sequence from 1 to 9 by 2:

```{r}
seq(
  from = 1,
  to = 9,
  by = 2)
```

Or, alternatively, you could specify the length of the desired sequence (5 values):

```{r}
seq(
  from = 1,
  to = 9,
  length.out = 5)
```

If you want to generate a series with decimal values, you could specify a long desired sequence of 81 values:

```{r}
seq(
  from = 1,
  to = 9,
  length.out = 81)
```

This is equivalent to specifying a sequence from 1 to 9 by 0.1:

```{r}
seq(
  from = 1,
  to = 9,
  by = 0.1)
```

Hopefully, that provides an example for how to learn about a particular function, its arguments, and how to use them.

## Create a Vector {#sec-createVector}

A vector is a series of elements that can be numeric or character.
Character elements should be specified in quotes.
A vector has one dimension (length).
To create a vector, use the `base::c()` function to combine elements into a vector ("c" stands for combine).
And, we use the assignment operator (`<-`) to assign the vector to an object named `exampleVector`, so we can access it later.
Anything on the right side of an assignment operator gets assigned to the object name on the left side of the assignment operator.

```{r}
exampleVector <- c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)

exampleVector2 <- c(
  "A","B1","B2","This is a sentence.","This is another sentence.")
```

We can then access the contents of the object by calling its name:

```{r}
exampleVector
```

## Create a Data Frame {#sec-createDF}

A data frame has two dimensions: rows and columns.
Here is an example of creating a data frame using the `base::data.frame()` function, while using the assignment operator (`<-`) to assign the data frame to an object so we can access it later:

```{r}
players <- data.frame(
  ID = 1:12,
  name = c(
    "Ken Cussion",
    "Ben Sacked",
    "Chuck Downfield",
    "Ron Ingback",
    "Rhonda Ball",
    "Hugo Long",
    "Lionel Scrimmage",
    "Drew Blood",
    "Chase Emdown",
    "Justin Time",
    "Spike D'Ball",
    "Isac Ulooz"),
  position = c("QB","QB","QB","RB","RB","WR","WR","WR","WR","TE","TE","LB"),
  age = c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)
  )

fantasyPoints <- data.frame(
  ID = c(2, 7, 13, 14),
  fantasyPoints = c(250, 170, 65, 15)
)

```

We can also create a data frame from all combinations of variables, using the `base::expand.grid()` function:

```{r}
fantasyPoints_weekly <- expand.grid(
  ID = 1:12,
  season = c(2022, 2023),
  week = 1:17
)
```

We can create a vector that is randomly sampled from values using the `base::sample()` function:

```{r}
set.seed(52242)
fantasyPoints_weekly$fantasyPoints <- sample(
  0:35,
  size = nrow(fantasyPoints_weekly),
  replace = TRUE
)
```

## Create a List {#sec-createList}

A list can store multiple data frames in one object using the `base::list()` function:

```{r}
exampleList <- list(players, fantasyPoints, fantasyPoints_weekly)
```

## Load a Data Frame {#sec-gettingStartedLoadData}

Here is how you load a `.RData` file using a *relative path* (i.e., a path relative to the working directory, where the working directory is represented by a period) using the `base::load()` function:

```{r}
load(file = "./data/nfl_players.RData")
```

The direction of the slashes matters—you should use forward slashes (not backslashes)!
To determine where you working directory is, you can use the `base::getwd()` function:

```{r}
getwd()
```

To change your working directory, you can use the `base::setwd()` function:

```{r}
#| eval: false

setwd("C:/Users/myusername/")
```

The following code loads a file from an *absolute path*:

```{r}
#| eval: false

load("C:/Users/myusername/nfl_players.RData")
```

Here is how you load a `.csv` file using the `utils::read.csv()` function:

```{r}
#| eval: false

nfl_players <- read.csv("./data/nfl_players.csv") # relative path
nfl_players <- read.csv("C:/Users/myusername/nfl_players.csv") # absolute path
```

## Save a Data Frame {#sec-gettingStartedSaveData}

Here is how you save a `.RData` file using a *relative path* (i.e., relative to your working directory) using the `base::save()` function:

```{r}
#| eval: false

save(
  nfl_players,
  file = "./data/nfl_players.RData")
```

The direction of the slashes matters—you should use forward slashes (not backslashes)!
The following code saves a file to an *absolute path*:

```{r}
#| eval: false

save(
  nfl_players,
  file = "C:/Users/myusername/nfl_players.RData")
```

Here is how you save a `.csv` file using the `utils::write.csv()` function:

```{r}
#| eval: false

write.csv(
  nfl_players,
  file = "./data/nfl_players.csv") # relative path

write.csv(
  nfl_players,
  file = "C:/Users/myusername/nfl_players.csv") # absolute path
```

## Variable Names {#sec-variableNames}

To see the names of variables in a data frame, use the `base::names()` function:

```{r}
names(nfl_players)

names(players)

names(fantasyPoints)
```

## Logical Operators {#sec-logicalOperators}

Logical Operators evaluate one or more elements for a condition, and return `TRUE`, `FALSE`, or if the element was missing, `NA`.

### Is Equal To: `==` {#sec-logicalOperatorEqual}

```{r}
players$position == "RB"
```

### Is Not Equal To: `!=` {#sec-logicalOperatorNotEqual}

```{r}
players$position != "RB"
```

### Is Greater Than: `>` {#sec-logicalOperatorGreaterThan}

```{r}
players$age > 30
```

### Is Less Than: `<` {#sec-logicalOperatorLessThan}

```{r}
players$age < 30
```

### Is Greater Than or Equal To: `>=` {#sec-logicalOperatorGreaterThanEqual}

```{r}
players$age >= 30
```

### Is Less Than or Equal To: `<=` {#sec-logicalOperatorLessThanEqual}

```{r}
players$age <= 30
```

### Is In a Value of Another Vector: `%in%` {#sec-logicalOperatorInVector}

```{r}
players$position %in% c("RB","WR")
```

### Is Not In a Value of Another Vector: `!(%in%)` {#sec-logicalOperatorNotInVector}

```{r}
!(players$position %in% c("RB","WR"))
```

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, `petersenlab::\`%ni%\``, for the logical operator of determining whether an element is *not* in a value of another vector:

```{r}
players$position %ni% c("RB","WR")
```

### Is Missing: `is.na()` {#sec-logicalOperatorMissing}

`base::is.na()`

```{r}
is.na(players$age)
```

### Is Not Missing: `!is.na()` {#sec-logicalOperatorNotMissing}

```{r}
!is.na(players$age)
```

### And: `&`{#sec-logicalOperatorAnd}

The "and" operator (`&`) is used to string together multiple logical operators that all must evaluate as `TRUE` in order for the combined evaluation to evaluate as `TRUE`; otherwise, the combined evaluation evaluates as `FALSE`.

```{r}
players$position == "WR" & players$age > 26
```

### Or: `|` {#sec-logicalOperatorOr}

The "or" operator (`|`) is used to string together multiple logical operators, any of which must evaluate as `TRUE` in order for the combined evaluation to evaluate as `TRUE`; otherwise, the combined evaluation evaluates as `FALSE`.

```{r}
players$position == "WR" | players$age > 23
```

### which() {#sec-logicalOperatorWhich}

The operator, `base::which()`, can be used to obtain the TRUE indices of an object, which is useful for [subsetting](#sec-subset).

```{r}
which(players$position == "RB")
```

## If...Else Conditions {#sec-ifelse}

We can use the construction, `if()...else if()...else()` if we want to perform conditional operations.
The typical construction of `if()...else if()...else()` operates such that it first checks if the first `if()` condition is true.
If the first `if()` condition is true, it performs the operation specified and terminates the process.
If the first `if()` condition is not true, it checks the `else if()` conditions in order until one of them is true.
There can be multiple `else if()` conditions.
For the first true `else if()` condition, it performs the operation specified and terminates the process.
If none of the `else if()` conditions is true, it performs the operation specified under `else()` and then terminates the process.
The construction, `if()...else if()...else()` can only be used on one value at a time.

```{r}
player_rank <- 15

if(player_rank <= 10){ # check this condition first
  tier <- 1
  print(tier)
} else if(player_rank <= 20){ # if first condition was not met, check this condition next
  tier <- 2
  print(tier)
} else if(player_rank <= 30){ # if first two conditions were not met, check this condition next
  tier <- 3
  print(tier)
} else{ # if all other conditions were not met, then do this
  print("Don't draft this player!")
}
```

To apply conditional operations to a vector, we can use the `base::ifelse()` function.

```{r}
player_rank <- c(1, 10, 20, 40, 100)

tier <- ifelse(
  player_rank <= 10, # check this condition
  1, # assign this value if true
  2) # assign this value if false

tier
```

## Piping {#sec-piping}

In base `R`, if you want to perform multiple operations, it is common to either a) nest the operations, or b) save the object at each step.

Below is an example of nested operations:

```{r}
length(names(nfl_players))
```

Below is an example of saving the intermediate object at each step:

```{r}
variableNames <- names(nfl_players)
variableNames

lengthOfVariableNames <- length(variableNames)
lengthOfVariableNames
```

Code for performing nested operations can be challenging to read.
Saving the intermediate object can be a waste of time to do if you are not interested in the intermediate object, and can take up unnecessary memory and computational resources.
An alternative approach is to use piping.
Piping allows taking the result from one computation and sending it to the next computation, thus allowing a chain of computations without saving the intermediate object at each step.

In base `R`, you can perform piping with the `base::\`|>\`` expression.
In `tidyverse`, you can perform piping with the `magrittr::\`%>%\`` expression.

#### Base `R` {#sec-pipingBaseR}

```{r}
nfl_players |>
  names() |>
  length()
```

#### Tidyverse {#sec-pipingTidyverse}

```{r}
nfl_players %>%
  names() %>%
  length()
```

## Subset {#sec-subset}

To subset a vector, use brackets to specify the elements to keep:

```{r}
#| eval: false

vector[elementsToKeep]
```

To subset a data frame, use brackets to specify the subset of rows and columns to keep, where the value/vector before the comma specifies the rows to keep, and the value/vector after the comma specifies the columns to keep:

```{r}
#| eval: false

dataframe[rowsToKeep, columnsToKeep]
```

You can subset by using any of the following:

- numeric indices of the elements/rows/columns to keep (or drop)
- names of the rows/columns to keep (or drop)
- values of `TRUE` and `FALSE` corresponding to which elements/rows/columns to keep

### One Variable {#sec-subsetOneVariable}

To subset one variable, use the following syntax:

```{r}
players$name
```

or:

```{r}
players[,"name"]
```

### Particular Rows of One Variable {#sec-subsetParticularRowsOneVariable}

To subset one variable, use the following syntax:

```{r}
players$name[which(players$position == "RB")]
```

or:

```{r}
players[which(players$position == "RB"), "name"]
```

### Particular Columns (Variables) {#sec-subsetParticularColumns}

To subset particular columns/variables, use the following syntax:

#### Base `R` {#sec-subsetParticularColumnsBaseR}

```{r}
subsetVars <- c("name","age")

players[,c(2,4)]
players[,c("name","age")]
players[,subsetVars]
```

Or, to drop columns:

```{r}
dropVars <- c("name","age")

players[,-c(2,4)]
players[,!(names(players) %in% c("name","age"))]
players[,!(names(players) %in% dropVars)]
```

#### Tidyverse {#sec-subsetParticularColumnsTidyverse}

In `tidyverse`, you cans select particular columns using the `dplyr::select()` function.

```{r}
players %>%
  dplyr::select(name, age)

players %>%
  dplyr::select(name:age)

players %>%
  dplyr::select(all_of(subsetVars))
```

Or, to drop columns:

```{r}
players %>%
  dplyr::select(-name, -age)

players %>%
  dplyr::select(-c(name:age))

players %>%
  dplyr::select(-all_of(dropVars))
```

### Particular Rows {#sec-subsetParticularRows}

To subset particular rows, use the following syntax:

#### Base `R` {#sec-subsetParticularRowsBaseR}

```{r}
subsetRows <- c(4,5)

players[c(4,5),]
players[subsetRows,]
players[which(players$position == "RB"),]
```

#### Tidyverse {#sec-subsetParticularRowsTidyverse}

In `tidyverse`, you cans select particular columns using the `dplyr::filter()` function.

```{r}
players %>%
  filter(position == "WR")

players %>%
  filter(position == "WR", age <= 26)

players %>%
  filter(position == "WR" | age >= 26)
```

### Particular Rows and Columns {#sec-subsetParticularRowsAndColumns}

To subset particular rows and columns, use the following syntax:

#### Base `R` {#sec-subsetParticularRowsAndColumnsBaseR}

```{r}
players[c(4,5), c(2,4)]
players[subsetRows, subsetVars]
players[which(players$position == "RB"), subsetVars]
```

#### Tidyverse {#sec-subsetParticularRowsAndColumnsTidyverse}

In `tidyverse`, you cans select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.

```{r}
players %>%
  dplyr::filter(position == "RB") %>%
  dplyr::select(all_of(subsetVars))
```

## View Data {#sec-viewData}

### All Data {#sec-viewDataAll}

To view data, use the `utils::View()` function:

```{r}
#| eval: false

View(players)
```

### First 6 Rows/Elements {#sec-viewDataHead}

To view only the first six rows (if a data frame) or elements (if a vector), use the `utils::head()` function:

```{r}
head(nfl_players)
head(nfl_players$display_name)
```

## Sort Data {#sec-sortData}

```{r}
players %>% 
  arrange(position, age) #sort by position (ascending) then by age (ascending)

players %>% 
  arrange(position, -age) #sort by position (ascending) then by age (descending)
```

## Data Characteristics {#sec-dataCharacteristics}

### Data Structure {#sec-dataCharacteristicsStructure}

`utils::str()`

```{r}
str(nfl_players)
```

### Data Dimensions {#sec-dataCharacteristicsDimensions}

Number of rows and columns (`base::dim()`):

```{r}
dim(nfl_players)
```

Number of rows (`base::nrow()`):

```{r}
nrow(nfl_players)
```

Number of columns (`base::ncol()`):

```{r}
ncol(nfl_players)
```

### Number of Elements {#sec-dataCharacteristicsLength}

`base::length()`

```{r}
length(nfl_players$display_name)
```

### Number of Missing Elements {#sec-dataCharacteristicsMissingElements}

Missing elements are stored in R as `NA`.
You can determine how many missing elements there are in a data frame using the following syntax:

```{r}
length(which(is.na(nfl_players)))
```

You can determine how many missing elements there are in a variable using the following syntax:

```{r}
length(nfl_players$college_name[which(is.na(nfl_players$college_name))])
```

### Number of Non-Missing Elements {#sec-dataCharacteristicsNonMissingElements}

You can determine how many non-missing elements there are in a data frame using the following syntax:

```{r}
length(which(!is.na(nfl_players)))
```

You can determine how many rows in the data frame have no missing data using the following syntax and the `stats::na.omit()` function:

```{r}
nrow(na.omit(nfl_players))
```

You can determine how many non-missing elements there are in a vector using the following syntax:

```{r}
length(nfl_players$college_name[which(!is.na(nfl_players$college_name))])
length(na.omit(nfl_players$college_name))
```

## Create New Variables {#sec-createNewVars}

To create a new variable, use the following syntax:

```{r}
players$newVar <- NA
```

Here is an example of creating a new variable:

```{r}
players$newVar <- 1:nrow(players)
```

```{r}
#| include: false

players$newVar <- NULL
```

## Recode Variables {#sec-recodeVars}

Here is an example of recoding a variable:

```{r}
players$oldVar1 <- NA
players$oldVar1[which(players$position == "QB")] <- "quarterback"
players$oldVar1[which(players$position == "RB")] <- "running back"
players$oldVar1[which(players$position == "WR")] <- "wide receiver"
players$oldVar1[which(players$position == "TE")] <- "tight end"

players$oldVar2 <- NA
players$oldVar2[which(players$age < 30)] <- "young"
players$oldVar2[which(players$age >= 30)] <- "old"
```

Recode multiple variables using the `dplyr::mutate()` and `dplyr::case_match()` functions:

```{r}
players %>%
  dplyr::mutate(across(c(
    oldVar1:oldVar2),
    ~ case_match(
      .,
      c("quarterback","old","running back") ~ 0,
      c("wide receiver","tight end","young") ~ 1)))
```

## Rename Variables {#renameVars}

`dplyr::rename()`

```{r}
players <- players %>% 
  dplyr::rename(
    newVar1 = oldVar1,
    newVar2 = oldVar2)
```

Using a vector of variable names with the `dplyr::rename_with()` function:

```{r}
#| eval: false

varNamesFrom <- c("oldVar1","oldVar2")
varNamesTo <- c("newVar1","newVar2")

players <- players %>% 
  dplyr::rename_with(~ varNamesTo, all_of(varNamesFrom))
```

## Convert the Types of Variables {#convertVarTypes}

One variable (`base::factor()`; `base::as.numeric()`; `base::as.integer()`; `base::as.character()`):

```{r}
players$factorVar <- factor(players$ID)
players$numericVar <- as.numeric(players$age)
players$integerVar <- as.integer(players$newVar1)
players$characterVar <- as.character(players$newVar2)
```

Multiple variables:

```{r}
players %>%
  dplyr::mutate(across(c(
    ID,
    age),
    as.numeric))

players %>%
  dplyr::mutate(across(
    age:newVar1,
    as.character))

players %>%
  dplyr::mutate(across(
    where(is.factor),
    as.character))
```

## Merging/Joins {#sec-merging}

### Overview {#sec-mergingOverview}

Merging (also called joining) merges two data objects using a shared set of variables called "keys."
The keys are the variable(s) that are used to align the rows from the two objects.
The data for the given key(s) in the first object get paired with (i.e., get placed in the same row as) the data for that same key in the second object.
In general, each row should have a value on each of the keys; there should be no missingness in the keys.
To merge two objects, the key(s) that will be used to match the records must be present in both objects.
The keys are used to merge the variables in object 1 (`x`) with the variables in object 2 (`y`).
Different merge types select different rows to merge.

For some data objects, you might want to combine information for the same player from multiple data objects.
If each data object is in `player` form (i.e., `player_id` uniquely identifies each row), you might merge by the player's identification number (e.g., `player_id`).
In this case, the key uniquely identifies each row.

However, some data objects have multiple keys.
For instance, in long form data objects, each player may have multiple rows corresponding to multiple seasons.
In this case, the keys may be `player_id` and `season`—that is, the data are in `player`-`season` form.
If object 1 and object 2 are both in `player`-`season` form, we would use `player_id` and `season` as the keys to merge the two objects.
In this case, the keys uniquely identify each row; that is, they account for the levels of nesting.

However, if the data objects are of different form, we would select the keys as the variable(s) that represent the lowest common denominator of variables used to join the data objects that are present in both objects.
For instance, assume that object 1 is in `player`-`season` form.
For object 2, each player has multiple rows corresponding to seasons and games/weeks—in this case, object 2 is in `player`-`season`-`week` form.
Object 1 does not have the `week` variable, so it cannot be used to join the objects.
Thus, we would use `player_id` and `season` as the keys to merge the two objects, because both variables are present in both objects.

It is important not to have rows with duplicate values on the keys.
For instance, if there is more than one row with the same `player_id` in each object (or multiple rows in object 2 with the same combination of `player_id`, `season`, and `week`), then each row with that `player_id` in object 1 gets paired with each row with that `player_id` in object 2.
The many possible combinations can lead to the resulting object greatly expanding in terms of the number of rows.
Thus, you want the keys to uniquely identify each row.
In the example below, `player` is present in each object, so we can merge by `player`; however, each object has multiple rows with the same player.
For example, `mergeExample1A` has three rows for `player` A; `mergeExample1B` has two rows for `player` A.
Thus, when we merge them, the resulting object has many more rows than each respective object (even though neither object has players that the other object does not).
Below, we merge data using the `dplyr::full_join()` function:

```{r}
mergeExample1A <- data.frame(
  player = c("A","A","A","B","B"),
  age = c(20,22,24,26,28)
)

mergeExample1B <- data.frame(
  player = c("A","A","B","B"),
  points = c(10,15,20,25)
)

mergeExample1 <- dplyr::full_join(
  mergeExample1A,
  mergeExample1B,
  by = "player")

mergeExample1

dim(mergeExample1)
```

Note: if the two objects include variables with the same name (apart from the keys), `R` will not know how you want each to appear in the merged object.
So, it will add a suffix (e.g., `.x`, `.y`) to each common variable to indicate which object (i.e., object `x` or object `y`) the variable came from, where object `x` is the first object—i.e., the object to which object `y` (the second object) is merged.
In general, apart from the keys, you should not include variables with the same name in two objects to be merged.
To prevent this, either remove or rename the shared variable in one of the objects, or include the shared variable as a key.
However, as described above, you should include it as a key ***only*** if you want to use its values to align the rows from each object.
Below is an example of merging two objects with the same variable name (i.e., points) that is not used as a key, using the `dplyr::full_join()` function.

```{r}
mergeExample2A <- data.frame(
  player = c("A","B","C","D","E"),
  points = c(20,22,24,26,28)
)

mergeExample2B <- data.frame(
  player = c("A","B","C","F"),
  points = c(10,15,20,25)
)

mergeExample2 <- dplyr::full_join(
  mergeExample2A,
  mergeExample2B,
  by = "player")

mergeExample2
```

When two objects are merged that have different formats, the resulting data object inherits the format of the data object that has more levels of nesting.
For instance, consider that you want to merge two objects, object A and object B.
Object A is in `player` form and object B is in `player`-`season`-`week` form.
When you merge them, the resulting data object will be in `player`-`season`-`week` form.
Below, we merge data using the `dplyr::full_join()` function:

```{r}
mergeExample3A <- data.frame(
  player = c("A","B","C","D","E"),
  weight = c(225,250,275,300,325)
)

mergeExample3B <- data.frame(
  player = c("A","A","A","A","B","B"),
  season = c(2023,2023,2024,2024,2024,2024),
  week = c(1,2,1,2,3,4),
  points = c(10,15,20,25,30,35)
)

mergeExample3 <- dplyr::full_join(
  mergeExample3A,
  mergeExample3B,
  by = "player")

mergeExample3
```

### Data Before Merging {#sec-mergingData}

Here are the data in the `players` object:

```{r}
players

dim(players)
```

The data are structured in ID form.
That is, every row in the dataset is uniquely identified by the variable, `ID`.

Here are the data in the `fantasyPoints` object:

```{r}
fantasyPoints

dim(fantasyPoints)
```

### Types of Joins {#sec-mergeTypes}

#### Visual Overview of Join Types {#sec-mergeTypesVisual}

@fig-joinTypes depicts various types of merges/joins.
Object `x` is the circle labeled as `x`.
Object `y` is the circle labeled as `y`.
The area of overlap in the Venn diagram indicates the rows on the keys that are shared between the two objects (e.g., the same `player_id`, `season`, and `week`).
The non-overlapping area indicates the rows on the keys that are unique to each object.
The shaded blue area indicates which rows (on the keys) are kept in the merged object from each of the two objects, when using each of the merge types.
For instance, a left outer join keeps the shared rows and the rows that are unique to object `x`, but it drops the rows that are unique to object `y`.

::: {#fig-joinTypes}
![](images/joinTypes.png){width=100% fig-alt="Types of Merges/Joins."}

Types of Merges/Joins.
:::

#### Full Outer Join {#sec-fullJoin}

A full outer join includes all rows in `x` **or** `y`.
It returns columns from `x` and `y`.
Here is how to merge two data frames using a full outer join (i.e., "full join"; `dplyr::full_join()`):

```{r}
fullJoinData <- dplyr::full_join(
  players,
  fantasyPoints,
  by = "ID")

fullJoinData

dim(fullJoinData)
```

#### Left Outer Join {#sec-leftJoin}

A left outer join includes all rows in `x`.
It returns columns from `x` and `y`.
Here is how to merge two data frames using a left outer join ("left join"; `dplyr::left_join()`):

```{r}
leftJoinData <- dplyr::left_join(
  players,
  fantasyPoints,
  by = "ID")

leftJoinData

dim(leftJoinData)
```

#### Right Outer Join {#sec-rightJoin}

A right outer join includes all rows in `y`.
It returns columns from `x` and `y`.
Here is how to merge two data frames using a right outer join ("right join"; `dplyr::right_join()`):

```{r}
rightJoinData <- dplyr::right_join(
  players,
  fantasyPoints,
  by = "ID")

rightJoinData

dim(rightJoinData)
```

#### Inner Join {#sec-innerJoin}

An inner join includes all rows that are in **both** `x` **and** `y`.
An inner join will return one row of `x` for each matching row of `y`, and can duplicate values of records on either side (left or right) if `x` and `y` have more than one matching record.
It returns columns from `x` and `y`.
Here is how to merge two data frames using an inner join (`dplyr::inner_join()`):

```{r}
innerJoinData <- dplyr::inner_join(
  players,
  fantasyPoints,
  by = "ID")

innerJoinData

dim(innerJoinData)
```

#### Semi Join {#sec-semiJoin}

A semi join is a filter.
A left semi join returns all rows from `x` **with** a match in `y`.
That is, it filters out records from `x` that are not in `y`.
Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of `x`, and it includes columns from only `x` (not from `y`).
Here is how to merge two data frames using a left semi join (`dplyr::semi_join()`):

```{r}
semiJoinData <- dplyr::semi_join(
  players,
  fantasyPoints,
  by = "ID")

semiJoinData

dim(semiJoinData)
```

#### Anti Join {#sec-antiJoin}

An anti join is a filter.
A left anti join returns all rows from `x` **without** a match in `y`.
That is, it filters out records from `x` that are in `y`.
It returns columns from only `x` (not from `y`).
Here is how to merge two data frames using a left anti join (`dplyr::anti_join()`):

```{r}
antiJoinData <- dplyr::anti_join(
  players,
  fantasyPoints,
  by = "ID")

antiJoinData

dim(antiJoinData)
```

#### Cross Join {#sec-crossJoin}

A cross join combines each row in `x` with each row in `y` (`dplyr::cross_join()`).

```{r}
crossJoinData <- dplyr::cross_join(
  players,
  fantasyPoints)

crossJoinData

dim(crossJoinData)
```

## Long and Wide Form {#sec-longWideForm}

Depending on the analysis, it may be important to restructure the data to be in long or wide form.
When the data are in wide form, each player has only one row.
When the data are in long form, each player has multiple rows—e.g., a row for each game.
The data structure is called wide or long form because a dataset in wide form has more columns and fewer rows (i.e., it appears wider and shorter), whereas a dataset in long form has more rows and fewer columns (i.e., it appears narrower and taller).

Here are the original data in long form.
The data are structured in "`player`-`season`-`week` form".
That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`—these are the keys.
This is an example of long form, because each player has multiple rows.

```{r}
dataLong <- dplyr::full_join(
  players %>% dplyr::select(-age),
  fantasyPoints_weekly,
  by = c("ID")
) %>% 
  arrange(ID, season, week)

dataLong

dim(dataLong)

names(dataLong)
```

### Transform Data from Long to Wide {#sec-longToWide}

Below, we transform the data from long form to wide form.
The long data are in "`player`-`season`-`week` form".
We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:

```{r}
dataWide <- dataLong %>% 
  tidyr::pivot_wider(
    names_from = c(season, week),
    names_glue = "{.value}_{season}_week{week}",
    values_from = fantasyPoints)

dataWide

dim(dataWide)

names(dataWide)
```

### Transform Data from Wide to Long {#sec-wideToLong}

Conversely, we can also restructure data from wide to long using the `tidyr::pivot_longer()` function.
Here are the data in long form, after they have been transformed from wide form using `tidyverse`:

```{r}
dataLong <- dataWide %>% 
  tidyr::pivot_longer(
    cols = fantasyPoints_2022_week1:fantasyPoints_2023_week17,
    names_to = c("season", "week"),
    names_pattern = "fantasyPoints_(.*)_week(.*)",
    values_to = "fantasyPoints")

dataLong

dim(dataLong)

names(dataLong)
```

## Loops {#sec-loops}

If you want to perform the same computation multiple times, it can be faster to do it in a loop compared to writing out the same computation many times.
For instance, here is a loop that runs from 1 to `{r} length(players$ID)` (the number of players in the `players` object), incrementing by 1 after each iteration.
The loop prints each element of a vector (i.e., the player's name) and the loop index (`i`) that indicates where the loop is in terms of its iterations:

```{r}
for(i in 1:length(players$ID)){
  print(paste("The loop is at index:", i, sep = " "))
  print(paste("My favorite player is:", players$name[i], sep = " "))
}
```

Loops are fine for basic computations, but other approaches (such as the `apply()` family of functions) can be even faster than loops.

## Create a Function {#sec-createFunction}

Now, let's put together what we have learned to create a useful function.
Functions are useful if you want to perform an operation multiple times, especially if you want to be able to manipulate various settings to make slight modifications each time.
Any operation that you want to perform multiple times, you can create a function to accomplish.
Use of a function can save you time without needed to retype out all of the code each time.
For instance, let's say you want to convert temperature between Fahrenheit and Celsius, you could create a function to do that.
In this case, our function has two arguments: `temperature` (in degrees) and `unit` of the original temperature (`F` for Fahrenheit or `C` for Celsius, where the default unit is Fahrenheit).
That allows us to make the slight modification of which unit is the input temperature and to change the calculation accordingly.

```{r}
convert_temperature <- function(temperature, unit = "F"){ # specify the arguments and any defaults
  if(unit == "F"){ # if the input temperature(s) in Fahrenheit
    newtemp <- (temperature - 32) / (9/5)
  } else if(unit == "C"){ # if the input temperature(s) in Celsius
    newtemp <- (temperature * (9/5)) + 32
  }

  return(newtemp) # this is what is returned by the function
}
```

Now we can use the function to convert temperatures between Fahrenheit and Celsius.
A temperature of 32°F is equal to 0°C.
A temperature of 0°C is equal to 89.6°F.

```{r}
convert_temperature(
  temperature = 32,
  unit = "F"
)

convert_temperature(
  temperature = 32,
  unit = "C"
)
```

We can also convert the temperature for a vector of values at once:

```{r}
convert_temperature(
  temperature = c(0, 10, 20, 30, 40, 50),
  unit = "F"
)

convert_temperature(
  temperature = c(0, 10, 20, 30, 40, 50),
  unit = "C"
)
```

Because the default unit is "F", we do not need to specify the unit if our input temperatures are in Fahrenheit:

```{r}
convert_temperature(
  c(0, 10, 20, 30, 40, 50)
)
```

## Commenting Code {#sec-comment}

To comment your code, using the `#` sign.
Anything on the line that appears after the `#` sign will be interpreted as a comment and will not be evaluated by R.
It is important to comment your code frequently with what you are doing doing and why so that you and others know how to read it.

```{r}
x <- 1 # this line will run
       # this line will NOT run: x <- 2

x
```

# Conclusion {#sec-gettingStartedConclusion}

This chapter provided an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
Most of the code throughout the book that performs data management operations leverages the techniques implemented in this chapter.

::: {.content-visible when-format="html"}

## Session Info {#sec-gettingStartedSessionInfo}

At the end of each chapter in which `R` code is used, I provide the session information, which describes the system and operating system the code was run on and the versions of each package.
That way, if you get different results from me, you can see which session settings differ, to help with reproducibility.
If you run the (all of) the exact same code as is provided in the text, in the exact same order, with the exact same setup (platform, operating system, package versions, etc.) and the exact same data, you should get the exact same answer as is in the text.
That is the idea of reproducibility—getting the exact same result with the exact same inputs.
Reproducibility is crucial for studies to achieve greater confidence in their findings and to ensure better replicability of findings across studies.

```{r}
sessionInfo()
```

:::