### 2025-09-16 data-visualization.qmd
20250916 - update visualization to 2025

{{< include _chunk-timing.qmd >}}

# Data Visualization {#sec-dataVisualization}

This chapter provides an overview of principles of data visualization and how to implement them using `R`.

## Getting Started {#sec-dataVisualizationGettingStarted}

### Load Packages {#sec-dataVisualizationLoadPackages}

```{r}
library("nflplotR")
library("plotly")
library("gghighlight")
library("ggridges")
library("ggExtra")
library("patchwork")
library("tidyverse")
```

### Load Data {#sec-dataVisualizationLoadData}

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
```

```{r}
load(file = "./data/nfl_pbp.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

## Overview {#sec-dataVisualizationOverview}

Graphics serve an important way of efficiently communicating findings to diverse audiences.
Effective graphics can help people with less technical knowledge better understand your findings.
As noted by @Tanney2021, it is important to communicate your work clearly to stakeholders (e.g., coaches, companies, or prospective clients); if you cannot explain your work clearly, you cannot expect a stakeholder to "buy in" to what you are selling.

### Principles of Graphic Design {#sec-dataVisualizationOverviewGraphicDesign}

When designing graphics, it is important to understand general principles of graphic design.
@AdobeExpress2020 describes important principles about graphic design at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
The important principles include:

> 1. Focus on alignment.
> 2. Use hierarchy to help focus your design.
> 3. Leverage contrast to accentuate important design elements.
> 4. Use repetition to your advantage.
> 5. Consider proximity when organizing your graphic elements.
> 6. Make sure your designs have balance.
> 7. Optimize color to support your design.
> 8. Leave negative space.

### Principles of Data Visualization {#sec-dataVisualizationOverviewPrinciples}

Data visualization involves graphic design in a particular domain—the visualization of data (numeric-derived information).
@Schwabish2021 describes five principles in data visualization:

> 1. Show the data.
> 2. Reduce the clutter.
> 3. Integrate the graphics and text.
> 4. Avoid the spaghetti chart.
> 5. Start with gray.

"Showing the data" involves showing the data that matter the most.
"Reducing the clutter" involves removing non-data things that obscure the data—for example, extraneous gridlines, tick marks, data markers (e.g., symbols to distinguish between series), and complex shadings (e.g., textured or filled gradients).
"Integrating the graphics and text" involves using headline titles, clear and useful labels (instead of legends), and helpful annotations.
Headline or newspaper-like titles are titles that are succinct with active phasing and that indicate the take-away message (e.g., "Quarterbacks Threw Fewer Touchdowns in 2024 than in Previous Years").
In terms of labels, @Schwabish2021 advocates to label the data directly instead of using a legend.
In terms of helpful annotations, you can provide additional text that helps explain the data (e.g., peaks or valleys, outliers, or other variations that deserve explanation), including how to interpret the chart.
"Avoiding the spaghetti chart" means avoiding packed charts with too much information that makes them difficult to interpret.
Spaghetti charts are lines with many lines that, make the plot look like a bunch of spaghetti.
However, @Schwabish2021 also advocates against using charts of other types that are complicated and difficult to interpret due to too much information, such as complicated maps or bar plots with too many colors, icons, or bars.
If there are too many lines or series, @Schwabish2021 advocates [breaking it up into multiple charts](#sec-faceting) (i.e., facets, trellis charts, or small multiples).
An example of [faceted charts](#sec-faceting) is depicted in @fig-facetedScatterplot.
"Starting with gray" refers to the idea of using gray as the default color for most lines/points/bars, so that you can use a color to highlight the lines/points/bars of interest.
In addition, as noted by @Schwabish2021, it is important to treat data as objectively as possible and not to present figures in a biased way as to mislead.

In his classic book, @Tufte2001 states that effective data visualizations should follow principles of graphical excellence and integrity.
He notes that "Graphical excellence is that which gives to the viewer the greatest number of ideas in the shortest time with the least ink in the smallest space." (p. 51).
That is, data visualizations should seek to maximize the data-to-ink ratio (within reason), and should spend less space on "fluff" (i.e., non-data things that can be erased without losing meaning, such as grid lines, redundancies, etc.).
This is consistent with Schwabish's -@Schwabish2021 principles of showing the data and reducing the clutter.
@Tufte2001 describes six principles of graphical integrity:

> 1. The representation of numbers, as physically measured on the surface of the graphic itself, should be directly proportional to the numerical quantities represented.
> 2. Clear, detailed, and thorough labeling should be used to defeat graphic distortion and ambiguity. Write out explanations of the data on the graphic itself. Label important events in the data.
> 3. Show data variation, not design variation.
> 4. In time-series displays of money, deflated and standardized units of monetary measurement are nearly always better than nominal units.
> 5. The number of information-carrying (variable) dimensions depicted should not exceed the number of dimensions in the data.
> 6. Graphics must not quote data out of context.
> 
> --- Tufte [-@Tufte2001, p. 77]

@Tufte2001 also provides recommendations for friendly, accessible graphics, including:

- spell words out (rather than using abbreviations)
- have words run from left to write (including the y-axis title)
- include little messages to help explain the data
- place labels on the graphic so no legend is needed
- avoid elaborately encoded shadings, cross-hatching, and colors
- avoid "chartjunk"—i.e., unnecessary or distracting elements (e.g., excessive decoration, overly complex graphics, graphical effects, and irrelevant information such as moiré vibration, heavy grids, and self-promoting graphs) that do not improve viewers' understanding of the data
- if colors are used, use colors that are distinguishable by color-deficient and color-blind viewers (red–green is a common form of color-blindness)
- use type (i.e., of the text) that is clear, precise, and modest
- use text that is upper-and-lower case, not all capitals

An example figure that applies these principles of data visualization is in @fig-exampleDataVisualization.

```{r}
#| label: fig-exampleDataVisualization
#| fig-cap: "Example Figure that Applies the Principles of Data Visualization."
#| fig-alt: "Example Figure that Applies the Principles of Data Visualization."
#| code-fold: true

confidenceLevel <- .95 # for 95% confidence interval

player_stats_seasonal_offense_summary <- player_stats_seasonal %>%
  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  group_by(position_group) %>%
  summarise(
    n = sum(!is.na(fantasyPoints)),
    mean = mean(fantasyPoints, na.rm = TRUE),
    sd = sd(fantasyPoints, na.rm = TRUE)
  ) %>%
  mutate(se = sd/sqrt(n)) %>%
  mutate(
    ci_lower = mean - qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
    ci_upper = mean + qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
    positionLabel = case_match(
      position_group,
      "QB" ~ "Quarterback",
      "RB" ~ "Running Back",
      "WR" ~ "Wide Receiver",
      "TE" ~ "Tight End"
      )
  )

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = positionLabel,
    y = mean,
    fill = positionLabel
    )
) +
  geom_bar(
    stat = "identity") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  gghighlight::gghighlight(
    positionLabel == "Quarterback",
    label_key = positionLabel) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Quarterbacks Score More Fantasy Points than Other Positions"
  ) +
  annotate(
    "segment",
    x = 3.5,
    xend = 3.2,
    y = 70,
    yend = 35,
    color = "blue",
    linewidth = 1.5,
    alpha = 0.6,
    arrow = arrow()) +
  annotate(
    "text",
    x = 2.75,
    y = 75,
    label = "Tight Ends score fewer fantasy\npoints than other positions",
    hjust = 0) + # left-justify
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Creating Data Visualizations in `R` {#sec-dataVisualizationOverviewInR}

The Data Visualization Catalogue, from Data to Viz, and Data Viz Project provide examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>, <https://www.data-to-viz.com>, and <https://datavizproject.com>.
The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
In this chapter, we will examine how to create statistical graphics to visualize data.
We will create the plots using the `ggplot2` package [@R-ggplot2; @ggplot3rdEd].
When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).

A key principle of graphic design and data visualization is the importance of contrast.
Each visual component (e.g., line) that is important to see should be easy to distinguish.
For instance, you can highlight lines or points of interest to draw people's attention to the target of interest [@Schwabish2021].
For examples of highlighting in figures, see Figures [-@fig-lineChartHighlighting] (@sec-lineChartHighlighting) and [-@fig-simulationOf10CoinFlips].

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by [ColorBrewer](https://colorbrewer2.org) (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
[ColorBrewer](https://colorbrewer2.org) allows users to select colorblind safe, print friendly, and photocopy safe palettes.
Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available from the [`viridis` palette](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale). 
There are a variety of resources for color schemes that are accessible to color-blind viewers:

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> [@Kilin2022; archived at <https://perma.cc/7VTA-Y8YS>]
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- [Okabe-Ito color palette](https://jfly.uni-koeln.de/color/) [@Okabe2008; archived at <https://perma.cc/VD2E-XZZU>]
- @Nunez2018

The Okabe-Ito color palette [@Okabe2008] can be specified with the following colors:

```{r}
palette_OkabeIto <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999")

palette_OkabeIto_black <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#000000")
```

The Okabe-Ito color palette [@Okabe2008] is depicted in Figures [-@fig-okabeItoPaletteGray] and [-@fig-okabeItoPaletteBlack].
You can see a pie chart of colors using the `graphics::pie()` function:

```{r}
#| label: fig-okabeItoPaletteGray
#| fig-cap: "Okabe-Ito Color Palette With Gray."
#| fig-alt: "Okabe-Ito Color Palette With Gray."

pie(
  rep(1, 8),
  col = palette_OkabeIto)
```

```{r}
#| label: fig-okabeItoPaletteBlack
#| fig-cap: "Okabe-Ito Color Palette With Black."
#| fig-alt: "Okabe-Ito Color Palette With Black."

pie(
  rep(1, 8),
  col = palette_OkabeIto_black)
```

## `ggplot2` Syntax {#sec-ggplot2Syntax}

### General Syntax {#sec-ggplot2SyntaxGeneral}

The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system (and aesthetic attributes), and 3) the geometric objects ("geoms")—visual marks that represent data points (e.g., points, lines, and bars).
The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:

```{r}
#| eval: false

ggplot2::ggplot(
  data = INSERT, # 1) data
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS # 3) geom(s)
```

The `ggplot2::coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
In addition, there are optional settings you can specify or modify, such as [axis titles](#sec-scatterplotPlotLabels), [axis labels](#sec-scatterplotAxes), [faceting](#sec-faceting), and the [theme](#sec-scatterplotTheme).

Here is a template including these potential modifications:

```{r}
#| eval: false

ggplot2::ggplot(
  data = INSERT, # 1) data
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS + # 3) geom(s)
  scale_x_continuous(
    breaks = INSERT # specify x-axis labels
  ) +
  scale_y_continuous(
    breaks = INSERT # specify y-axis labels
  ) +
  labs( # add plot labels
    x = "INSERT", # x-axis title
    y = "INSERT", # y-axis title
    title = "INSERT", # plot title
    subtitle = "INSERT", # plot subtitle
    caption = "Add a caption below plot", # plot caption
    alt = "Add alt text to the plot" # plot alt-text
  ) +
  INSERT_THEME_NAME + # apply theme
  facet_wrap(vars(INSERT_VARIABLES_TO_FACET)) # facet by one or more variables
```

You can layer multiple geoms in the same plot, layer by layer.
You can have a different data set and aesthetics associated with each geom, by specifying the `data` and `mapping` arguments for the function of that particular geom.
For an example of a plot that uses different data sets for different geoms, see @fig-bootstrapRegressionCoefficients (@sec-multipleRegressionRobustRegression).

### Geometric Objects {#sec-ggplot2SyntaxGeoms}

Here is a list of common geometric objects ("geoms") available in `ggplot2` (there are many others):

- [histogram](#sec-histogram): `ggplot2::geom_histogram()`
- [density plot](#sec-densityPlot): `ggplot2::geom_density()`
- [box-and-whisker plot](#sec-boxAndWhiskerPlot): `ggplot2::geom_boxplot()`
- [violin plot](#sec-violinPlot): `ggplot2::geom_violin()`
- [bar plot](#sec-barPlot): `ggplot2::geom_bar()`
- [scatterplot](#sec-scatterplot): `ggplot2::geom_point()`
- [rug plot](#sec-scatterplotRugPlot): `ggplot2::geom_rug()`
- [line chart](#sec-lineChart): `ggplot2::geom_line()`
- statistical summary (e.g., [best-fit line](#sec-scatterplotBestFitLine)): `ggplot2::geom_smooth()`

### Color {#sec-ggplot2SyntaxColor}

In `ggplot2`, you can specify the color of a geom using one of two parameters: 1) `fill` and 2) `color`.
The `fill` parameter specifies the color inside the geom.
The `color` parameter specifies the outline/border color of the geom.

The colors to use in the `fill` and `color` parameters can be specified by name (as in @fig-colorByName), hexadecimal (HEX) codes (as in @fig-colorByHEX), values in red-green-blue (RGB) space (as in @fig-colorByRGB), or palettes (e.g., ColorBrewer or viridis; as in Figures [-@fig-multipleColorsByPaletteColorBrewer] and [-@fig-multipleColorsByPaletteViridis]).

```{r}
#| label: fig-colorByName
#| fig-cap: "Specifying the Color/Fill by Name."
#| fig-alt: "Specifying the Color/Fill by Name."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_histogram(
    color = "black",
    fill = "forestgreen"
  )
```

You can see a full list of color names available using the `grDevices::colors()` function:

```{r}
colors()
```

However, you can use any color by specifying its HEX code, as in @fig-colorByHEX.

```{r}
#| label: fig-colorByHEX
#| fig-cap: "Specifying the Color/Fill by Hexadecimal Code."
#| fig-alt: "Specifying the Color/Fill by Hexadecimal Code."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_histogram(
    color = "#000000",
    fill = "#228B22"
  )
```

You can use any color by specifying its values in red-green-blue (RGB) color space using the `grDevices::rgb()` function, as in @fig-colorByRGB.

```{r}
#| label: fig-colorByRGB
#| fig-cap: "Specifying the Color/Fill by Value in Red-Green-Blue (RGB) Space."
#| fig-alt: "Specifying the Color/Fill by Value in Red-Green-Blue (RGB) Space."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints)
) +
  geom_histogram(
    color = rgb(
      red = 0,
      green = 0,
      blue = 0,
      maxColorValue = 255
    ),
    fill = rgb(
      red = 34,
      green = 139,
      blue = 34,
      maxColorValue = 255
    )
  )
```

We can allow the colors to differ by group by specifying the `fill` parameter in the mapping aesthetics, as in @fig-colorByGroup.

```{r}
#| label: fig-colorByGroup
#| fig-cap: "Specifying Color By Group."
#| fig-alt: "Specifying Color By Group."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_density(
    color = "black",
    alpha = 0.7
  )
```

We can specify multiple colors manually using the `ggplot2::scale_fill_manual()` function, as in @fig-multipleColorsManually.

```{r}
#| label: fig-multipleColorsManually
#| fig-cap: "Specifying Multiple Colors Manually."
#| fig-alt: "Specifying Multiple Colors Manually."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_density(
    color = "black",
    alpha = 0.7
  ) +
  scale_fill_manual(
    values = c(
      "QB" = "red", # or could use hex, rgb, etc.
      "RB" = "blue", # or could use hex, rgb, etc.
      "WR" = "green", # or could use hex, rgb, etc.
      "TE" = "yellow") # or could use hex, rgb, etc.
    )
```

We can also specify the colors using a palette, such as by the ColorBrewer palette using the `ggplot2::scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `ggplot2::scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).

```{r}
#| label: fig-multipleColorsByPaletteColorBrewer
#| fig-cap: "Specifying Multiple Colors Using the ColorBrewer Palette."
#| fig-alt: "Specifying Multiple Colors Using the ColorBrewer Palette."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_density(
    color = "black",
    alpha = 0.7
  ) +
  scale_fill_brewer(palette = "Set1") # see ?scale_fill_brewer for list of available palettes
```

```{r}
#| label: fig-multipleColorsByPaletteViridis
#| fig-cap: "Specifying Multiple Colors Using the Viridis Palette."
#| fig-alt: "Specifying Multiple Colors Using the Viridis Palette."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_density(
    color = "black",
    alpha = 0.7
  ) +
  scale_fill_viridis_d(option = "viridis") # see ?scale_fill_viridis_d for list of available palettes
```

### Line Type {#sec-ggplot2SyntaxLineType}

Common line types include "solid" lines, "dashed" lines, and "dotted" lines.
A full list of line types available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>).
To specify the line type, you can use the `linetype` parameter, as in @fig-lineType.

```{r}
#| label: fig-lineType
#| fig-cap: "Specifying Line Type."
#| fig-alt: "Specifying Line Type."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name)
) +
  geom_line(
    linetype = "dashed"
  )
```

We can allow the line type to differ by group by specifying the `linetype` parameter in the mapping aesthetics, as in @fig-lineTypeByGroup.

```{r}
#| label: fig-lineTypeByGroup
#| fig-cap: "Specifying Line Type By Group."
#| fig-alt: "Specifying Line Type By Group."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    linetype = player_display_name)
) +
  geom_line()
```

We can specify the line types manually using the `ggplot2::scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.

```{r}
#| label: fig-lineTypeByGroupManually
#| fig-cap: "Specifying Line Type By Group Manually."
#| fig-alt: "Specifying Line Type By Group Manually."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name,
    linetype = player_display_name)
) +
  geom_line() +
  scale_linetype_manual(
    values = c(
      "Tom Brady" = "solid",
      "Peyton Manning" = "dashed")
    )
```

### Line Width {#sec-ggplot2SyntaxLineWidth}

To specify the line width, you can use the `linewidth` parameter, as in @fig-linewidth.

```{r}
#| label: fig-linewidth
#| fig-cap: "Specifying Line Width."
#| fig-alt: "Specifying Line Width."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name)
) +
  geom_line(
    linewidth = 4
  )
```

### Point Shape {#sec-ggplot2SyntaxPointShape}

Common point shapes include circles, triangles, squares, etc.
A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>) and is in @fig-pointShapeOptions.

::: {#fig-pointShapeOptions}
![](images/pointShapes.png){width=100% fig-alt="Available Point Shapes in ggplot2."}

Available Point Shapes in ggplot2.
:::

To specify the point shape, you can use the `shape` parameter, as in @fig-pointShape.

```{r}
#| label: fig-pointShape
#| fig-cap: "Specifying Point Shape."
#| fig-alt: "Specifying Point Shape."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name)
) +
  geom_point(
    shape = 2
  )
```

We can allow the point shape to differ by group by specifying the `shape` parameter in the mapping aesthetics, as in @fig-pointShapeByGroup.

```{r}
#| label: fig-pointShapeByGroup
#| fig-cap: "Specifying Point Shape By Group."
#| fig-alt: "Specifying Point Shape By Group."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    shape = player_display_name)
) +
  geom_point()
```

We can specify the point shape manually using the `ggplot2::scale_shape_manual()` function as in @fig-pointShapeByGroupManually.

```{r}
#| label: fig-pointShapeByGroupManually
#| fig-cap: "Specifying Point Shape By Group Manually."
#| fig-alt: "Specifying Point Shape By Group Manually."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    shape = player_display_name)
) +
  geom_point() +
  scale_shape_manual(
    values = c(
      "Tom Brady" = 1,
      "Peyton Manning" = 3)
    )
```

### Size of Points and Text {#sec-ggplot2SyntaxSize}

To specify the size of points and text, you can use the `size` and `base_size` parameters, respectively, as in @fig-sizePointsText.

```{r}
#| label: fig-sizePointsText
#| fig-cap: "Specifying Size of Points and Text."
#| fig-alt: "Specifying Size of Points and Text."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name)
) +
  geom_point(
    size = 2 # size of points
  ) +
  theme_classic(
    base_size = 20 # font size
  )
```

You can also specify the font size separately for particular elements, as in @fig-sizeTextManually.

```{r}
#| label: fig-sizeTextManually
#| fig-cap: "Specifying Size of Text Manually for Particular Elements."
#| fig-alt: "Specifying Size of Text Manually for Particular Elements."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name %in% c("Tom Brady")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_display_name)
) +
  geom_point() +
  theme(
    axis.title.x = element_text(size = 30), # font size for x-axis title
    axis.title.y = element_text(size = 30), # font size for y-axis title
    axis.text.x  = element_text(size = 12), # font size for x-axis labels
    axis.text.y  = element_text(size = 12)  # font size for y-axis labels
  )
```

## Univariate Distribution {#sec-univariateDistribution}

@Scherer2021 describes various ways of visualizing univariate distributions; see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).

### Histogram {#sec-histogram}

A histogram of fantasy points is depicted in @fig-histogram.

```{r}
#| label: fig-histogram
#| fig-cap: "Histogram of Fantasy Points."
#| fig-alt: "Histogram of Fantasy Points."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints)
) +
  geom_histogram(
    color = "#000000",
    fill = "#0099F8"
  ) +
  labs(
    x = "Fantasy Points",
    y = "Count",
    title = "Histogram of Fantasy Points"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Density Plot {#sec-densityPlot}

A histogram of fantasy points is depicted in @fig-density.

```{r}
#| label: fig-density
#| fig-cap: "Density Plot of Fantasy Points by Position."
#| fig-alt: "Density Plot of Fantasy Points by Position."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    fill = position_group)
) +
  geom_density(alpha = 0.7) + # add transparency
  labs(
    x = "Fantasy Points",
    y = "Density",
    fill = "Position",
    title = "Density Plot of Fantasy Points by Position"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Density Plot with Histogram and Rug Plot {#sec-densityHistogram}

A density plot of fantasy points with a histogram and rug plot is depicted in @fig-densityHistogram.

```{r}
#| label: fig-densityHistogram
#| fig-cap: "Density Plot with Histogram and Rug Plot."
#| fig-alt: "Density Plot with Histogram and Rug Plot."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  labs(
    x = "Fantasy Points",
    y = "Density",
    title = "Density Plot of Fantasy Points with Histogram and Rug Plot"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Histogram with Overlaid Density and Rug Plot {#sec-histogramDensity}

A histogram of fantasy points with an overlaid density and rug plot is depicted in @fig-histogramDensity.

```{r}
#| label: fig-histogramDensity
#| fig-cap: "Histogram with Overlaid Density and Rug Plot."
#| fig-alt: "Histogram with Overlaid Density and Rug Plot."

binWidth <- 15

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints)
) +
  geom_histogram(
    color = "#000000",
    fill = "#0099F8",
    binwidth = binWidth
  ) +
  geom_density(
    aes(y = after_stat(count) * binWidth),
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  labs(
    x = "Fantasy Points",
    y = "Count",
    title = "Histogram of Fantasy Points with Overlaid Density and Rug Plot"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Box-and-Whisker Plot {#sec-boxAndWhiskerPlot}

In a box-and-whisker plot, the box is created using the 1st and 3rd quartiles (i.e., the 25th and 75th percentiles, respectively).
The length of the box is equal to the interquartile range, which is calculated as: $\text{IQR} = Q_3 - Q_1$, where $Q_3$ and $Q_1$ are the third and first quartiles, respectively.
The line in the middle of the box is located at the median (i.e., the 2nd quartile or 50th percentile).
The whiskers commonly extend $1.5 \times \text{IQR}$ units from the box.
That is, the upper whisker is commonly located at $1.5 \times \text{IQR}$ units above the third quartile.
The lower whisker is commonly located at $1.5 \times \text{IQR}$ units below the first quartile.
The points represent extreme values (i.e., outliers) that are outside the whiskers.

A box-and-whisker plot of fantasy points is depicted in @fig-boxAndWhisker.

```{r}
#| label: fig-boxAndWhisker
#| fig-cap: "Box-and-Whisker Plot."
#| fig-alt: "Box-and-Whisker Plot."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasyPoints,
    fill = position_group)
) +
  geom_boxplot(staplewidth = 0.25) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Box-and-Whisker Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Violin Plot {#sec-violinPlot}

A violin plot of fantasy points is depicted in @fig-violin.

```{r}
#| label: fig-violin
#| fig-cap: "Violin Plot. Lines represent the 25th, 50th, and 75th quantiles."
#| fig-alt: "Violin Plot. Lines represent the 25th, 50th, and 75th quantiles."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = fantasyPoints,
    fill = position_group)
) +
  geom_violin(
    quantiles = c(0.25, 0.5, 0.75),
    quantile.linetype = "solid"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Violin Plot of Fantasy Points by Position",
    subtitle = "Lines represent the 25th, 50th, and 75th quantiles"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Ridgeline Plot {#sec-ridgelinePlot}

A ridgeline plot of fantasy points is depicted in @fig-ridgeline1 using the `ggridges` package [@R-ggridges].

```{r}
#| label: fig-ridgeline1
#| fig-cap: "Ridgeline Plot."
#| fig-alt: "Ridgeline Plot."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    y = position_group,
    group = position_group,
    fill = position_group)
) +
  ggridges::geom_density_ridges(
    rel_min_height = 0.0085, # remove trailing tails
  ) +
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

We can add lines at the quartiles, as depicted in @fig-ridgeline2.

```{r}
#| label: fig-ridgeline2
#| fig-cap: "Ridgeline Plot. Lines represent the 25th, 50th, and 75th quantiles."
#| fig-alt: "Ridgeline Plot. Lines represent the 25th, 50th, and 75th quantiles."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = fantasyPoints,
    y = position_group,
    group = position_group,
    fill = factor(after_stat(quantile)))
) +
  ggridges::stat_density_ridges(
    rel_min_height = 0.0085, # remove trailing tails
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = 4,
    quantile_lines = TRUE
  ) +
   scale_fill_viridis_d() + # use viridis color scheme
  labs(
    x = "Fantasy Points",
    y = "Position",
    title = "Ridgeline Plot of Fantasy Points by Position",
    subtitle = "Vertical lines represent the 25th, 50th, and 75th quantiles",
    fill = "Quartile"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

## Scatterplot {#sec-scatterplot}

As a tutorial, we walk through some of the (many) modifications that can be made to create an advanced, customized plot in `ggplot2`.

First, we prepare the data:

```{r}
# Subset Data
rb_seasonal <- player_stats_seasonal %>% 
  filter(position_group == "RB")
```

### Base Layer {#sec-scatterplotBaseLayer}

Second, we create the base layer of the plot using the `ggplot2::ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
We specify the data object and the variables in the data object that are associated with the x- and y-axes:

```{r}
#| label: fig-scatterplot1
#| fig-cap: "Base Plot."
#| fig-alt: "Base Plot."

ggplot2::ggplot(
  data = rb_seasonal, # specify data object
  aes(
    x = age, # specify variable on x-axis
    y = rushing_yards)) # specify variable on y-axis
```

### Add Points {#sec-scatterplotAddPoints}

Third, we create a scatterplot using the `ggplot2::geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:

```{r}
#| label: fig-scatterplot2
#| fig-cap: "Scatterplot."
#| fig-alt: "Scatterplot."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() # add points for scatterplot
```

### Best-Fit Line {#sec-scatterplotBestFitLine}

Fourth, we add a linear best-fit line using the `geom_smooth()`, as in @fig-scatterplot3:

```{r}
#| label: fig-scatterplot3
#| fig-cap: "Scatterplot with Linear Best-Fit Line."
#| fig-alt: "Scatterplot with Linear Best-Fit Line."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth(method = "lm") # add linear best-fit line
```

We could also estimate a quadratic polynomial best-fit line, as in @fig-scatterplot4:

```{r}
#| label: fig-scatterplot4
#| fig-cap: "Scatterplot with Quadratic Best-Fit Line."
#| fig-alt: "Scatterplot with Quadratic Best-Fit Line."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ poly(x, 2)) # add quadratic best-fit line
```

Or, we could estimate a smooth best-fit line using locally estimated scatterplot smoothing (LOESS) to allow for any form of nonlinearity, as in @fig-scatterplot5:

```{r}
#| label: fig-scatterplot5
#| fig-cap: "Scatterplot with Best-Fit Line Using Locally Estimated Scatterplot Smoothing (LOESS)."
#| fig-alt: "Scatterplot with Best-Fit Line Using Locally Estimated Scatterplot Smoothing (LOESS)."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth(method = "loess") # add smooth best-fit (LOESS) line
```

By default, the best-fit line is based on a generalized additive model, which allows for nonlinearity, as in @fig-scatterplot6:

```{r}
#| label: fig-scatterplot6
#| fig-cap: "Scatterplot with Best-Fit Line from Generalized Additive Model."
#| fig-alt: "Scatterplot with Best-Fit Line from Generalized Additive Model."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() # add GAM best-fit line; same as specifying method = "gam"
```

### Modify Axes {#sec-scatterplotAxes}

Then, we can change the axes, as in @fig-scatterplot7:

```{r}
#| label: fig-scatterplot7
#| fig-cap: "Scatterplot with Modified Axes."
#| fig-alt: "Scatterplot with Modified Axes."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40), # set limits of x-axis
    ylim = c(0,NA), # set limits of y-axis
    expand = FALSE) + # don't add space between axes and data
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5) # specify x-axis labels
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250) # specify y-axis labels
  )
```

### Plot Labels {#sec-scatterplotPlotLabels}

Then, we can add plot labels, as in @fig-scatterplot8:

```{r}
#| label: fig-scatterplot8
#| fig-cap: "Scatterplot with Plot Labels."
#| fig-alt: "Scatterplot with Plot Labels."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA),
    expand = FALSE) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs( # add plot labels
    x = "Running Back's Age (years)", # x-axis title
    y = "Rushing Yards (Season)", # y-axis title
    title = "NFL Rushing Yards (Season) by Player Age", # plot title
    subtitle = "(Among Running Backs)" # plot subtitle
  )
```

### Theme {#sec-scatterplotTheme}

Then, we can use a theme such as the classic theme (`theme_classic()`) to make it more visually presentable, as in @fig-scatterplot9:

```{r}
#| label: fig-scatterplot9
#| fig-cap: "Scatterplot with Classic Theme."
#| fig-alt: "Scatterplot with Classic Theme."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA),
    expand = FALSE) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic() # use the classic theme
```

Or, we could use a different theme, such as the dark theme (`theme_dark()`) in @fig-scatterplot10.
For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes> [@ggplot3rdEd].

```{r}
#| label: fig-scatterplot10
#| fig-cap: "Scatterplot with Dark Theme."
#| fig-alt: "Scatterplot with Dark Theme."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA),
    expand = FALSE) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_dark() # use the dark theme
```

### Interactive {#sec-scatterplotInteractive}

::: {.content-visible when-format="html:js"}

After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.

```{r}
#| label: fig-scatterplot11
#| fig-cap: "Interactive Scatterplot Using Plotly."
#| fig-alt: "Interactive Scatterplot Using Plotly."

plot_ypcByPlayerAge <- ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season)) + # add season for mouse over tooltip
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA),
    expand = FALSE) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic()

plotly::ggplotly(plot_ypcByPlayerAge)
```

:::

::: {.content-visible when-format="pdf"}

After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].

```{r}
#| eval: false

plot_ypcByPlayerAge <- ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season)) + # add season for mouse over tooltip
  geom_smooth() +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA),
    expand = FALSE) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic()

ggplotly(plot_ypcByPlayerAge)
```

:::

### Scatterplot with Rug Plot {#sec-scatterplotRugPlot}

We can also add a rug plot, as in @fig-scatterplot12.

```{r}
#| label: fig-scatterplot12
#| fig-cap: "Scatterplot With Rug Plot."
#| fig-alt: "Scatterplot With Rug Plot."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  geom_rug(
    color = "#800000",
    alpha = 0.2) +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA)) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic()
```

### Marginal Density/Histogram {#sec-scatterplotMarginalDensigram}

We can also add a marginal density/histogram, as in @fig-scatterplot13 using the `ggExtra` package [@R-ggExtra]:

```{r}
#| label: fig-scatterplot13
#| fig-cap: "Scatterplot With Marginal Density/Histogram."
#| fig-alt: "Scatterplot With Marginal Density/Histogram."

ggExtra::ggMarginal(
  plot_ypcByPlayerAge,
  type = "densigram")
```

Or we can create the plots manually, as in @fig-scatterplot14 using the `patchwork` package [@R-patchwork]:

```{r}
#| label: fig-scatterplot14
#| fig-cap: "Scatterplot With Marginal Density/Histogram."
#| fig-alt: "Scatterplot With Marginal Density/Histogram."

dens1 <- ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age)) +
  coord_cartesian(
    xlim = c(20,40),
    expand = FALSE) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  theme_void()

dens2 <- ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = rushing_yards)) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  theme_void() +
  coord_flip(
    ylim = c(0, NA),
    expand = FALSE)

dens1 + patchwork::plot_spacer() + plot_ypcByPlayerAge + dens2 + 
  patchwork::plot_layout(
    ncol = 2, 
    nrow = 2, 
    widths = c(4, 1),
    heights = c(1, 4)
  )
```

## Line Chart {#sec-lineChart}

A bar plot of Tom Brady's fantasy points by season is depicted in @fig-lineChart.

```{r}
#| label: fig-lineChart
#| fig-cap: "Line Chart."
#| fig-alt: "Line Chart."

ggplot2::ggplot(
  data = player_stats_seasonal %>%
    filter(player_display_name == "Tom Brady"),
  mapping = aes(
    x = season,
    y = fantasyPoints
    )
) +
  geom_line(
    linewidth = 1.5,
    color = "blue"
  ) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Bar Plot of Tom Brady's Fantasy Points by Season"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### With Highlighting {#sec-lineChartHighlighting}

We use the `gghighlight` package [@R-gghighlight] to highlight particular elements:

```{r}
#| label: fig-lineChartHighlighting
#| fig-cap: "Line Chart with Highlighting."
#| fig-alt: "Line Chart with Highlighting."

ggplot2::ggplot(
  data = player_stats_seasonal %>% 
    filter(position_group %in% c("QB")),
  mapping = aes(
    x = season,
    y = fantasyPoints,
    group = player_id,
    color = player_display_name)
) +
  geom_line(linewidth = 2) +
  gghighlight::gghighlight(
    player_display_name == "Tom Brady",
    label_key = player_display_name,
    unhighlighted_params = list(linewidth = 0.5)) +
  labs(
    x = "Season",
    y = "Fantasy Points",
    title = "Fantasy Points by Season and Player",
    subtitle = "(Tom Brady in Red)"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

## Bar Plot {#sec-barPlot}

To create a bar plot, we first compute summary statistics:

```{r}
confidenceLevel <- .95 # for 95% confidence interval

player_stats_seasonal_offense_summary <- player_stats_seasonal %>%
    filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  group_by(position_group) %>%
  summarise( 
    n = sum(!is.na(fantasyPoints)),
    mean = mean(fantasyPoints, na.rm = TRUE),
    sd = sd(fantasyPoints, na.rm = TRUE)
  ) %>%
  mutate(se = sd/sqrt(n)) %>%
  mutate(
    ci_lower = mean - qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
    ci_upper = mean + qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se
  )
```

The summary statistics are in @tbl-summaryStats.

```{r}
#| label: tbl-summaryStats
#| tbl-cap: "Table of Summary Statistics."

player_stats_seasonal_offense_summary
```

A bar plot of fantasy points by position is depicted in @fig-barPlot.

```{r}
#| label: fig-barPlot
#| fig-cap: "Bar Plot."
#| fig-alt: "Bar Plot."

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary,
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### With Error Bars {#sec-barPlotErrorBars}

Based on the summary statistics in @tbl-summaryStats, we create a bar plot with bars representing the 95% confidence interval in @fig-barPlotConfidenceInterval.

```{r}
#| label: fig-barPlotConfidenceInterval
#| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."
#| fig-alt: "Bar Plot with Bars Representing the 95% Confidence Interval."

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

### Modified Color Scheme {#sec-barPlotModifiedColorScheme}

We can also modify the color scheme, as in @fig-barPlotColorScheme

```{r}
#| label: fig-barPlotColorScheme
#| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."
#| fig-alt: "Bar Plot with Bars Representing the 95% Confidence Interval."

ggplot2::ggplot(
  data = player_stats_seasonal_offense_summary %>%
    filter(position_group %in% c("QB","RB","WR","TE")),
  mapping = aes(
    x = position_group,
    y = mean,
    fill = position_group
    )
) +
  geom_bar(
    stat = "identity") +
  scale_fill_brewer(palette = "Dark2") +
  geom_errorbar(
    aes(
      ymin = ci_lower,
      ymax = ci_upper),
    width = 0.2,
    color = "black"
  ) +
  labs(
    x = "Position",
    y = "Fantasy Points",
    title = "Bar Plot of Fantasy Points by Position"
  ) +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
```

## Faceting {#sec-faceting}

```{r}
#| label: fig-facetedScatterplot
#| fig-cap: "Faceted Scatterplot."
#| fig-alt: "Faceted Scatterplot."

ggplot2::ggplot(
  data = player_stats_seasonal %>% 
    filter(position_group %in% c("QB","RB","WR","TE")),
  aes(
    x = age,
    y = fantasyPoints)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(
    ylim = c(0,NA), # set limits of y-axis
    expand = FALSE) + # don't add space between axes and data
  labs(
    x = "Player's Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age"
  ) +
  theme_bw() +
  facet_wrap(vars(position_group)) # facet by position_group
```

## Examples {#sec-scatterplotExamples}

### Players {#sec-scatterplotPlayers}

#### Running Back Performance By Player Age {#sec-plotRBperformanceByAge}

```{r}
# Prepare Data
rushing_attempts <- nfl_pbp %>% 
  dplyr::filter(season_type == "REG") %>% 
  dplyr::filter(
    rush == 1,
    rush_attempt == 1,
    qb_scramble == 0,
    qb_dropback == 0,
    !is.na(rushing_yards))

rb_yardsPerCarry <- rushing_attempts %>% 
  dplyr::group_by(rusher_id, season) %>% 
  dplyr::summarise(
    ypc = mean(rushing_yards, na.rm = TRUE),
    rush_attempts = n(),
    .groups = "drop") %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(
    player_stats_seasonal,
    by = c("rusher_id" = "player_id", "season")
  ) %>% 
  dplyr::filter(
    position_group == "RB",
    rush_attempts >= 50)
```

##### Rushing Yards Per Carry {#sec-plotYPCbyAge}

::: {.content-visible when-format="html:js"}

Rushing yards per carry over the course of the season is depicted as a function of the Running Back's age in @fig-rushYPCperSeasonByAge.

```{r}
#| label: fig-rushYPCperSeasonByAge
#| fig-cap: "Rushing Yards Per Carry (Season) by Player Age."
#| fig-alt: "Rushing Yards Per Carry (Season) by Player Age."

plot_ypcByPlayerAge2 <- ggplot2::ggplot(
  data = rb_yardsPerCarry,
  aes(
    x = age,
    y = ypc)) +
  geom_point(
    aes(
      text = player_display_name,
      label = season)) +
  geom_smooth() +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards Per Carry (Season)",
    title = "NFL Rushing Yards Per Carry (Season) by Player Age",
    subtitle = "(minimum 50 rushing attempts)"
  ) +
  theme_classic()

ggplotly(plot_ypcByPlayerAge2)
```

:::

::: {.content-visible when-format="pdf"}

Rushing yards per carry over the course of the season is depicted as a function of the Running Back's age in @fig-rushYPCperSeasonByAgePDF.

```{r}
#| label: fig-rushYPCperSeasonByAgePDF
#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age."
#| fig-alt: "NFL Rushing Yards Per Carry Per Season by Player Age."

ggplot2::ggplot(
  data = rb_yardsPerCarry,
  aes(
    x = age,
    y = ypc)) +
  geom_point(
    aes(
      text = player_display_name,
      label = season)) +
  geom_smooth() +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards Per Carry (Season)",
    title = "NFL Rushing Yards Per Carry (Season) by Player Age",
    subtitle = "(minimum 50 rushing attempts)"
  ) +
  theme_classic()
```

:::

##### Rushing EPA Per Season {#sec-plotRushEPAbyAge}

::: {.content-visible when-format="html:js"}

Rushing expected points added (EPA) over the course of the season is depicted as a function of the Running Back's age in @fig-rushEPAperSeasonByAge.

```{r}
#| label: fig-rushEPAperSeasonByAge
#| fig-cap: "Rushing Expected Points Added (Season) by Player Age."
#| fig-alt: "Rushing Expected Points Added (Season) by Player Age."

plot_rushEPAbyPlayerAge <- ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_epa)) +
  geom_point(
    aes(
      text = player_display_name,
      label = season)) +
  geom_smooth() +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing EPA (Season)",
    title = "NFL Rushing Expected Points Added (Season) by Player Age"
  ) +
  theme_classic()

ggplotly(plot_rushEPAbyPlayerAge)
```

:::

::: {.content-visible when-format="pdf"}

Rushing expected points added (EPA) over the course of the season is depicted as a function of the Running Back's age in @fig-rushEPAperSeasonByAgePDF.

```{r}
#| label: fig-rushEPAperSeasonByAgePDF
#| fig-cap: "NFL Rushing Expected Points Added (EPA) During a Season by Player Age."
#| fig-alt: "NFL Rushing Expected Points Added (EPA) During a Season by Player Age."

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_epa)) +
  geom_point(
    aes(
      text = player_display_name,
      label = season)) +
  geom_smooth() +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing EPA (Season)",
    title = "NFL Rushing Expected Points Added (Season) by Player Age"
  ) +
  theme_classic()
```

:::

### Teams {#sec-scatterplotTeams}

#### Defensive and Offensive EPA per Play {#sec-plotDefOffEPA}

Expected points added (EPA) per play by the team with possession.

```{r}
pbp_regularSeason <- nfl_pbp %>% 
  dplyr::filter(
    season == 2025,
    season_type == "REG") %>%
  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))

epa_offense <- pbp_regularSeason %>%
  dplyr::group_by(team = posteam) %>%
  dplyr::summarise(off_epa = mean(epa, na.rm = TRUE))

epa_defense <- pbp_regularSeason %>%
  dplyr::group_by(team = defteam) %>%
  dplyr::summarise(def_epa = mean(epa, na.rm = TRUE))

epa_combined <- epa_offense %>%
  dplyr::inner_join(
    epa_defense,
    by = "team")
```

Defensive EPA per play during the 2025 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA using the `nflplotR` package [@R-nflplotR].

```{r}
#| label: fig-plotDefOffEPA
#| fig-cap: "2025 NFL Offensive and Defensive EPA Per Play."
#| fig-alt: "2025 NFL Offensive and Defensive EPA Per Play."

ggplot2::ggplot(
  data = epa_combined,
  aes(
    x = off_epa,
    y = def_epa)) +
  nflplotR::geom_mean_lines(
    aes(
      x0 = off_epa,
      y0 = def_epa)) +
  nflplotR::geom_nfl_logos(
    aes(
      team_abbr = team),
      width = 0.065,
      alpha = 0.7) +
  labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    title = "2025 NFL Offensive and Defensive EPA per Play"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
  scale_y_reverse()
```

## Conclusion {#sec-dataVisualizationConclusion}

This chapter provided an overview of principles of data visualization and how to implement them using `R`.
Key principles of data visualization include: ensuring adequate contrast, highlighting key elements of interest (and using gray for other elements), reducing clutter, and adding helpful titles and annotations.
We provided examples of data visualizations, including [histograms](#sec-histogram), [density plots](#sec-densityPlot), [rug plots](#sec-densityHistogram), [box-and-whisker plots](#sec-boxAndWhiskerPlot), [violin plots](#sec-violinPlot), [ridgeline plots](#sec-ridgelinePlot), [scatterplots](#sec-scatterplot), [line charts](#sec-lineChart), and [bar plots](#sec-barPlot).

::: {.content-visible when-format="html"}

## Session Info {#sec-dataVisualizationSessionInfo}

```{r}
sessionInfo()
```

:::